### 2020-03-06 Vanella_Article_Atmosphere_2020.tex
Initial commit of manuscript

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[journal,article,atmosphere,submit,moreauthors,pdftex]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, animals, antibiotics, antibodies, antioxidants, applmech, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejbc, ejihpe, electrochem, electronics, endocrines, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, oceans, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, sustainableworld, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2020}
\copyrightyear{2020}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Modeling Terrain using an Immersed Boundary Method}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Marcos Vanella $^1$, Kevin McGrattan $^1$, Randall McDermott $^1$, Glenn Forney $^1$, and William Mell $^2$}

% Authors, for metadata in PDF
\AuthorNames{Marcos Vanella, Kevin McGrattan, Randall McDermott, Glenn Forney, and William Mell}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad National Institute of Standards and Technology, Gaithersburg, Maryland, USA \\
$^{2}$ \quad U.S. Forest Service, Seattle, Washington, USA}

% Contact information of the corresponding author
\corres{Correspondence: marcos.vanella@nist.gov}

\abstract{A method for simulating the spread of wildfire over complex terrain is presented. An immersed boundary method (IBM) is used to render the complex terrain on a rectilinear Cartesian grid. Fire spread is modeled using one of three basic approaches: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set technique in which the fire spreads as a function of terrain, vegetation type, and wind speed.}

\keyword{complex terrain; fire spread; immersed boundary method; level sets}

\begin{document}

\section{Introduction}

The work of Mell et al.~\cite{Mell:IJWF2007}.

\section{The Mean Forcing Concept}

The process of steering the solution of the mass, momentum, and energy equations to match wind speed and direction gathered at random weather stations is known as \emph{data assimilation}~\cite{Kalnay:2003}. FDS uses a relatively simple data assimilation technique, a method known as \emph{nudging}, where you add a forcing term to the momentum equation to ``nudge'' the flow in the direction of the specified wind. FDS automatically drives the mean velocity components toward the desired values by adding a forcing term to the momentum equation. For example, the $u$ component equation is modified as follows:
\begin{equation}
   \frac{\partial u}{\partial t} + \ldots = \frac{u_0(z,t)-\overline{u}(z,t)}{\tau} \label{mean_forcing_u}
\end{equation}
Here, $u(\mathbf{x},t)$ is the computed velocity component, $u_0$ is the specified wind field component that can vary with height and time, and $\overline{u}$ is the average of the computed velocity component at the height $z$. The relaxation time scale, $\tau$, is an important parameter. The shorter this time scale, the faster the flow field will ``catch up'' to the mean wind, but at the expense of possibly washing out important flow structures. There is no firm guidance in the atmospheric modeling literature.

\section{Wildland Fire Spread}

There are three ways of simulating wildland fire spread.

\begin{enumerate}
\item {\bf Particle Method:} The vegetation is represented by a collection of Lagrangian particles that are heated via convection and radiation.
\item {\bf Boundary Fuel Model:} Ground vegetation is modeled like a porous solid with a thickness equal to the height of the vegetation.
\item {\bf Level Set Method:} The fire front propagates using purely empirical rules.
\end{enumerate}
The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels and the fire spread rate is {\em predicted} by the model. The Level Set Method relies on a set of experimentally-determined spread rates for different types of vegetation and wind speeds.

\reftitle{References}

\externalbibliography{yes}
\bibliography{Vanella_Article_Atmosphere_2020}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2020-03-06 Vanella_Article_Atmosphere_2020.tex
Atmosphere .tex : My changes to Title, abstract.

@@
 % Full title of the paper (Capitalized)
-\Title{Modeling Terrain using an Immersed Boundary Method}
+\Title{Application of an immersed boundary technique to wildfire simulation over complex terrains}
 
@@
 
-\abstract{A method for simulating the spread of wildfire over complex terrain is presented.
-An immersed boundary method (IBM) is used to render the complex terrain on a rectilinear Cartesian grid.
-Fire spread is modeled using one of three basic approaches: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set technique in which the fire spreads as a function of terrain, vegetation type, and wind speed.}
+\abstract{A method for the large eddy simulation (LES) of wildfire spread over complex terrains is presented.
+In this scheme, a cut-cell immersed boundary method (IBM) is used to render the complex terrain, defined by a tessellation, on a rectilinear Cartesian grid.
+Discretization of scalar transport equations for chemical species is done via a finite volume scheme on cut-cells defined by the intersection of the terrain and Cartesian cells.
+Momentum transport and heat transfer close to the immersed terrain are handled using a dynamic wall model and direct forcing IBM.
+Further, atmospheric wind conditions are specified using a mean wind forcing concept.
+Fire spread is modeled using one of three basic approaches: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set technique in which the fire spreads as a function of terrain, vegetation type, and wind speed.
+Several test and validation cases are reported to demonstrate the capabilities of this novel wildfire simulation methodology.}
 

### 2020-03-09 Vanella_Article_Atmosphere_2020.tex
Change title

--- 2020-03-06 Vanella_Article_Atmosphere_2020.tex
@@
 % Full title of the paper (Capitalized)
-\Title{Modeling Terrain using an Immersed Boundary Method}
+\Title{Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over Complex Terrain}
 
@@
 
-\abstract{A method for simulating the spread of wildfire over complex terrain is presented.
-An immersed boundary method (IBM) is used to render the complex terrain on a rectilinear Cartesian grid.
-Fire spread is modeled using one of three basic approaches: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set technique in which the fire spreads as a function of terrain, vegetation type, and wind speed.}
+\abstract{A method for the large-eddy simulation (LES) of wildfire spread over complex terrain is presented.
+In this scheme, a cut-cell immersed boundary method (CC-IBM) is used to render the complex terrain, defined by a tessellation, on a rectilinear Cartesian grid.
+Discretization of scalar transport equations for chemical species is done via a finite volume scheme on cut-cells defined by the intersection of the terrain geometry and the Cartesian cells.
+Momentum transport and heat transfer close to the immersed terrain are handled using dynamic wall models and a direct forcing immersed boundary method.
+Further, atmospheric wind conditions are specified using a mean forcing concept.
+Fire spread is modeled using one of three basic approaches: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set method in which the fire spreads as a function of terrain slope, vegetation type, and wind speed.
+Several test and validation cases are reported to demonstrate the capabilities of this novel wildfire simulation methodology.}
 

### 2020-06-03 Vanella_Article_Atmosphere_2020.tex

@@
 
-The work of Mell et al.~\cite{Mell:IJWF2007}.
+% Motivation.
+In the last decades, wild land and wild land-urban interface (WUI) fires have received substantial attention due to their destructive nature and extensive cost in lives and property~\cite{thomas_2017,mcdermott_2019,boghani_2019,richards_2020}.
+Over the years, modeling and simulation activities of such phenomena have been tailored to the needs of planners, scientists and first responders, and by the available computing capabilities.
+% Wildfire modeling in the literature.
+A large body of literature can be found on the subject.
+See for example reviews~\cite{Papadopoulos_2011,Bakhshaii_2019,mcdermott_2019} on classical and new models with associated software.
+Due to inherent complex nature of the problem, most models of wildfires have historically relied on simplified field-tested rules and correlations.
 
-\section{The Mean Forcing Concept}
+Among these, the fire spread model by Rothermel~\cite{Rothermel:1972} assumes quasi-steady state ground fire conditions and takes as inputs fuel properties, wind, terrain slope and moisture that can be measured in situ.
+Fuel models have been cataloged~\cite{Anderson:1982}, and their input parameters defined in a manner the user can select and combine the most appropriate listed fuels for her case of study.
+Rothermels fire line model is a widely used wildfire spread model, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013} \textbf{OTHERS?}, enabling a range of planning and operational forecasting capabilities.
+
+% Level set as a simplified fire spread model, operational forecasting.
+Evolution of wild land fires is usually tied to a fire front, which can have a complex shape moving/deforming with the local flame spread velocities.
+Knowing the location of this fire front and local properties of the fire, fuel and terrain allows, by use of simplified models like Rothermels, to predict local ignition and fire spread.
+When modeling the problem numerically, identifying such fire front at a given moment in time requires a mathematical representation and an associated discretization or mesh.
+The Lagrangian treatment of the motion of a fire front requires specifying a set of control markers and connectivity that define and move with such front.
+Therefore, the fire front is represented by this discrete mesh.
+For example, if the domain of motion is a surface terrain, this Lagrangian mesh can be a polygon that moves with the front along this surface~\cite{Finney:FARSITE,Bova:IJWF2015}.
+On the other hand, an Eulerian representation of motion allows for the mesh to be fixed in space (i.e.
+in a terrain a polygonal patch mesh of said terrain).
+Then, the location of the fire front can be portrayed implicitly by using a scalar field defined in the fixed two dimensional mesh.
+This scalar field is called level set function and numerical methods that evolve it in time are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
+Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}.
+A practical comparison of the two techniques as implemented in WFDS~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.
+
+% Resolving the flow features on terrains using LES over complex terrains. Describe what we do, FDS, cut-cell schemes, immersed boundary methods.
+As the rate of spread of the front line is dependent in the local velocity on empirical fire spread models like Rothermels, a better local resolution of the velocity field around complex terrains should be beneficial to the accuracy of the overall model.
+In reality, besides the obstructive character of the terrain itself, compounded effects due to the fire affect the local velocity field.
+\textbf{Some of these are ...}.
+The de facto physics-based technique for practical simulation of outdoor flows is Large Eddy Simulation (LES)~\cite{coen_2013,mcgratta_2013}.
+Within FDS~\cite{mcgratta_2013},  the filtered Navier-Stokes equations for buoyant, combusting flows are evolved using a Low Mach approximation.
+Chemical species are handled using scalar transport equations.
+Energy transport is accounted for by means of the solution of radiative transport equations and a thermodynamic divergence proxy.
+\textbf{Comment on WFDS for LOF.
+The work of Mell et al.~\cite{Mell:IJWF2007}.}.
+
+% Unstructured geometries, capability can be tuned for use with LOF and fire spread models.
+% What has been done, cut-cell scheme IBM, connection to LS and Rothermels, wind.
+Lately, substantial work has been done to add to FDS the option of simulating flows around unstructured geometries.
+These are represented by surface triangulations which is a typical representation in computer aided design (CAD) and simulation software.
+This added flexibility is still compliant with the structured cartesian grids and differentiation schemes within FDS.
+We implemented a cut-cell scheme~\cite{berger_2016} to advance the scalar transport and energy of multicomponent mixtures on the polyhedra next to the geometries.
+These polyhedrons arise from geometry intersection with the cartesian fluid grid.
+As geometry boundaries are not changed, the method has the property of exact volume and area conservation.
+Velocities next to geometries are reconstructed using an immersed boundary method (IBM)~\cite{fadlun_2000,balaras_2004} coupled with an equilibrium wall model for turbulent flows~\cite{mcdermo_2018}.
+Variables related to IBM forcing and heat transfer modeling at the wall are sampled through interpolation and a normal probe approach~\cite{balaras_2004}.
+Also, radiation transport boundary conditions have been adapted to this technique.
+For fire simulation over complex terrains a range of options in terms of physics fidelity of the model are being developed, in particular, in the representation of fire spread.
+For fast simulation of outdoor fires, this new capability has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and Rothermels fire spread model.
+The wind speed input for Rothermels model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution.
+These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept.
+
+% Outline of the manuscript.
+In the following section a description on the mathematical model for fire over terrain is provided.
+The terrain description and model discretization is provided in section~\ref{sec:terraindisc}.
+An account of the fire spread model is given in section~\ref{sec:firespread}, as well as the wind atmospheric boundary conditions in section~\ref{sec:wind}.
+Numerical experiments testing the methodology on a simple example and a real case of fire in Cogoleto, Italy are provided in section~\ref{sec:numexp}.
+Finally a summary of this work is outlined in section~\ref{sec:summary}.
+
+\section{Mathematical model} \label{sec:matmodel}
+
+% Model equations, fluid side : scalar transport, momentum, energy. Stratification.
+
+% Combustion, fuel, fire spread model equations.
+
+\section{Terrain description and discretization} \label{sec:terraindisc}
+
+% How do we describe terrains - geometries. Methodology to get from DEM a geometry.
+
+% What FDS does for Scalar Transport - 1 paragraph Modification close to the terrain surface: scalar transport, momentum, divergence.
+
+% Connection with level set fire spread.
+
+\section{Wildland Fire Spread} \label{sec:firespread}
+
+There are three ways of simulating wildland fire spread.
+
+\begin{enumerate}
+\item {\bf Particle Method:} The vegetation is represented by a collection of Lagrangian particles that are heated via convection and radiation.
+\item {\bf Boundary Fuel Model:} Ground vegetation is modeled like a porous solid with a thickness equal to the height of the vegetation.
+\item {\bf Level Set Method:} The fire front propagates using purely empirical rules.
+\end{enumerate}
+The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels and the fire spread rate is {\em predicted} by the model.
+The Level Set Method relies on a set of experimentally-determined spread rates for different types of vegetation and wind speeds.
+
+% Here, methodology to get fuels in the terrain from online sources.
+
+% Defining atmospheric conditions:
+\section{Defining atmospheric wind : The Mean Forcing Concept} \label{sec:wind}
 
@@
 
-\section{Wildland Fire Spread}
+\section{Numerical Experiments} \label{sec:numexp}
 
-There are three ways of simulating wildland fire spread.
+\subsection{Simple example}  \label{sec:simexp}
 
-\begin{enumerate}
-\item {\bf Particle Method:} The vegetation is represented by a collection of Lagrangian particles that are heated via convection and radiation.
-\item {\bf Boundary Fuel Model:} Ground vegetation is modeled like a porous solid with a thickness equal to the height of the vegetation.
-\item {\bf Level Set Method:} The fire front propagates using purely empirical rules.
-\end{enumerate}
-The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels and the fire spread rate is {\em predicted} by the model.
-The Level Set Method relies on a set of experimentally-determined spread rates for different types of vegetation and wind speeds.
+\subsection{Cogoleto Fire}  \label{sec:cogo}
+
+\section{Summary} \label{sec:summary}
 

### 2020-06-09 Vanella_Article_Atmosphere_2020.tex

@@
 
-% Model equations, fluid side : scalar transport, momentum, energy. Stratification.
+For fire modeling on the gas phase, FDS employs a Low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}.
+Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\overline{p}(z,t)$ used on the equation of state (the ideal gas law is used), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.
 
-% Combustion, fuel, fire spread model equations.
+Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$.
+If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, were $\rho_\alpha(\mathbf{x},t)$ is the species density for the local thermodynamic conditions.
+The scalar transport and momentum equations take the form:
+\begin{eqnarray}
+   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \mathbf{J_{c}}_\alpha ) &=& - \nabla \cdot (\mathbf{J_{d}}_\alpha) + \dot{m}_\alpha'''  + 
+    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
+    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
+    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau} \right] \label{eqn:momtran}
+\end{eqnarray}
+where $\mathbf{J_{c}}_\alpha=\rho Y_\alpha  \mathbf{u}$, $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ are the convective and diffusive fluxes for component $\alpha$.
+Ficks Law for binary diffusion respect to a background species is assumed, $D_\alpha$ is the diffusivity of $\alpha$ respect to the background species, $\dot{m}_\alpha'''(\mathbf{x},t)$ is the combustion volumetric source for species $\alpha$ and $ \dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from sub-grid particles gasification.
+Also, in the momentum equation the head $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, and $\tilde{p}$ is the perturbation pressure, $\rho_0(z)$ is an equlibrium density, $\mathbf{g}=(0,0,g_z)$ gravity vector, $\mathbf{f}_{b}$ is the particle force term and $\boldsymbol{\tau}$ tensor accounting for molecular and sub-grid turbulent stresses parameterized by an eddy viscosity model.
+This last model is employed by default in large eddy simulation (LES) of outdoor fires.
+Details of the model equations and their derivation are found in reference~\cite{mcgratta_2013tr}.
+
+Further, the flow field is subjected to the following divergence constraint:
+%
+\begin{eqnarray}
+    ( \nabla \cdot \mathbf{u} )^{th} &=& 
+    \left[ \frac{1}{\rho c_p T} - \frac{1}{\bar{p}} \right] 
+    \frac{\partial \bar{p}}{\partial t} + \frac{w \rho_0 g_z}{\rho c_p T} \nonumber \\
+    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
+    &+& \frac{1}{\rho} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right) \left[ \dot{m}_\alpha'''  + \dot{m}_{b,\alpha}''' - \nabla \cdot (\mathbf{J_{d}}_\alpha) - \mathbf{u} \cdot \nabla (\rho Y_\alpha) \right] \label{eq:divth}
+\end{eqnarray}
+%
+where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux, sum of conduction, convection and radiation, and $T$ is the local temperature.
+The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively, built from the individual species counterparts as defined.
+Finally, $w,\rho_0,g_z$ on the stratification term are respectively, the local vertical velocity, density for standard conditions (depends on vertical coordinate), and vertical gravity acceleration.
+This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}.
+Being $R$ the ideal gas constant, problem is closed with the ideal gas law $ \overline{p} = {\rho R T}/{\overline{W}}$.
+
+% Fire spread model equations.
+As commented in the introduction, the location of a fire front can be defined implicitly along the terrain surface through the zero of a two dimensional level set scalar function $\phi(x,y,t)$ in the form $\phi(x,y,t)=0$.
+This implicit form can be used to define and advection equation for the level set function:
+%
+\begin{equation}
+   \frac{D \phi(x,y,t)}{D t} =  \frac{\partial \phi}{\partial t} + \mathbf{R}_{uv} \cdot \nabla (\phi) = 0 \label{eqn:lset}
+\end{equation}
+%
+where $\mathbf{R}_{uv}=(R_u,R_v)$ is the rate of spread vector along the terrain plane.
+The magnitude of the rate of spread vector is computed by $|\mathbf{R}_{uv}|=R_0 \left(1 + \sqrt{(\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) \cdot (\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) } \right)$, being $R_0$ the zero velocity and zero slope rate of spread (baseline, dependent of fuel properties), and vectors $\boldsymbol{\phi}_W,\boldsymbol{\phi}_S \ge 0$ are computed using local wind, slope through empirical rules~\cite{Wilson:1980}.
+See Appendix 1 of reference~\cite{Bova:IJWF2015} for details on these calculations, as implemented in WFDS-LS and in this work.
 
@@
 % How do we describe terrains - geometries. Methodology to get from DEM a geometry.
+%
+In figure~\ref{Fig:figure_1}a, a terrain is represented by its surface triangulation within an FDS cartesian mesh.
+Some cartesian cells and faces belonging to this mesh are transversed by the terrain surface.
+The remainder polyhedra and polygons that lay on the gas side of this intersected cartesian entities are called cut-cells and cut-faces.
+They discretize the fluid domain in the region surrounding the solid terrain.
+A detail of this unstructured mesh is.
+shown in figure~\ref{Fig:figure_1}b.
+The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where more than one cut-cell per cartesian cell are not uncommon.
+As described later, small cells pose a temporal stability constraint in explicit time integration schemes that needs to be addressed.
+It is noted that a two level of refinement grid hierarchy emerges.
+The coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured counterparts.
+Methods that solve discrete model equations on these grids are called cut-cell or embedded boundary methods~\cite{berger_2016}.
+Reliably defining the cut-cell mesh geometric properties and topology is a complex problem on itself, omitted here in order to keep the discussion succinct.
+%
+\begin{figure}[h]
+   \centering
+   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{./figures/SketchFig1.pdf} 
+   \put(-305,-5){(a)}    
+   \put(-95,-4){(b)}
+   \caption{Sketch of regular Cartesian and cut-cell grids around terrain: (a) An unstructured cut-cell region is defined on the gas phase side around a terrain $\mathbf{T}$.
+(b) Detail showing a slice representative of cut-cells and regular gas cells, terrain triangulation and two refinement level interpretation.}
+   \label{Fig:figure_1}
+\end{figure}  
+%
 
-% What FDS does for Scalar Transport - 1 paragraph Modification close to the terrain surface: scalar transport, momentum, divergence.
+\subsection{Scalar transport and energy discretization in unstructured region}
 
-% Connection with level set fire spread.
+The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}.
+See figure~\ref{Fig:figure_2}a.
+Integrating equation~\eqref{eqn:spectran} over a cut-cell $ii$ with control volume $\Omega_{ii}$:
+%
+\begin{equation}
+ \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
+      } d \Omega  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega + \int_{\Omega_{ii}} {( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''') } d \Omega \label{eq:intconvdiff}
+\end{equation}
+%
+For a control volume that does not change with time, the time derivative and source terms are approximated by
+%
+\begin{eqnarray}
+  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega & = & \frac{\partial}{\partial t} \int_{\Omega_{ii}} {\rho Y_\alpha} d \Omega
+  = \frac{\partial \widetilde{\rho \: Y_\alpha }_{ii}}{\partial t} V_{ii} \\
+  \int_{\Omega_{ii}} {( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''')} d \Omega & = & (\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii}) V_{ii} \label{eq:intcons}
+\end{eqnarray}
+%
+where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages.
+In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged.
+Consider the FV discretization of the diffusive term (eqn.~\eqref{eq:intconvdiff}) on cut-cell $ii$ of figure~\ref{Fig:figure_2}a:
+%
+\begin{equation}
+\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega =
+    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega = \sum^{nf_c}_{k=1}
+    \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvdiffcc}
+\end{equation}
+%
+The integral over the cut-cell volume has been transformed in an area integral on its $nf_c=5$ faces using the divergence theorem.
+As these $k$ faces with areas $A_k$ and normals out of the control volume $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization.
+A centroid to centroid (i.e.
+$\Delta x_{cc}$ in figure~\ref{Fig:figure_2}a) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase.
+Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces.
+%
+\begin{figure}[h]
+   \centering
+   \includegraphics[trim = 30mm 42mm 35mm 45mm, clip,width=1\linewidth]{./figures/SketchFig2.pdf} 
+   \put(-335,-5){(a)}    
+   \put(-95,-4){(b)}
+   \caption{Sketch of cut-cell and faces: (a) Cut-cell $ii$ surrounded by gas phase regular and cut-faces (3-5), and boundary cut-faces (1-2).
+(b) Interpolation sketch for wall modeled immersed boundary reconstruction of cut-face velocities.}
+   \label{Fig:figure_2}
+\end{figure}  
+%
+
+Similarly, the discretization of the advective term is:
+%
+\begin{equation}
+    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} d \Omega =
+     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega =
+     \sum^{nf_c}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvadvcc}
+\end{equation}
+%
+where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux limited interpolation to the cut-face~\cite{mcgratta_2013}.
+In the cut-cell region we use Godunov interpolation for the advective term.
+As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of equations~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known.
+The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:
+%
+%
+\begin{eqnarray}
+    ( \nabla \cdot \mathbf{u} )_{ii}^{th} \; V_{ii} &=&
+    \left[ \frac{1}{(\rho c_p T)_{ii}} - \frac{1}{\bar{p}_{ii}} \right]
+    \frac{\partial \bar{p}_{ii}}{\partial t} \; V_{ii} +
+    \frac{w_{ii} \: \rho_0 g_z}{(\rho c_p T)_{ii}} V_{ii} \nonumber \\
+    &+& \frac{1}{(\rho c_p T)_{ii}} \left[ (\dot{q}'''+ \dot{q}_b''')_{ii} V_{ii} -
+    \sum_{k=1}^{nf_c} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
+    - \overline{\mathbf{u} \cdot \nabla (\rho h_s)} V_{ii} \right]  \\
+    &+& \frac{1}{\rho_{ii}} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right)_{ii} \left[ (\dot{m}_\alpha'''+\dot{m}_{b,\alpha}''')_{ii} V_{ii} -
+    \sum_{k=1}^{nf_c} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
+    - \overline{\mathbf{u} \cdot \nabla (\rho Y_\alpha)} V_{ii} \right] \nonumber \label{eq:divth2}
+\end{eqnarray}
+%
+where the over line terms refer to flux limited interpolation of corresponding scalars and terms defined with subscript $ii$ refer to cell defined quantities.
+All heat and mass fluxes and scalars are assumed known.
+Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid.
+Details of the FDS time integration scheme can be found in references~\cite{mcgratta_2013,mcdermo_2014}.
+For each RK2 substep, the species transport equations are advanced in all FDS cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region.
+A similar procedure is done for the thermodynamic divergence.
+% Small cell linking for explicit time integration.
+As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step.
+For these we link their unknown number to larger surrounding cells, which in turn defines mathematically an evolution equation for the union of both linked cells.
+Cell linking in the context of momentum equations can be found in reference~\cite{kirk_2003}.
+In FDS, the momentum equations and pressure Poisson equation are solved in the cartesian mesh.
+Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region.
+As will be shown next, a divergence integral equivalence argument is used to transfer the the divergence from cut-cells to under laying cartesian cells, and build the source of the Poisson equation for the pressure.
+
+\subsection{Immersed boundary method on cut-cell region and wall modeling}
+
+Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within equation~\ref{eqn:momtran}, the model momentum transport problem can be casted as~\cite{mcgratta_2013}:
+%
+\begin{eqnarray}
+  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \left[ \mathbf{F}(\mathbf{u},\mathbf{x},t) + \boldsymbol{\nabla} H(\mathbf{x},t) \right ]\label{eq:LowMachMom} \\
+         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{th} \label{eq:LowMachDiv}
+\end{eqnarray}
+%
+where equation~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by equation~\eqref{eq:divth}. 
+Boundary conditions are prescribed for $\mathbf{u}(\mathbf{x},t)$ on boundaries, including the immersed terrain.
+%Classical fractional step methods for time integration of incompressible or low Mach flow are based on two operations: First, momentum transport to obtain intermediate velocities, and second, projection of velocities into a target divergence space.
+The previous equations are advanced in time using a fractional step method.
+As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$.
+Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$ known
+%
+\begin{eqnarray}
+  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{eq:LowMachMomEu}\\
+  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} \label{eq:LowMachDivEu}
+\end{eqnarray}
+%
+where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$.
+This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator.
+As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible of enforcing the divergence condition and used on the projection step.
+Taking the divergence of equation~\eqref{eq:LowMachMomEu} and considering  the constraint~\eqref{eq:LowMachDivEu}, the two steps of the method are
+%
+\begin{enumerate}
+  \item Solve Poisson equation for $H^n$:
+
+\begin{equation}
+   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
+\end{equation}
+
+  \item Obtain final velocity for step:
+
+  \begin{equation}
+     \mathbf{u}^{n+1} = \mathbf{u}^{n} - \Delta t \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{it:FSProject}
+   \end{equation}
+
+   %The term $\hat{\mathbf{u}}^{n+1}=\mathbf{u}^{n} - \Delta t \mathbf{F}^n$ is known as intermediate velocity, and is a non divergence matching approximation to $\mathbf{u}^{n+1}$ (i.e.
+$\nabla \cdot \hat{\mathbf{u}}^{n+1} \neq \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$).
+\end{enumerate}
+%
+A consequence of the projection scheme is that boundary conditions are required on the Poisson equation of step~\eqref{it:FSPoisson}.
+For explicit methods and stationary \textit{solid} boundaries, the corresponding boundary condition is \textit{homogeneous} Neumann for $H^n$ in $\partial \Omega,\partial \Omega_1,...,\partial \Omega_{nbods}$~\cite{perot_1993}.
+Next, an approximation to the no-slip boundary condition at the immersed terrain surface is needed.
+To this end, a direct forcing immersed boundary method (IBM) for the momentum equations~\cite{fadlun_2000} is employed.
+A force field is computed on the discrete momentum equations on grid faces crossed by the immersed surfaces to approximate the no-slip boundary condition on these.
+In LES, the surrounding velocity field is modeled using an equilibrium boundary layer solution.
+
+In figure~\ref{Fig:figure_2}b the velocity update in each of the gas phase cut-face centroids $d$ is done individualizing the point $B$ on the boundary and normal direction through these.
+Also, and external point $e_x$ through the normal $\mathbf{\hat{n}}$ is defined at a distance $\delta_{ex}$, of the order of the cartesian cells size.
+Known velocities and fluid parameters are interpolated from the surrounding fluid points $e_1,\dots,e_4$ to $e_x$.
+This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution~\cite{mcgratta_2013}.
+The target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux matched to the under-laying cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$.
+Finally, an immersed boundary force
+%
+\begin{equation}
+F_E^k=-\frac{\hat{u}_E^{n+1,k-1}-u_E^{n}}{\Delta t} - \frac{\partial H^{n,k-1}}{\partial x}
+\label{eq:FIBM}
+\end{equation}
+%
+can be computed and used in equation~\eqref{it:FSPoisson} to take into account the presence of the body.
+The index $k$ refers in this context to the subiteration that can be performed in the IB force~\eqref{eq:FIBM} and projection~\eqref{it:FSPoisson}-\eqref{it:FSProject} to match final velocities with the wall modeled targets.
 
@@
 
+% Coupling level set fire spread model with fluid solver.
+
 % Defining atmospheric conditions:
@@
 The shorter this time scale, the faster the flow field will ``catch up'' to the mean wind, but at the expense of possibly washing out important flow structures.
-There is no firm guidance in the atmospheric modeling literature.
 

### 2020-06-10 Vanella_Article_Atmosphere_2020.tex
Minor changes

@@
 % Fire spread model equations.
-As commented in the introduction, the location of a fire front can be defined implicitly along the terrain surface through the zero of a two dimensional level set scalar function $\phi(x,y,t)$ in the form $\phi(x,y,t)=0$.
+As commented in the introduction, the location of a fire front can be defined implicitly along the terrain surface through the zero of a two dimensional level set scalar function $\phi(x,y,t)$.
 This implicit form can be used to define and advection equation for the level set function:
@@
 Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region.
-As will be shown next, a divergence integral equivalence argument is used to transfer the the divergence from cut-cells to under laying cartesian cells, and build the source of the Poisson equation for the pressure.
+A divergence integral equivalence argument is used to transfer the the divergence from cut-cells to under laying cartesian cells, and build the source of the Poisson equation for the pressure.
 

### 2020-06-10 Vanella_Article_Atmosphere_2020.tex
edits to section 4

@@
 
+\subsection{Particle Method}
+
+Lagrangian particles can be used to represent different types of vegetation, like leaves, grass, and so on.
+In practice each type of vegetation is represented by a single particle in each computational grid cell.
+Most often the particle is assumed to be cylindrical in shape.
+The particle is assigned a diameter based on the measured surface area to volume ratio, $\sigma$.
+The length is relatively unimportant so long as it is assumed to be much greater than the diameter.
+Material properties are assigned for the wet vegetation which is assumed to dry, decompose to char, and then exothermally oxidize when exposed to heat from an oncoming fire.
+The drag force exerted by the collection of particles is given by:
+\begin{equation}
+   \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \mathbf{u} \, ||\mathbf{u}||
+\end{equation}
+where $\rho$ is the air density, $C_{\rm d}$ is the drag coefficient, $C_{\rm s}$ is the shape factor (ratio of projected area to surface area), $\beta$ is the packing ratio, $\sigma$ is the surface area to volume ratio, and $\mathbf{u}$ is the air velocity.
+The diameter is specified via the surface area to volume ratio, $\sigma=2/r$ for a cylinder.
+The packing ratio is the volume of solid needles divided by the volume they occupy, typically denoted $\beta$.
+It is calculated by dividing the dry mass of vegetation per unit volume, the so-called ``bulk density,'' by the density of the dry vegetation, $\rho_{\rm d}$.
+
+A coniferous tree (e.g. pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.
+
 % Here, methodology to get fuels in the terrain from online sources.

### 2020-06-10 Vanella_Article_Atmosphere_2020.tex
Work on Section 4

@@
 
-There are three ways of simulating wildland fire spread.
-
+There are various ways of simulating wildland fire spread depending on the resolution of the underlying numerical grid. 
 \begin{enumerate}
-\item {\bf Particle Method:} The vegetation is represented by a collection of Lagrangian particles that are heated via convection and radiation.
+\item {\bf Particle Model:} The vegetation is represented by a collection of Lagrangian particles that are heated via convection and radiation.
+This model is appropriate for grid resolution of the order of 1~m or less.
 \item {\bf Boundary Fuel Model:} Ground vegetation is modeled like a porous solid with a thickness equal to the height of the vegetation.
-\item {\bf Level Set Method:} The fire front propagates using purely empirical rules.
+This model is appropriate for grid resolution of the order of 1~m to 10~m.
+\item {\bf Level Set Method:} The fire front propagates using purely empirical rules. This model is appropriate for grid resolution of the order of roughly 10~m and greater. 
 \end{enumerate}
@@
 
-\subsection{Particle Method}
+\subsection{Particle Model}
 
 Lagrangian particles can be used to represent different types of vegetation, like leaves, grass, and so on.
-In practice each type of vegetation is represented by a single particle in each computational grid cell.
+In the model each type of vegetation is represented by a single particle in each computational grid cell.
 Most often the particle is assumed to be cylindrical in shape.
@@
    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \mathbf{u} \, ||\mathbf{u}||
+   \label{drag}
 \end{equation}
@@
 A coniferous tree (e.g. pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.
+
+\subsection{Boundary Fuel Model}
+
+In many simulations of wildland fire, the ground vegetation layer is too shallow to be resolved explicitly, as is done when using Lagrangian particles to represent the vegetation.
+In such cases, the ground vegetation can be modeled as a porous boundary consisting of a layer of dry vegetation, moisture, and air, underneath of which is hard ground~\cite{Perez-Ramirez:FT2017}.
+The drag exerted by the vegetation is modeled using a special velocity boundary condition, and convective heat transfer is modeled via a source term in the one-dimensional heat conduction equation that is solved through the layer of vegetation and solid ground.
+Thermal radiation penetrates the vegetation layer via a 1-D radiative transport equation that is used for semi-transparent solids.
+
+The Boundary Fuel and Particle Models share the same basic input parameters and pyrolyis model.
+The drag exerted on the wind flowing through the vegetation is imposed as a force term in the gas phase grid cell adjacent to the boundary:
+\begin{equation}
+    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \frac{h_{\rm b}}{\delta z} \, \mathbf{u} \, ||\mathbf{u}||
+\end{equation}
+where the parameters are the same as in Eq.~(\ref{drag}) except for the additional depth of vegetation, $h_{\rm b}$, height of a grid cell, $\delta z$, and $\mathbf{u}$ is the gas velocity in the first grid cell.
+
+Thermal radiation is absorbed in depth according to a 1-D radiative transport solver. The absorption coefficient is given by:
+\begin{equation}
+   \kappa = C_{\rm s} \, \sigma \, \beta
+\end{equation}
+Thermal convection is not imposed at the interface between the gas phase and the vegetation layer, but rather is imposed via a source term in the 1-D heat conduction solver:
+\begin{equation}
+   \langle \dot{q}_{\rm c,b}''' \rangle = \sigma \, \beta \, \dot{q}_{\rm c}'' \quad ; \quad \dot{q}_{\rm c}'' = h \, (T_{\rm g} - T_{\rm s}) \quad \hbox{W/m}^2 \quad ; \quad  h = \frac{k}{L} \, {\rm Nu}
+\end{equation}
+where $\dot{q}_{\rm c}''$ the convective heat transferred to a small, cylindrical solid under the assumption that the subgrid-scale vegetation is cylindrical and the gas velocity and temperature are taken from the first gas phase grid cell adjacent to the boundary.
+
+\subsection{Level Set Model}
+
+For simulations of wildland fires spanning large areas that cannot be gridded finely enough to predict fire spread using a physics-based model, there is an empirical model built based on level sets~\cite{Bova:IJWF2015}.
+The methodology reproduces in an Eulerian framework the approach used in the Lagrangian-based fire front-tracking model FARSITE~\cite{Finney:FARSITE}.
+This approach makes use of the Rothermel-Albini~\cite{Rothermel:1972,Albini:1976} surface fire spread rate formulae and the assumption that a surface fire spreading from a point under certain wind, slope and vegetation conditions does so with an ellipse-shaped fire front with, for a given wind speed, a fixed length-to-breadth ratio~\cite{Richards:1990}.
+
+The level set spread model can be applied in various ways:
+\begin{enumerate}
+\item Only the level set simulation is performed. The wind is not affected by the terrain, and there is no fire.
+\item The wind field is established over the terrain, but it is ``frozen'' when the fire ignites.
+\item The wind field follows the terrain, but there is no actual fire in the simulation. It is just front-tracking.
+\item The wind and fire are fully-coupled. When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
+\end{enumerate}
+When fully coupled to the CFD model, the level set function acts as an igniter as it reaches a new surface cell.
+The cell burns for a pre-determined amount of time.
+For cases where the depth of the fireline is less than the size of a grid cell, the burning time is extended while the burning rate is decreased so as to maintain conservation of fuel mass.
 

### 2020-10-27 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 % Motivation.
-In the last decades, wild land and wild land-urban interface (WUI) fires have received substantial attention due to their destructive nature and extensive cost in lives and property~\cite{thomas_2017,mcdermott_2019,boghani_2019,richards_2020}.
+In the last decades, wildland and wildland-urban interface (WUI) fires have received substantial attention due to their destructive nature and extensive cost in lives and property~\cite{thomas_2017,mcdermott_2019,boghani_2019,richards_2020}.
 Over the years, modeling and simulation activities of such phenomena have been tailored to the needs of planners, scientists and first responders, and by the available computing capabilities.
@@
 Fuel models have been cataloged~\cite{Anderson:1982}, and their input parameters defined in a manner the user can select and combine the most appropriate listed fuels for her case of study.
-Rothermels fire line model is a widely used wildfire spread model, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013} \textbf{OTHERS?}, enabling a range of planning and operational forecasting capabilities.
+Rothermel's fire line model is a widely used wildfire spread model, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013,Coen:2,Coen:2015,CAWFE,ELMFIRE,WRF-Fire,Coen:2013,Mandel:2009,WRF-SFIRE,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
 
@@
 
+In July and August of 1986, the Commonwealth Scientific and Industrial Research Organisation (CSIRO) of Australia conducted controlled grassland fire experiments near Darwin, Northern Territory~\cite{Cheney:IJWF1993}.
+July and August are in the middle of the dry season when the grasses are fully cured (dried) and the weather is warm and dry.
+The experiments were conducted on flat plots measuring 100~m by 100~m, 200~m by 200~m, or 200~m by 300~m.
+Two cases have been simulated.
+Case~C064 was conducted on a 100~m by 100~m plot of kerosene grass ({\it Eriachne burkittii}); Case~F19 was conducted on a 200~m by 200~m plot of kangaroo grass ({\it Themeda australis}).
+
+Two of these experiments were originally simulated with FDS by Mell~et~al.~\cite{Mell:IJWF2007}.
+These simulations modeled the grass as a collection of cylindrical Lagrangian particles.
+Now these two experiments are also simulated using the Boundary Fuel Model (BFM)~\cite{Perez-Ramirez:FT2017} and the Rothermel-Albini fire spread algorithm~\cite{Rothermel:1972,Albini:1976}.
+For the experiment labelled Case~C064, fuel index 1 (Short Grass) is used, with a modified moisture fraction of 0.063.
+For F19, fuel index 3 (Tall Grass) is used, with a modified moisture fraction of 0.058.
+
+Measured properties for the specific types of grasses burned in the two experiments are listed in Table~\ref{Properties_Grasses}.
+Properties that were not measured are listed in Table~\ref{Assumed_Properties_Grasses}.
+These assumed properties are typically for wood or cellulosic fuels.
+The moisture is modeled as water.
+The grass is assumed to be composed primarily of cellulose.
+
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{figures/Case_F19_0358.png}
+\caption[Snapshot of the simulation of CSIRO Grassland Fire F19]{Snapshot of the simulation of CSIRO Grassland Fire F19.}
+\label{F19}
+\end{figure}
+
+A snapshot of the Lagrangian particle simulation of Case~F19 is shown in Fig.~\ref{F19}.
+The computational domain in this case is 240~m by 240~m by 20~m.
+The grid cells are 0.5~m cubes.
+The domain is subdivided into 36 individual meshes and run in parallel.
+The grass is represented 1 simulated blade per grid cell.
+The radius of the cylinder is derived from the measured surface area to volume ratio.
+Each simulated blade of grass represents many more actual blades of grass.
+The weighting factor is determined from the measured bulk mass per unit area.
+The fires in the experiments were ignited by two men carrying drip torches walking in opposite directions along the upwind boundary of the plot (the red strip in Fig.~\ref{F19}).
+In FDS, this action was modeled using a specified spread rate along the strip.
+
+\begin{table}[ht]
+\begin{center}
+\caption[Measured properties for the CSIRO Grassland Fire cases]{Measured properties for the CSIRO Grassland Fire cases~\cite{Cheney:IJWF1993}.}
+\label{Properties_Grasses}
+\begin{tabular}{|l|c|c|c|}
+\hline
+Property                        & Units         & Case F19      & Case C064     \\ \hline \hline
+Wind Speed                      & m/s           & 4.8           & 4.6           \\ \hline
+Ambient Temperature             & $^\circ$C     & 34            & 32            \\ \hline
+Surface Area to Volume Ratio    & m$^{-1}$      & 12240         & 9770          \\ \hline
+Grass Height                    & m             & 0.51          & 0.21          \\ \hline
+Bulk Mass per Unit Area         & kg/m$^2$      & 0.313         & 0.283         \\ \hline
+Moisture Fraction               & \%            & 5.8           & 6.3           \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\caption[Assumed properties for dry grass and soil]{Assumed properties for various types of dried grass and soil.
+Note that the Pyrolysis Temperature is taken to be the temperature at which the mass loss rate peaks in the TGA experiments of Morvan and Dupuy~\cite{Morvan:CF2004}.}
+\label{Assumed_Properties_Grasses}
+\begin{tabular}{|l|c|c|c|}
+\hline
+Property                        & Units                 & Value                     & Reference                             \\ \hline \hline
+Chemical Composition            & --                    & C$_6$H$_{10}$O$_5$        & Assumption                            \\ \hline
+Heat of Combustion              & kJ/kg                 & 15600                     & \cite{Susott:FS1982}                  \\ \hline
+Soot Yield                      & kg/kg                 & 0.015                     & \cite{SFPE:Tewarson}                  \\ \hline
+Char Yield                      & kg/kg                 & 0.2                       & \cite{Susott:FS1982}                  \\ \hline
+Specific Heat                   & kJ/(kg$\cdot$K)       & 1.5                       & Various sources                       \\ \hline
+Conductivity                    & W/(m$\cdot$K)         & 0.1                       & Assumption                            \\ \hline
+Density                         & kg/m$^3$              & 512                       & \cite{Rothermel:1972}                 \\ \hline
+Heat of Pyrolysis               & kJ/kg                 & 418                       & \cite{Morvan:CF2004}                  \\ \hline
+Pyrolyis Temperature            & $^\circ$C             & 200                       & \cite{Morvan:CF2004}                  \\ \hline \hline
+Obukhov Length                  & m                     & -500                      & Assumption                            \\ \hline
+Aerodynamic Roughness Length    & m                     & 0.03                      & Assumption                            \\ \hline
+Drag Coefficient                & --                    & 2.8                       & \cite{Falkenstein-Smith:2018}         \\ \hline \hline
+Soil Specific Heat              & kJ/(kg$\cdot$K)       & 2.0                       & \cite{Farouki:1981}                   \\ \hline
+Soil Conductivity               & W/(m$\cdot$K)         & 0.25                      & \cite{Farouki:1981}                   \\ \hline
+Soil Density                    & kg/m$^3$              & 1300                      & \cite{Farouki:1981}                   \\ \hline
+
+\end{tabular}
+\end{center}
+\end{table}
+
 \subsection{Cogoleto Fire}  \label{sec:cogo}
@@
 \externalbibliography{yes}
-\bibliography{Vanella_Article_Atmosphere_2020}
+\bibliography{Vanella_Article_Atmosphere_2020,LOFM_Workshop_Refernces}
 

### 2020-10-27 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 
-Among these, the fire spread model by Rothermel~\cite{Rothermel:1972} assumes quasi-steady state ground fire conditions and takes as inputs fuel properties, wind, terrain slope and moisture that can be measured in situ.
+Among these, the fire spread rate model by Rothermel~\cite{Rothermel:1972} assumes quasi-steady state surface fire conditions and takes as inputs fuel properties, wind, terrain slope and moisture that can be measured in situ.
 Fuel models have been cataloged~\cite{Anderson:1982}, and their input parameters defined in a manner the user can select and combine the most appropriate listed fuels for her case of study.
-Rothermel's fire line model is a widely used wildfire spread model, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013,Coen:2,Coen:2015,CAWFE,ELMFIRE,WRF-Fire,Coen:2013,Mandel:2009,WRF-SFIRE,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
+Rothermel's head fire rate of spread formula is a widely used, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013,Coen:2,Coen:2015,LAUTENBERGER_2013,Coen:2013,Mandel:2009,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
 
@@
 Evolution of wild land fires is usually tied to a fire front, which can have a complex shape moving/deforming with the local flame spread velocities.
-Knowing the location of this fire front and local properties of the fire, fuel and terrain allows, by use of simplified models like Rothermels, to predict local ignition and fire spread.
+Knowing the location of this fire front and local properties of the fire, fuel and terrain allows, by use of simplified models like FARSITE~\cite{Finney:FARSITE}, to predict local ignition and fire spread.
 When modeling the problem numerically, identifying such fire front at a given moment in time requires a mathematical representation and an associated discretization or mesh.
@@
 
-For fire modeling on the gas phase, FDS employs a Low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}.
-Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\overline{p}(z,t)$ used on the equation of state (the ideal gas law is used), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.
+For fire modeling on the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}.
+Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used on the equation of state (the ideal gas law is used), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.
 
 Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$.
-If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, were $\rho_\alpha(\mathbf{x},t)$ is the species density for the local thermodynamic conditions.
+If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, where $\rho_\alpha(\mathbf{x},t)$ is the species mass density and $\rho = \sum \rho_\alpha$.
 The scalar transport and momentum equations take the form:
 \begin{eqnarray}
-   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \mathbf{J_{c}}_\alpha ) &=& - \nabla \cdot (\mathbf{J_{d}}_\alpha) + \dot{m}_\alpha'''  +
+   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \rho Y_\alpha  \mathbf{u} ) &=& - \nabla \cdot \mathbf{J_{d}}_\alpha + \dot{m}_\alpha'''  +
     \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
@@
 \end{eqnarray}
-where $\mathbf{J_{c}}_\alpha=\rho Y_\alpha  \mathbf{u}$, $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ are the convective and diffusive fluxes for component $\alpha$.
-Ficks Law for binary diffusion respect to a background species is assumed, $D_\alpha$ is the diffusivity of $\alpha$ respect to the background species, $\dot{m}_\alpha'''(\mathbf{x},t)$ is the combustion volumetric source for species $\alpha$ and $ \dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from sub-grid particles gasification.
-Also, in the momentum equation the head $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, and $\tilde{p}$ is the perturbation pressure, $\rho_0(z)$ is an equlibrium density, $\mathbf{g}=(0,0,g_z)$ gravity vector, $\mathbf{f}_{b}$ is the particle force term and $\boldsymbol{\tau}$ tensor accounting for molecular and sub-grid turbulent stresses parameterized by an eddy viscosity model.
-This last model is employed by default in large eddy simulation (LES) of outdoor fires.
-Details of the model equations and their derivation are found in reference~\cite{mcgratta_2013tr}.
+where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ are the diffusive fluxes for component $\alpha$.
+Fick's Law for binary diffusion respect to a background species is assumed, $D_\alpha$ is the diffusivity of $\alpha$ respect to the background species.
+To enforce realizability, $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$, we lump errors in diffusive transport into the most abundant species locally \cite{McDermott:2015}.
+The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification.
+In the momentum equation the head is $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an equlibrium densit.
+The gravity vector is $\mathbf{g}=(0,0,g_z)$, $\mathbf{f}_{b}$ is the particle force term, and $\boldsymbol{\tau}$ is a tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model.
+This last model is employed by default in large-eddy simulation (LES) of outdoor fires.
+Details of the model equations and their derivation are found in \cite{mcgratta_2013tr}.
 
-Further, the flow field is subjected to the following divergence constraint:
+Conservation of energy is achieved by forcing flow field to obey the following thermodynamic divergence constraint:
 %
@@
 This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}.
-Being $R$ the ideal gas constant, problem is closed with the ideal gas law $ \overline{p} = {\rho R T}/{\overline{W}}$.
+Equation (\ref{eq:divth}) is derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law.
+Given a background pressure and a local mass density the local temperature is computed from the ideal gas law as $T = \bar{p} \overline{W} / (\rho R) $.
 
@@
 Moisture Fraction               & \%            & 5.8           & 6.3           \\ \hline
+Measured ROS                   & m/s           & 1.0           & 1.0           \\ \hline
+Calculated ROS, Particle Method & m/s           & 1.0           & 1.0           \\ \hline
+Calculated ROS, Boundary Fuel Method & m/s           & 1.0           & 1.0           \\ \hline
+Calculated ROS, Level Set Method & m/s           & 1.0           & 1.0           \\ \hline
 \end{tabular}
@@
 
-\subsection{Cogoleto Fire}  \label{sec:cogo}
+\subsection{Cogoleto Fire}  
+\label{sec:cogo}
+
+At approximately 23:00 (local time) on March 25, 2019, a wildfire started in the seaside town of Cogoleto, near Geno, Italy.
+Strong winds, gusting up to 100~km/h, spread the fire from its ignition point on a ridge above the town towards the sea.
+
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{figures/cogoleto_fire_2019_ls4_1000.png}
+\caption{Satellite photograph of the region where the Cogoleto Fire occurred. The area outlined in red is the extent of the actual fire; green is the simulation.}
+\label{Cogoleto_satellite}
+\end{figure}
+
+Shown in Fig.~\ref{Cogoleto_satellite} are the results of a level set simulation of the fire superimposed on a satellite photograph of the area and the outline of the extent of the actual fire.
+The computational domain is assembled using the open-source GIS (Geographic Information System) program called QGIS~\cite{QGIS}.
+The wind speed and direction are based on a single weather station near the point of ignition.
+The vegetation type is provided by ??.
+In the simulation, the single point wind data is taken as the prevailing wind, and the local wind field is obtained via a CFD computation.
+The fire is simulated based on the position of the level set front; that is, when the front arrives at a given location, a fire is ignited and burns for a duration of time consistent with the specified fuel loading of that particular point on the map.
+
+The extent of the simulated fire, shown in green in Fig.~\ref{Cogoleto_satellite}, is greater than the actual fire, shown in red.
+This is not surprising given that the simulation does not include fire suppression and the rate of spread of the level set front is determined empirically for vegetation that may not exactly resemble that of this particular region.
 
@@
 \externalbibliography{yes}
-\bibliography{Vanella_Article_Atmosphere_2020,LOFM_Workshop_Refernces}
+\bibliography{Vanella_Article_Atmosphere_2020,LOFM_Workshop_References}
 

### 2020-10-30 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 % Authors, for the paper (add full first names)
-\Author{Marcos Vanella $^1$, Kevin McGrattan $^1$, Randall McDermott $^1$, Glenn Forney $^1$, and William Mell $^2$}
+\Author{Marcos Vanella $^1$, Kevin McGrattan $^1$, Randall McDermott $^1$, Glenn Forney $^1$, William Mell $^2$, Emanuele Gissi $^3$, and Paolo Fiorucci $^4$}
 
 % Authors, for metadata in PDF
-\AuthorNames{Marcos Vanella, Kevin McGrattan, Randall McDermott, Glenn Forney, and William Mell}
+\AuthorNames{Marcos Vanella, Kevin McGrattan, Randall McDermott, Glenn Forney, William Mell, Emanuele Gissi, and Paolo Fiorucci}
 
@@
 $^{1}$ \quad National Institute of Standards and Technology, Gaithersburg, Maryland, USA \\
-$^{2}$ \quad U.S. Forest Service, Seattle, Washington, USA}
+$^{2}$ \quad U.S. Forest Service, Seattle, Washington, USA \\
+$^{3}$ \quad Corpo nazionale dei Vigili del fuoco, Savona, Italy \\
+$^{4}$ \quad CIMA Research Foundation, Savona, Italy}
 
@@
 Further, atmospheric wind conditions are specified using a mean forcing concept.
-Fire spread is modeled using one of three basic approaches: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set method in which the fire spreads as a function of terrain slope, vegetation type, and wind speed.
+Additionally, three basic approaches have been explored to model fire spread: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set method in which the fire spreads as a function of terrain slope, vegetation type, and wind speed.
 Several test and validation cases are reported to demonstrate the capabilities of this novel wildfire simulation methodology.}
@@
 
-% Motivation.
-In the last decades, wildland and wildland-urban interface (WUI) fires have received substantial attention due to their destructive nature and extensive cost in lives and property~\cite{thomas_2017,mcdermott_2019,boghani_2019,richards_2020}.
-Over the years, modeling and simulation activities of such phenomena have been tailored to the needs of planners, scientists and first responders, and by the available computing capabilities.
-% Wildfire modeling in the literature.
-A large body of literature can be found on the subject.
-See for example reviews~\cite{Papadopoulos_2011,Bakhshaii_2019,mcdermott_2019} on classical and new models with associated software.
-Due to inherent complex nature of the problem, most models of wildfires have historically relied on simplified field-tested rules and correlations.
+In the past few decades, wildland and wildland-urban interface (WUI) fires have received substantial attention due to their destructive nature and extensive cost in lives and property~\cite{thomas_2017,mcdermott_2019,richards_2020}.
+A large body of literature can be found on the subject, as seen in various review articles~\cite{Papadopoulos_2011,Bakhshaii_2019,mcdermott_2019}.
+Due to the complex nature of fire, limited computing resources, and the needs of planners and first responders, most models of wildfires have historically relied on simplified field-tested rules and correlations.
+Among these, the rate of spread model of Rothermel~\cite{Rothermel:1972} assumes quasi-steady state surface fire conditions and takes as inputs fuel properties, wind, terrain slope, and moisture that can be measured {\em in situ}.
+Fuel models have been cataloged~\cite{Anderson:1982}, and their parameters defined so that model users can select and combine the most appropriate listed fuels.
+Rothermel's rate of spread formula is still widely used, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013,Coen:2,Coen:2015,LAUTENBERGER_2013,Coen:2013,Mandel:2009,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
 
-Among these, the fire spread rate model by Rothermel~\cite{Rothermel:1972} assumes quasi-steady state surface fire conditions and takes as inputs fuel properties, wind, terrain slope and moisture that can be measured in situ.
-Fuel models have been cataloged~\cite{Anderson:1982}, and their input parameters defined in a manner the user can select and combine the most appropriate listed fuels for her case of study.
-Rothermel's head fire rate of spread formula is a widely used, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013,Coen:2,Coen:2015,LAUTENBERGER_2013,Coen:2013,Mandel:2009,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
+Wildland fire modeling typically focuses on tracking the fire front, which can have a complex shape that moves and deforms based on local conditions like fuel loading, moisture content, wind, and terrain.
+The fire front requires a mathematical representation.
+The Lagrangian approach is based on a polygonal mesh of control markers that represents the front~\cite{Finney:FARSITE,Bova:IJWF2015}.
+On the other hand, an Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field.
+This scalar field is called the level set function and numerical methods that solve for it in time are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
+Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}.
+A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.
 
-% Level set as a simplified fire spread model, operational forecasting.
-Evolution of wild land fires is usually tied to a fire front, which can have a complex shape moving/deforming with the local flame spread velocities.
-Knowing the location of this fire front and local properties of the fire, fuel and terrain allows, by use of simplified models like FARSITE~\cite{Finney:FARSITE}, to predict local ignition and fire spread.
-When modeling the problem numerically, identifying such fire front at a given moment in time requires a mathematical representation and an associated discretization or mesh.
-The Lagrangian treatment of the motion of a fire front requires specifying a set of control markers and connectivity that define and move with such front.
-Therefore, the fire front is represented by this discrete mesh.
-For example, if the domain of motion is a surface terrain, this Lagrangian mesh can be a polygon that moves with the front along this surface~\cite{Finney:FARSITE,Bova:IJWF2015}.
-On the other hand, an Eulerian representation of motion allows for the mesh to be fixed in space (i.e.
-in a terrain a polygonal patch mesh of said terrain).
-Then, the location of the fire front can be portrayed implicitly by using a scalar field defined in the fixed two dimensional mesh.
-This scalar field is called level set function and numerical methods that evolve it in time are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
-Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}.
-A practical comparison of the two techniques as implemented in WFDS~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.
+Because the spread rate of the fire front depends in part on the local wind conditions, a well resolved wind field over complex terrain should improve the accuracy of the overall model.
+The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{coen_2013,Mell:IJWF2007}.
+For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind.
+For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model.
+The wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution.
+These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept.
+This is a first step at implementing a level set approach in a LES wind calculation with combustion.
+It is recognized that, in the empirically-based development of the Rothermel formula, the wind speed was not influenced by the local buoyancy induced flow.
 
-% Resolving the flow features on terrains using LES over complex terrains. Describe what we do, FDS, cut-cell schemes, immersed boundary methods.
-As the rate of spread of the front line is dependent in the local velocity on empirical fire spread models like Rothermels, a better local resolution of the velocity field around complex terrains should be beneficial to the accuracy of the overall model.
-In reality, besides the obstructive character of the terrain itself, compounded effects due to the fire affect the local velocity field.
-\textbf{Some of these are ...}.
-The de facto physics-based technique for practical simulation of outdoor flows is Large Eddy Simulation (LES)~\cite{coen_2013,mcgratta_2013}.
-Within FDS~\cite{mcgratta_2013},  the filtered Navier-Stokes equations for buoyant, combusting flows are evolved using a Low Mach approximation.
-Chemical species are handled using scalar transport equations.
-Energy transport is accounted for by means of the solution of radiative transport equations and a thermodynamic divergence proxy.
-\textbf{Comment on WFDS for LOF.
-The work of Mell et al.~\cite{Mell:IJWF2007}.}.
+In the following section the mathematical model for fire spread over complex terrain is described.
+The description of the terrain discretization is provided in section~\ref{sec:terraindisc}.
+The various fire spread methods are described in section~\ref{sec:firespread}, and the atmospheric boundary conditions in section~\ref{sec:wind}.
+Some flat terrain simulations are compared to experimental data, and a complex terrain simulation is compared to an actual wildfire in section~\ref{sec:numexp}.
 
-% Unstructured geometries, capability can be tuned for use with LOF and fire spread models.
-% What has been done, cut-cell scheme IBM, connection to LS and Rothermels, wind.
-Lately, substantial work has been done to add to FDS the option of simulating flows around unstructured geometries.
-These are represented by surface triangulations which is a typical representation in computer aided design (CAD) and simulation software.
-This added flexibility is still compliant with the structured cartesian grids and differentiation schemes within FDS.
-We implemented a cut-cell scheme~\cite{berger_2016} to advance the scalar transport and energy of multicomponent mixtures on the polyhedra next to the geometries.
-These polyhedrons arise from geometry intersection with the cartesian fluid grid.
-As geometry boundaries are not changed, the method has the property of exact volume and area conservation.
-Velocities next to geometries are reconstructed using an immersed boundary method (IBM)~\cite{fadlun_2000,balaras_2004} coupled with an equilibrium wall model for turbulent flows~\cite{mcdermo_2018}.
-Variables related to IBM forcing and heat transfer modeling at the wall are sampled through interpolation and a normal probe approach~\cite{balaras_2004}.
-Also, radiation transport boundary conditions have been adapted to this technique.
-For fire simulation over complex terrains a range of options in terms of physics fidelity of the model are being developed, in particular, in the representation of fire spread.
-For fast simulation of outdoor fires, this new capability has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and Rothermels fire spread model.
-The wind speed input for Rothermels model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution.
-These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept.
-
-% Outline of the manuscript.
-In the following section a description on the mathematical model for fire over terrain is provided.
-The terrain description and model discretization is provided in section~\ref{sec:terraindisc}.
-An account of the fire spread model is given in section~\ref{sec:firespread}, as well as the wind atmospheric boundary conditions in section~\ref{sec:wind}.
-Numerical experiments testing the methodology on a simple example and a real case of fire in Cogoleto, Italy are provided in section~\ref{sec:numexp}.
-Finally a summary of this work is outlined in section~\ref{sec:summary}.
-
-\section{Mathematical model} \label{sec:matmodel}
+\section{Mathematical Model} \label{sec:matmodel}
 
@@
 The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification.
-In the momentum equation the head is $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an equlibrium densit.
+In the momentum equation the head is $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an equlibrium density.
 The gravity vector is $\mathbf{g}=(0,0,g_z)$, $\mathbf{f}_{b}$ is the particle force term, and $\boldsymbol{\tau}$ is a tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model.
@@
 Conservation of energy is achieved by forcing flow field to obey the following thermodynamic divergence constraint:
-%
 \begin{eqnarray}
@@
 \end{eqnarray}
-%
 where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux, sum of conduction, convection and radiation, and $T$ is the local temperature.
@@
 
-% Fire spread model equations.
-As commented in the introduction, the location of a fire front can be defined implicitly along the terrain surface through the zero of a two dimensional level set scalar function $\phi(x,y,t)$.
-This implicit form can be used to define and advection equation for the level set function:
-%
-\begin{equation}
-   \frac{D \phi(x,y,t)}{D t} =  \frac{\partial \phi}{\partial t} + \mathbf{R}_{uv} \cdot \nabla (\phi) = 0 \label{eqn:lset}
-\end{equation}
-%
-where $\mathbf{R}_{uv}=(R_u,R_v)$ is the rate of spread vector along the terrain plane.
-The magnitude of the rate of spread vector is computed by $|\mathbf{R}_{uv}|=R_0 \left(1 + \sqrt{(\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) \cdot (\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) } \right)$, being $R_0$ the zero velocity and zero slope rate of spread (baseline, dependent of fuel properties), and vectors $\boldsymbol{\phi}_W,\boldsymbol{\phi}_S \ge 0$ are computed using local wind, slope through empirical rules~\cite{Wilson:1980}.
-See Appendix 1 of reference~\cite{Bova:IJWF2015} for details on these calculations, as implemented in WFDS-LS and in this work.
-
 \section{Terrain description and discretization} \label{sec:terraindisc}
 
-% How do we describe terrains - geometries. Methodology to get from DEM a geometry.
-%
 In figure~\ref{Fig:figure_1}a, a terrain is represented by its surface triangulation within an FDS cartesian mesh.
@@
 They discretize the fluid domain in the region surrounding the solid terrain.
-A detail of this unstructured mesh is.
-shown in figure~\ref{Fig:figure_1}b.
+A detail of this unstructured mesh is shown in figure~\ref{Fig:figure_1}b.
 The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where more than one cut-cell per cartesian cell are not uncommon.
@@
 Reliably defining the cut-cell mesh geometric properties and topology is a complex problem on itself, omitted here in order to keep the discussion succinct.
-%
-\begin{figure}[h]
+\begin{figure}[ht]
    \centering
@@
 \end{figure}
-%
 
@@
 Integrating equation~\eqref{eqn:spectran} over a cut-cell $ii$ with control volume $\Omega_{ii}$:
-%
 \begin{equation}
@@
 \end{equation}
-%
 For a control volume that does not change with time, the time derivative and source terms are approximated by
-%
 \begin{eqnarray}
@@
 \end{eqnarray}
-%
 where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages.
@@
 Consider the FV discretization of the diffusive term (eqn.~\eqref{eq:intconvdiff}) on cut-cell $ii$ of figure~\ref{Fig:figure_2}a:
-%
 \begin{equation}
@@
 \end{equation}
-%
 The integral over the cut-cell volume has been transformed in an area integral on its $nf_c=5$ faces using the divergence theorem.
@@
 Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces.
-%
-\begin{figure}[h]
+\begin{figure}[ht]
    \centering
@@
 \end{figure}
-%
-
 Similarly, the discretization of the advective term is:
-%
 \begin{equation}
@@
 \end{equation}
-%
 where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux limited interpolation to the cut-face~\cite{mcgratta_2013}.
@@
 The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:
-%
-%
 \begin{eqnarray}
@@
 \end{eqnarray}
-%
 where the over line terms refer to flux limited interpolation of corresponding scalars and terms defined with subscript $ii$ refer to cell defined quantities.
@@
 The index $k$ refers in this context to the subiteration that can be performed in the IB force~\eqref{eq:FIBM} and projection~\eqref{it:FSPoisson}-\eqref{it:FSProject} to match final velocities with the wall modeled targets.
+This velocity reconstruction procedure has been combined with the level set method for fire spread described in next section.
 
@@
 The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels and the fire spread rate is {\em predicted} by the model.
-The Level Set Method relies on a set of experimentally-determined spread rates for different types of vegetation and wind speeds.
+The Level Set Method relies on a set of predetermined spread rates for different types of vegetation and wind speeds.
 
@@
 Lagrangian particles can be used to represent different types of vegetation, like leaves, grass, and so on.
-In the model each type of vegetation is represented by a single particle in each computational grid cell.
+In the model, each type of vegetation is represented by a single particle in each computational grid cell.
 Most often the particle is assumed to be cylindrical in shape.
@@
 \begin{equation}
-   \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \mathbf{u} \, ||\mathbf{u}||
+   \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \mathbf{u} \, \|\mathbf{u}\|
    \label{drag}
@@
 
-For simulations of wildland fires spanning large areas that cannot be gridded finely enough to predict fire spread using a physics-based model, there is an empirical model built based on level sets~\cite{Bova:IJWF2015}.
-The methodology reproduces in an Eulerian framework the approach used in the Lagrangian-based fire front-tracking model FARSITE~\cite{Finney:FARSITE}.
+For simulations of wildland fires spanning large areas that cannot be gridded finely enough to predict fire spread using a physics-based model, fire spread via level sets is an option~\cite{Bova:IJWF2015}.
 This approach makes use of the Rothermel-Albini~\cite{Rothermel:1972,Albini:1976} surface fire spread rate formulae and the assumption that a surface fire spreading from a point under certain wind, slope and vegetation conditions does so with an ellipse-shaped fire front with, for a given wind speed, a fixed length-to-breadth ratio~\cite{Richards:1990}.
-
-The level set spread model can be applied in various ways:
+The location of a fire front is defined implicitly via a two dimensional level set scalar function $\phi(x,y,t)=0$.
+The advection equation for the level set function is
+\begin{equation}
+   \frac{D \phi(x,y,t)}{D t} =  \frac{\partial \phi}{\partial t} + \mathbf{R}_{uv} \cdot \nabla (\phi) = 0 \label{eqn:lset}
+\end{equation}
+where $\mathbf{R}_{uv}=(R_u,R_v)$ is the spread rate vector. The magnitude of the vector is
+\begin{equation}
+  \|\mathbf{R}_{uv}\|=R_0 \left(1 + \sqrt{(\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) \cdot (\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) } \right)
+\end{equation}
+where $R_0$ is the zero velocity, zero slope rate of spread that depends on the fuel properties.
+The vectors $\boldsymbol{\phi}_W,\boldsymbol{\phi}_S \ge 0$ are computed using the local wind and slope through empirical rules~\cite{Wilson:1980}.
+See Appendix 1 of reference~\cite{Bova:IJWF2015} for details on these calculations.
+In FDS, the level set spread model can be applied in various ways:
 \begin{enumerate}
@@
 
-% Here, methodology to get fuels in the terrain from online sources.
-
-% Coupling level set fire spread model with fluid solver.
-
-% Defining atmospheric conditions:
 \section{Defining atmospheric wind : The Mean Forcing Concept} \label{sec:wind}
@@
 
-\subsection{Simple example}  \label{sec:simexp}
+\subsection{Flat Terrain Fire Spread}  \label{sec:simexp}
 
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{figures/Case_F19_0358.png}
-\caption[Snapshot of the simulation of CSIRO Grassland Fire F19]{Snapshot of the simulation of CSIRO Grassland Fire F19.}
+\includegraphics[width=\textwidth]{figures/F19_collage.png}
+\caption{Photographs of the experiment and snapshots of the simulation of CSIRO Grassland Fire F19, 56~s, 86~s, and 138~s following ignition.}
 \label{F19}
@@
 
-A snapshot of the Lagrangian particle simulation of Case~F19 is shown in Fig.~\ref{F19}.
-The computational domain in this case is 240~m by 240~m by 20~m.
+Snapshots of the Lagrangian particle simulation of Case~F19 compared to photographs of the experiment are shown in Fig.~\ref{F19}.
+The plot of grass is 200~m by 200~m and the computational domain is 240~m by 240~m by 20~m high.
 The grid cells are 0.5~m cubes.
 The domain is subdivided into 36 individual meshes and run in parallel.
-The grass is represented 1 simulated blade per grid cell.
-The radius of the cylinder is derived from the measured surface area to volume ratio.
+A blade of grass is represented by a single cylindrically-shaped particle within a grid cell.
+The radius of the cylinder is derived from the measured surface area to volume ratio of the grass.
 Each simulated blade of grass represents many more actual blades of grass.
@@
 Moisture Fraction               & \%            & 5.8           & 6.3           \\ \hline
-Measured ROS                   & m/s           & 1.0           & 1.0           \\ \hline
-Calculated ROS, Particle Method & m/s           & 1.0           & 1.0           \\ \hline
-Calculated ROS, Boundary Fuel Method & m/s           & 1.0           & 1.0           \\ \hline
-Calculated ROS, Level Set Method & m/s           & 1.0           & 1.0           \\ \hline
+Measured ROS                    & m/s           & 1.5           & 1.1           \\ \hline
+Calculated ROS, Particle Method & m/s           & 1.5           & 1.0           \\ \hline
+Calculated ROS, Boundary Fuel Method & m/s      & 3.7           & 2.5           \\ \hline
+Calculated ROS, Level Set Method & m/s          & 1.3           & 0.7           \\ \hline
 \end{tabular}
@@
 
-\begin{table}
+The predicted rates of spread for the two cases are listed in Table~\ref{Properties_Grasses}.
+For these simulations, the particle method is the most accurate, and the boundary fuel method over-predicts by about a factor of two.
+The level set method under-predicts the rate of spread, but only because of the somewhat arbitrary choice of fuel type.
+For a simple flat terrain scenario like this, the choice of fuel model to accompany the level set calculation largely determines the spread rate.
+The original FDS simulations of this experiment ~\cite{Mell:IJWF2007} also used a boundary fuel approach, with better results, work is underway to reconcile the differences.
+It is not surprising that the particle method does well in this scenario because it is closest to the actual physics.
+However, for larger domains, the boundary fuel and level set methods are better able to approximate the fire spread while the particle method would require an excessive amount of computer resources.
+
+\begin{table}[ht]
 \begin{center}
@@
 
-\subsection{Cogoleto Fire}  
+\subsection{Complex Terrain Fire Spread}  
 \label{sec:cogo}
 
-At approximately 23:00 (local time) on March 25, 2019, a wildfire started in the seaside town of Cogoleto, near Geno, Italy.
-Strong winds, gusting up to 100~km/h, spread the fire from its ignition point on a ridge above the town towards the sea.
+At approximately 23:00 (local time) on March 25, 2019, a wildfire started in the seaside town of Cogoleto, near Genova, Italy.
+Strong winds, gusting up to 100~km/h from the North, spread the fire from its ignition point due to a faulty electric line on a ridge above the town towards the sea.
+Several houses were destroyed, hundreds of residents had to be evacuated, the firefighter crews worked for three days to tame the flames, no casualties were reported.
 
@@
 Shown in Fig.~\ref{Cogoleto_satellite} are the results of a level set simulation of the fire superimposed on a satellite photograph of the area and the outline of the extent of the actual fire.
-The computational domain is assembled using the open-source GIS (Geographic Information System) program called QGIS~\cite{QGIS}.
+
+The computational domain is assembled using the open-source GIS (Geographic Information System) program called QGIS~\cite{QGIS}, extended by the qgis2fds~\cite{qgis2fds} plugin.
+The open-source qgis2fds plugin was specifically developed in the framework of the WUIFI-21 project\footnote{WUIFI-21: {\it High fidelity computational fluid dynamics modeling of forest fires for Wildland-Urban Interface communities resilience and protection}, an Italy-US research project financed by the Italian Ministry of Foreign Affairs and International Cooperation} for facilitating the use of geographical data for forest fire simulation and smoke pollutants dispersion in FDS.
+
 The wind speed and direction are based on a single weather station near the point of ignition.
-The vegetation type is provided by ??.
+The vegetation type is provided by a database maintained at the CIMA\footnote{International Centre on Environmental Monitoring.} Research Foundation, derived from regional forest management maps.
 In the simulation, the single point wind data is taken as the prevailing wind, and the local wind field is obtained via a CFD computation.
@@
 
+The simulation results were compared with those obtained from CIMA Propagator~\cite{Trucchia:2020}, an experimental propagation model that has provided several European civil protection organizations with real-time fire predictions since 2009.
+The model has been implemented by the CIMA Foundation, its propagation model is based on stochastic cellular automata.
+
 \section{Summary} \label{sec:summary}
+
+The numerical methods described in this paper span the range of length scales from tens of centimeters to tens of kilometers.
+Obviously, these techniques support differing descriptions of the underlying fire physics, but all can be embedded within the same code base.
+This allows researchers to combine the different techniques in ways that were not possible when separate codes were maintained by separate organizations.
+With FDS, one can simulate 24 hours of fire spread in minutes of computation time using the level set function and a wind field that varies only temporally, not spatially, with no fire-atmospheric coupling.
+At the same time, using the same set of input parameters, one can simulate the fire at far greater resolution with much greater physical fidelity, albeit for shorter time periods and spatial extent.
 

### 2020-11-01 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 
+\newcommand{\dV}{{\rm d}V}
+\newcommand{\dS}{{\rm d}S}
+
 \begin{document}
@@
 On the other hand, an Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field.
-This scalar field is called the level set function and numerical methods that solve for it in time are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
+This scalar field is called the level set function and numerical methods that compute the function are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
 Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}.
@@
 Because the spread rate of the fire front depends in part on the local wind conditions, a well resolved wind field over complex terrain should improve the accuracy of the overall model.
-The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{coen_2013,Mell:IJWF2007}.
+The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{Linn:2007,coen_2013,Mell:IJWF2007}.
 For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind.
@@
 These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept.
-This is a first step at implementing a level set approach in a LES wind calculation with combustion.
-It is recognized that, in the empirically-based development of the Rothermel formula, the wind speed was not influenced by the local buoyancy induced flow.
+This is a first step at implementing a level set approach in an LES wind calculation with combustion.
+It is recognized that in the empirically-based development of the Rothermel formula the wind speed is not influenced by the local buoyancy induced flow.
 
@@
 
-For fire modeling on the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}.
-Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used on the equation of state (the ideal gas law is used), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.
+For fire modeling in the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}.
+Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used in the equation of state (ideal gas law), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.
 
@@
 \end{eqnarray}
-where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ are the diffusive fluxes for component $\alpha$.
-Fick's Law for binary diffusion respect to a background species is assumed, $D_\alpha$ is the diffusivity of $\alpha$ respect to the background species.
-To enforce realizability, $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$, we lump errors in diffusive transport into the most abundant species locally \cite{McDermott:2015}.
+where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ is the diffusive flux for component $\alpha$.
+Fick's Law for binary diffusion with respect to a background species is assumed, and $D_\alpha$ is the diffusivity of $\alpha$ with respect to the background species.
+To enforce realizability, $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}.
 The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification.
-In the momentum equation the head is $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an equlibrium density.
+In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an ambient density of the atmosphere.
 The gravity vector is $\mathbf{g}=(0,0,g_z)$, $\mathbf{f}_{b}$ is the particle force term, and $\boldsymbol{\tau}$ is a tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model.
-This last model is employed by default in large-eddy simulation (LES) of outdoor fires.
-Details of the model equations and their derivation are found in \cite{mcgratta_2013tr}.
+This last model is employed by default in LES of outdoor fires.
+Details of the model equations and their derivation are found in Ref.~\cite{mcgratta_2013tr}.
 
-Conservation of energy is achieved by forcing flow field to obey the following thermodynamic divergence constraint:
+Conservation of energy is achieved by forcing the flow field to obey the following thermodynamic divergence constraint:
 \begin{eqnarray}
-    ( \nabla \cdot \mathbf{u} )^{th} &=&
+    ( \nabla \cdot \mathbf{u} )^{\rm th} &=&
     \left[ \frac{1}{\rho c_p T} - \frac{1}{\bar{p}} \right]
@@
 \end{eqnarray}
-where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux, sum of conduction, convection and radiation, and $T$ is the local temperature.
-The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively, built from the individual species counterparts as defined.
-Finally, $w,\rho_0,g_z$ on the stratification term are respectively, the local vertical velocity, density for standard conditions (depends on vertical coordinate), and vertical gravity acceleration.
+where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux vector (conduction, convection and radiation), and $T$ is the local gas temperature.
+The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively.
 This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}.
-Equation (\ref{eq:divth}) is derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law.
-Given a background pressure and a local mass density the local temperature is computed from the ideal gas law as $T = \bar{p} \overline{W} / (\rho R) $.
+Equation~(\ref{eq:divth}) is derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law.
+Given a background pressure and a local mass density, the local temperature is computed from the ideal gas law as $T = \bar{p} \overline{W} / (\rho R) $.
 
@@
 
-In figure~\ref{Fig:figure_1}a, a terrain is represented by its surface triangulation within an FDS cartesian mesh.
-Some cartesian cells and faces belonging to this mesh are transversed by the terrain surface.
-The remainder polyhedra and polygons that lay on the gas side of this intersected cartesian entities are called cut-cells and cut-faces.
-They discretize the fluid domain in the region surrounding the solid terrain.
-A detail of this unstructured mesh is shown in figure~\ref{Fig:figure_1}b.
-The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where more than one cut-cell per cartesian cell are not uncommon.
-As described later, small cells pose a temporal stability constraint in explicit time integration schemes that needs to be addressed.
-It is noted that a two level of refinement grid hierarchy emerges.
-The coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured counterparts.
+In Fig.~\ref{Fig:figure_1}, the terrain is represented by its surface triangulation within an FDS Cartesian mesh.
+Some Cartesian cells and faces belonging to this mesh are transversed by the terrain surface.
+The remaining polyhedra and polygons that lie on the gas side of these intersected geometries are called cut-cells and cut-faces.
+They discretize the fluid domain in the region surrounding the solid surface.
+A detail of this unstructured mesh is shown in Fig.~\ref{Fig:figure_1}.
+The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where there are more than one cut-cell per Cartesian cell are common.
+As described later, small cells impose a temporal stability constraint on explicit time integration schemes.
+A two-level grid refinement hierarchy emerges.
+The coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured components.
 Methods that solve discrete model equations on these grids are called cut-cell or embedded boundary methods~\cite{berger_2016}.
-Reliably defining the cut-cell mesh geometric properties and topology is a complex problem on itself, omitted here in order to keep the discussion succinct.
+Reliably defining the cut-cell mesh geometric properties and topology is a complex problem in and of itself, but beyond the scope of this paper.
 \begin{figure}[ht]
@@
 
-The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}.
-See figure~\ref{Fig:figure_2}a.
-Integrating equation~\eqref{eqn:spectran} over a cut-cell $ii$ with control volume $\Omega_{ii}$:
+The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}, as outlined in the diagram of Fig.~\ref{Fig:figure_2}.
+Integrating Eq.~(\ref{eqn:spectran}) over a cut-cell $ii$ with control volume $\Omega_{ii}$:
 \begin{equation}
- \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
-      } d \Omega  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega + \int_{\Omega_{ii}} {( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''') } d \Omega \label{eq:intconvdiff}
+ \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
+      } \, \dV  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \mathbf{J_{d}}_\alpha   } \, \dV + \int_{\Omega_{ii}} { \left( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''' \right) } \, \dV \label{eq:intconvdiff}
 \end{equation}
 For a control volume that does not change with time, the time derivative and source terms are approximated by
-\begin{eqnarray}
-  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega & = & \frac{\partial}{\partial t} \int_{\Omega_{ii}} {\rho Y_\alpha} d \Omega
-  = \frac{\partial \widetilde{\rho \: Y_\alpha }_{ii}}{\partial t} V_{ii} \\
-  \int_{\Omega_{ii}} {( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''')} d \Omega & = & (\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii}) V_{ii} \label{eq:intcons}
-\end{eqnarray}
+\begin{equation}
+  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \widetilde{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
+  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons}
+\end{equation}
 where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages.
 In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged.
-Consider the FV discretization of the diffusive term (eqn.~\eqref{eq:intconvdiff}) on cut-cell $ii$ of figure~\ref{Fig:figure_2}a:
+Consider the FV discretization of the diffusive term of Eq.~(\ref{eq:intconvdiff}) on cut-cell $ii$ of Fig.~\ref{Fig:figure_2}:
 \begin{equation}
-\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega =
-    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega = \sum^{nf_c}_{k=1}
+\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot   \mathbf{J_{d}}_\alpha   } \, \dV =
+    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \, \dS = \sum^{n_{\rm f}}_{k=1}
     \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvdiffcc}
 \end{equation}
-The integral over the cut-cell volume has been transformed in an area integral on its $nf_c=5$ faces using the divergence theorem.
-As these $k$ faces with areas $A_k$ and normals out of the control volume $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization.
-A centroid to centroid (i.e.
-$\Delta x_{cc}$ in figure~\ref{Fig:figure_2}a) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase.
+The integral over the cut-cell volume has been transformed in an area integral on its $n_{\rm f}=5$ faces using the divergence theorem.
+As these $k$ faces with areas $A_k$ and outward normals $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization.
+Centroid to centroid (i.e.
+$\Delta x_{\rm cc}$ in Fig.~\ref{Fig:figure_2}) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase.
 Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces.
@@
 \begin{equation}
-    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} d \Omega =
-     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega =
-     \sum^{nf_c}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvadvcc}
+    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} \, \dV =
+     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \, \dS \approx
+     \sum^{n_{\rm f}}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \, A_k \label{eq:discfvadvcc}
 \end{equation}
-where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux limited interpolation to the cut-face~\cite{mcgratta_2013}.
+where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{mcgratta_2013}.
 In the cut-cell region we use Godunov interpolation for the advective term.
-As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of equations~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known.
+As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of Eqs.~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known.
 The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:
 \begin{eqnarray}
-    ( \nabla \cdot \mathbf{u} )_{ii}^{th} \; V_{ii} &=&
+    ( \nabla \cdot \mathbf{u} )_{ii}^{\rm th} \; V_{ii} &=&
     \left[ \frac{1}{(\rho c_p T)_{ii}} - \frac{1}{\bar{p}_{ii}} \right]
@@
     &+& \frac{1}{(\rho c_p T)_{ii}} \left[ (\dot{q}'''+ \dot{q}_b''')_{ii} V_{ii} -
-    \sum_{k=1}^{nf_c} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
+    \sum_{k=1}^{n_{\rm f}} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
     - \overline{\mathbf{u} \cdot \nabla (\rho h_s)} V_{ii} \right]  \\
     &+& \frac{1}{\rho_{ii}} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right)_{ii} \left[ (\dot{m}_\alpha'''+\dot{m}_{b,\alpha}''')_{ii} V_{ii} -
-    \sum_{k=1}^{nf_c} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
+    \sum_{k=1}^{n_{\rm f}} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
     - \overline{\mathbf{u} \cdot \nabla (\rho Y_\alpha)} V_{ii} \right] \nonumber \label{eq:divth2}
 \end{eqnarray}
-where the over line terms refer to flux limited interpolation of corresponding scalars and terms defined with subscript $ii$ refer to cell defined quantities.
+where the over line terms refer to flux limited interpolation of corresponding scalars, and terms defined with subscript $ii$ refer to cell defined quantities.
 All heat and mass fluxes and scalars are assumed known.
@@
 Details of the FDS time integration scheme can be found in references~\cite{mcgratta_2013,mcdermo_2014}.
-For each RK2 substep, the species transport equations are advanced in all FDS cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region.
+For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region.
 A similar procedure is done for the thermodynamic divergence.
@@
 As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step.
-For these we link their unknown number to larger surrounding cells, which in turn defines mathematically an evolution equation for the union of both linked cells.
-Cell linking in the context of momentum equations can be found in reference~\cite{kirk_2003}.
-In FDS, the momentum equations and pressure Poisson equation are solved in the cartesian mesh.
+These cut-cells are linked to larger surrounding cells.
+Cell linking in the context of momentum equations can be found in Ref.~\cite{kirk_2003}.
+In FDS, the momentum equations and pressure Poisson equation are solved on the Cartesian mesh.
 Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region.
-A divergence integral equivalence argument is used to transfer the the divergence from cut-cells to under laying cartesian cells, and build the source of the Poisson equation for the pressure.
+A divergence integral equivalence argument is used to transfer the divergence from cut-cells to the underlying Cartesian cells in order to build the source term of the Poisson equation for the pressure.
 
@@
 
-Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within equation~\ref{eqn:momtran}, the model momentum transport problem can be casted as~\cite{mcgratta_2013}:
-%
+Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within Eq.~(\ref{eqn:momtran}), the model momentum transport problem can be written~\cite{mcgratta_2013}:
 \begin{eqnarray}
-  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \left[ \mathbf{F}(\mathbf{u},\mathbf{x},t) + \boldsymbol{\nabla} H(\mathbf{x},t) \right ]\label{eq:LowMachMom} \\
-         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{th} \label{eq:LowMachDiv}
+  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \mathbf{F}(\mathbf{u},\mathbf{x},t) - \boldsymbol{\nabla} H(\mathbf{x},t)  \label{eq:LowMachMom} \\
+         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{\rm th} \label{eq:LowMachDiv}
 \end{eqnarray}
-%
-where equation~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by equation~\eqref{eq:divth}.
+where Eq.~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by Eq.~\eqref{eq:divth}.
 Boundary conditions are prescribed for $\mathbf{u}(\mathbf{x},t)$ on boundaries, including the immersed terrain.
-%Classical fractional step methods for time integration of incompressible or low Mach flow are based on two operations: First, momentum transport to obtain intermediate velocities, and second, projection of velocities into a target divergence space.
+
 The previous equations are advanced in time using a fractional step method.
 As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$.
-Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$ known
-%
+Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th}$ known
 \begin{eqnarray}
-  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{eq:LowMachMomEu}\\
-  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} \label{eq:LowMachDivEu}
+  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \mathbf{F}^n -  \boldsymbol{\nabla} H^n \label{eq:LowMachMomEu}\\
+  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} \label{eq:LowMachDivEu}
 \end{eqnarray}
-%
 where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$.
 This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator.
-As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible of enforcing the divergence condition and used on the projection step.
-Taking the divergence of equation~\eqref{eq:LowMachMomEu} and considering  the constraint~\eqref{eq:LowMachDivEu}, the two steps of the method are
-%
+As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible for enforcing the divergence condition and used on the projection step.
+Taking the divergence of Eq.~\eqref{eq:LowMachMomEu} and considering  the constraint, Eq.~\eqref{eq:LowMachDivEu}, the two steps of the method are
 \begin{enumerate}
@@
 \begin{equation}
-   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
+   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
 \end{equation}
@@
 
-   %The term $\hat{\mathbf{u}}^{n+1}=\mathbf{u}^{n} - \Delta t \mathbf{F}^n$ is known as intermediate velocity, and is a non divergence matching approximation to $\mathbf{u}^{n+1}$ (i.e.
-$\nabla \cdot \hat{\mathbf{u}}^{n+1} \neq \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$).
 \end{enumerate}
-%
-A consequence of the projection scheme is that boundary conditions are required on the Poisson equation of step~\eqref{it:FSPoisson}.
-For explicit methods and stationary \textit{solid} boundaries, the corresponding boundary condition is \textit{homogeneous} Neumann for $H^n$ in $\partial \Omega,\partial \Omega_1,...,\partial \Omega_{nbods}$~\cite{perot_1993}.
+A consequence of the projection scheme is that boundary conditions are required on the Poisson equation, \eqref{it:FSPoisson}.
+For explicit methods and stationary \textit{solid} boundaries, the corresponding boundary condition is \textit{homogeneous} Neumann for $H^n$ in $\partial \Omega,\partial \Omega_1,...,\partial \Omega_{\rm nbods}$~\cite{perot_1993}.
 Next, an approximation to the no-slip boundary condition at the immersed terrain surface is needed.
@@
 
-In figure~\ref{Fig:figure_2}b the velocity update in each of the gas phase cut-face centroids $d$ is done individualizing the point $B$ on the boundary and normal direction through these.
-Also, and external point $e_x$ through the normal $\mathbf{\hat{n}}$ is defined at a distance $\delta_{ex}$, of the order of the cartesian cells size.
+In Fig.~\ref{Fig:figure_2}, the velocity update in each of the gas phase cut-face centroids $d$ is done individualizing the point $B$ on the boundary and normal direction through these.
+Also, and external point $e_x$ through the normal $\mathbf{\hat{n}}$ is defined at a distance $\delta_{ex}$, of the order of the Cartesian cells size.
 Known velocities and fluid parameters are interpolated from the surrounding fluid points $e_1,\dots,e_4$ to $e_x$.
 This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution~\cite{mcgratta_2013}.
-The target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux matched to the under-laying cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$.
+The target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux-matched to the underlying Cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$.
 Finally, an immersed boundary force
-%
 \begin{equation}
@@
 \end{equation}
-%
-can be computed and used in equation~\eqref{it:FSPoisson} to take into account the presence of the body.
-The index $k$ refers in this context to the subiteration that can be performed in the IB force~\eqref{eq:FIBM} and projection~\eqref{it:FSPoisson}-\eqref{it:FSProject} to match final velocities with the wall modeled targets.
+can be computed and used in Eq.~\eqref{it:FSPoisson} to take into account the presence of the body.
+The index $k$ refers in this context to the sub-iteration that can be performed in the IB force, Eq.~\eqref{eq:FIBM}, and projection, Eqs.~\eqref{it:FSPoisson}-\eqref{it:FSProject}, to match final velocities with the wall modeled targets.
 This velocity reconstruction procedure has been combined with the level set method for fire spread described in next section.
@@
 
-A coniferous tree (e.g. pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.
+A coniferous tree (e.g., pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.
 
@@
 In many simulations of wildland fire, the ground vegetation layer is too shallow to be resolved explicitly, as is done when using Lagrangian particles to represent the vegetation.
-In such cases, the ground vegetation can be modeled as a porous boundary consisting of a layer of dry vegetation, moisture, and air, underneath of which is hard ground~\cite{Perez-Ramirez:FT2017}.
+In such cases, the ground vegetation can be modeled as a porous boundary consisting of a layer of dry vegetation, moisture, and air, underneath which is hard ground~\cite{Perez-Ramirez:FT2017}.
 The drag exerted by the vegetation is modeled using a special velocity boundary condition, and convective heat transfer is modeled via a source term in the one-dimensional heat conduction equation that is solved through the layer of vegetation and solid ground.
@@
 \begin{equation}
-    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \frac{h_{\rm b}}{\delta z} \, \mathbf{u} \, ||\mathbf{u}||
+    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \frac{h_{\rm b}}{\delta z} \, \mathbf{u} \, \|\mathbf{u}\|
 \end{equation}
@@
 This approach makes use of the Rothermel-Albini~\cite{Rothermel:1972,Albini:1976} surface fire spread rate formulae and the assumption that a surface fire spreading from a point under certain wind, slope and vegetation conditions does so with an ellipse-shaped fire front with, for a given wind speed, a fixed length-to-breadth ratio~\cite{Richards:1990}.
-The location of a fire front is defined implicitly via a two dimensional level set scalar function $\phi(x,y,t)=0$.
+The location of a fire front at time $t$ is defined by the set of points $x,y$ where the level set scalar function $\phi(x,y,t)=0$.
 The advection equation for the level set function is
@@
 \begin{enumerate}
-\item Only the level set simulation is performed. The wind is not affected by the terrain, and there is no fire.
+\item Only the level set simulation is performed, with a constant and uniform specified wind and slope. The wind is not affected by the terrain, and there is no fire.
 \item The wind field is established over the terrain, but it is ``frozen'' when the fire ignites.
-\item The wind field follows the terrain, but there is no actual fire in the simulation. It is just front-tracking.
+\item The wind field follows the terrain, but there is no actual fire in the simulation, just front-tracking.  The level set evolves continuously in time with the flow field.
 \item The wind and fire are fully-coupled. When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
@@
 The process of steering the solution of the mass, momentum, and energy equations to match wind speed and direction gathered at random weather stations is known as \emph{data assimilation}~\cite{Kalnay:2003}.
-FDS uses a relatively simple data assimilation technique, a method known as \emph{nudging}, where you add a forcing term to the momentum equation to ``nudge'' the flow in the direction of the specified wind.
+FDS uses a relatively simple data assimilation technique, a method known as \emph{nudging}, where a forcing term is added to the momentum equation to ``nudge'' the flow in the direction of the specified wind.
 FDS automatically drives the mean velocity components toward the desired values by adding a forcing term to the momentum equation.

### 2020-12-08 Response_to_A_Maranghides.tex
Update on Overleaf.

\documentclass[12pt]{article}

\setlength{\textwidth}{6.25in}
\setlength{\textheight}{8.75in}
\setlength{\topmargin}{0.in}
\setlength{\headheight}{0.in}
\setlength{\headsep}{0.in}
\setlength{\parindent}{0.in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\parskip}{\baselineskip}
\setlength{\leftmargini}{\parindent} % Controls the indenting of the "bullets" in a list

\usepackage{xcolor}
\newcommand\hl[1]{\textcolor{red}{#1}}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf Response to comments by WERB Reviewer A. Maranghides}
\end{center}

I focused my review on the later part of the paper, because of my technical expertise and also because of context/ impact.

I am specifically interested in the case of the Cogoleto Fire. I think the case presented here is \textbf{oversimplified and can potentially direct the reader to make certain assumptions about the model capabilities}. Even though this is a small part of the paper in many ways it is maybe the most important in terms of representing the model usefulness.

What is lacking here are:
%
\begin{itemize}

   \item Information about the scale of the event: how large was the fire from origin to furthest point, how long it took to reach the final perimeter. A scale on the figure 4 would be helpful.

   From origin to furthest point measured distance is ~ 1200 m.

   \item Wind context: was the single measurement consistent both in speed and direction throughout the event. How representative was this single point measurement, given the second closest weather station and how that data compared. was the station at a specific location (ravine or local acceparation zone). I can tell you that in CA, in the Camp Fire (as an example here) 5 -8 km can make all the difference in the world (100 km/h vs 10 km/h)

   \item information on defensive actions. Defensive actions not only impact the final perimeter but can also impact the direction as firefighters typically try to direct the fire away from high value targets. This can be done with both ground and aerial resources. I do not know what was used and when.

   \item There is no time scale provided on the runs or event progression so it is very difficult to judge what is presented.

   \item There is no mention of spotting within the fire perimeter and also outside. Short, intermediate and far field spotting can play critical roles in the evolution of a wildfire and/or WUI incident and in my mind is one of the key issues that realistically limits the applicability of predictive tools on that scale. Since we cannot resolve where embers will result in ignitions and the location of the actual ignitions can have dramatic impact on how the event develops, not knowing limits the usefulness of these tools. A probabilistic approach does not improve the usefulness in my mind.

   \item Explicit listing of the model limitations of dealing with fuel breaks both natural and manmade. While I appreciate that if the wind was truly 100km/h fuel breaks may have had little impact on the overall fire spread, the effect on fuel breaks can still provide critical information to understanding how fire behaved and should be discussed.

   \item I would also like to see some statement explicitly addressing the complexity of dealing with non-uniform fuels (not only vegetative but also WUI). Even using the level set approach one needs to be mindful of the potential change in fire spread as the fire migrates between different fuels.

\end{itemize}

My review recommendations are the following:

\begin{enumerate}
    \item[A:] Explicitly address the above mentioned issues.

    \item[B:] Use this opportunity to highlight the value of developing these large scale outdoor tools which are (in my mind):
    \item[B1.] Adding value to prescribed burns planning, where conditions are more benign (specifically wind) and data can be collected to refine and develop the models
    \item[B2.] Use the ability to model complex geometries at a multi-parcel level using physics to highly resolve fire spread and ultimately ignition.

\hl{For the previous two questions Kevin added a paragraph: "The strategy that has been adopted in FDS is to accommodate simulation options ranging from a level set calculation of fire spread over tens of kilometers that can be run in minutes, similar to models like BEHAVE developed by the U.S. Forest Service, all the way to detailed simulations of fire spread at sub-meter resolution for relatively short time periods (minutes to hours) and small areas (tens of hectares)." at the end of the Cogoleto section}

\end{enumerate}

Even if you are not interested/agree to go with B, at a minimum consider addressing A. While it may appear that I have been critical, I think the capabilities added to deal with complex geometries can add very much needed value to FDS at a different scale where we can test and reliably validate the code.

\end{document}

### 2020-12-08 Response_to_Reviewer_1.tex
Update on Overleaf.

\documentclass[12pt]{article}

\setlength{\textwidth}{6.25in}
\setlength{\textheight}{8.75in}
\setlength{\topmargin}{0.in}
\setlength{\headheight}{0.in}
\setlength{\headsep}{0.in}
\setlength{\parindent}{0.in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\parskip}{\baselineskip}
\setlength{\leftmargini}{\parindent} % Controls the indenting of the "bullets" in a list

\usepackage{xcolor}
\newcommand\hl[1]{\textcolor{red}{#1}}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf Response to comments by Reviewer 1}
\end{center}

The article "Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over
Complex Terrain" is clearly written and well structured. It shows noticeable effort to describe an
original immersed boundary method, which is the core of its purpose. However, it is difficult to
assess the relevance of this work, for three main reasons.

\hl{Our original intent was to present a broad overview of the our approach to modeling wildland fire. The technical details of the model, verification and validation, and a user's guide is part of the package that one downloads when using the model, and these documents are available at the model web site. Nevertheless, we have added more details to the paper to better describe our approach.}

1. A lack of model description

The IBM together with the main solved PDE are correctly presented and a large amount of
previous articles globally allows understanding the demarch. But many submodels and
parametrization choices, that may have a strong influence on the final result, are weakly
depicted. In particular:

The volume combustion source term is not described.

\hl{A new section (Sec.~2.2 ``Subgrid Parameterizations'') discusses subgrid parameterizations and combustion details.}

Radiative modelling in the thermal divergence equation is not given. In particular, indications
should be given on the gas and soot absorption coefficients, if required, or on a radiative fraction

\hl{Further details on the radiation model have been added.}

The pyrolysis model used for the lagrangian particles or the porous boundary is not given. In
particular, the pyrolysis model shape (volume model ? surface model ?) is not depicted, and
neither the chemical degradation path for the pyrolysing vegetation, nor the mass and heat
transfer models in the porous boundary or the solid particles, are explained, whereas all these
aspects may have a strong influence on the mass of fuel released in the atmosphere, and therefore
on the fire spread rate.

\hl{Details of the pyrolysis model have been added.}

The heat exchange modelling between the gas phase and the solid particles is not depicted.

\hl{Details have been added.}

Same problem for many semi-empirical coefficients used for fluid/solid interface models: the
drag coefficient and the shape factor in the drag force model (they probably differ between the
particle model and the boundary flux model since they do not consider the same solid objects),
the Nusselt correlation for the heat exchange coefficient.

\hl{We added more details about how the convective and radiative heat transfer is done for the particle and boundary fuel models.}

2. The influence of the numerical model parameters

A sensitivity analysis on (at least) the following parameters would be helpful to assess the
relevance of the numerical model:

grid cell size

\hl{The flat terrain fire spread (CSIRO) cases were all run at three different grid resolutions and showed no significant difference in spread rate. }

pyrolysis model parameters, more precisely those which are made under an a priori assumption
(conductivity, specific heat)

\hl{We added a reference to a paper that McGrattan wrote in 2017 for the U.S. Combustion Institute meeting on sensitivity of parameters in the particle-based CSIRO calculations}

Obukhov length and roughness length

In particular, a comparison with the parameters used in a previous study [22] would be
interesting, insofar as the model seems to be very close, except on the management on the
boundary condition.

\hl{Physical parameters describing the vegetation in the Australian grass land experiments are the same in the current study and the past study. However, many aspects of the modeling approach have changed or advanced, making it difficult to make useful comparisons regarding the influence of some model parameters or methods. These include the modeling approach for the ambient wind. This is discussed in the text.}

3. A lack of quantitative comparisons

In section 6.1, the only quantitative comparison is made on the average fire rate of spread. But,
similarly with Ref [22], many comparisons could be made:

Mass loss rate/heat release rate time histories

fire front positions

velocity in the low atmospheric layer

\hl{We added more details about the flat terrain fire spread simulations, in particular a two-dimensional comparison of the fire front versus aerial observations.}

All these comparisons between the two first approaches and the previous work could bring
important information to explain the observed discrepancies: in particular, if these differences
are explained by interface heat transfer issues, or pyrolysis model issues.

\hl{We understand the reviewer's point. The differences between the current model and the version of the model in Mell et al.(2007) could, potentially, shed light on physical processes that need model advancements, such as heat transfer or thermal degradation. However, there are numerous differences between the 2007 and the current version of the model and these prohibitively confound such a comparison.}

Finally, section 6.2 is of poor interest, because no reason for the differences between the actual
fire and the simulation is raised, and no comparison with other numerical tools is given.

\hl{Kevin}

To conclude with, this paper needs strong improvements in the model description, the
consideration of the effects of uncertainties related to model parametrization on the numerical
results, and in the quantitative description of the results and the differences between the
numerical model. In particular, it should include a precise comparison with simulations of ref.
[22], since the observed differences cannot be related, at this stage, to the implementation of the
IBM or to differences in the model description or parametrization. In this respect, the interest of
the IBM approach, which is the main originality of this paper, is not shown.

\end{document}

### 2020-12-08 Response_to_Reviewer_2.tex
Update on Overleaf.

\documentclass[12pt]{article}

\setlength{\textwidth}{6.25in}
\setlength{\textheight}{8.75in}
\setlength{\topmargin}{0.in}
\setlength{\headheight}{0.in}
\setlength{\headsep}{0.in}
\setlength{\parindent}{0.in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\parskip}{\baselineskip}
\setlength{\leftmargini}{\parindent} % Controls the indenting of the "bullets" in a list

\usepackage{xcolor}
\newcommand\hl[1]{\textcolor{red}{#1}}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf Response to comments by Reviewer 2}
\end{center}

Early attempts at a mechanistic modeled wildfire spread was in the late 1960s. After these few
years of active development and advancement a lagged occurred for 15 to 20 years. The early.
development was generally due to two engineers Rothermel and Albini. The delay was caused
by the fact that a mechanistic approach is not in the training background for most Foresters. In
the 1980s with the major advances in numerical techniques applied to combustion and fluid
dynamics the gap with the Forester’s got even larger. However, in the last few decades
engineers and fluid dynamicists have stepped in and made important progress. We are now at a
stage where mechanistic models are becoming common and designed for different needs of
wildfire managers.

This paper is one of these developments. It uses large eddy simulations in the Fire Dynamic
Simulator using three ways of describing the fuel and fire spread: fuel as a range of Lagrangian
particles, fuel as a porous media and the traditional (empirical) fire boundary from which the
fire spreads, the so-called level set approach. The arguments in the paper is well-developed
using a novel boundary technique—cut- cell immersed method. Considering the sophistication
of the rest of the model the assimilation technique used seems to be very simplistic and I
assume temporary. Wind is such an important process in wildfires one assumes that this will be
better developed. Overall, I found the model section difficult but very exciting.

\hl{Since the time of the initial manuscript submission a significant improvement has been made in the way we treat wind boundary conditions.  As the reviewer noted, the previous approach was a rudimentary form of data assimilation where nudging was was used for momentum throughout the computational domain.  On closer inspection of wind validation cases we discovered two significant issues with this approach: First, the forcing terms did not work well together with the ``open'' boundary conditions we were using at the time for inflow and outflow convective boundaries---unstable inflow waves would persist if any flow reversals due to turbulence were encountered.  Second, the forcing terms in the momentum equation were clearly suppressing turbulence on the leeward side of hills. \\ \\
To overcome these issues we have developed an improved ``open'' boundary condition, which is described in the revised manuscript (Sec.~5 on ``Atmospheric Wind Boundary Conditions'').  This new boundary condition solves both of the problems mentioned above, it is easier to implement, and it still allows for one-way downscaling of microscale meteorological model output or weather observations.  However, as the reviewer alludes, a more advanced modeling approach would allow for the possibility of a two-way coupling with the at least the microscale (1 km), but also the mesoscale (10 km) meteorologic models.  This is presently beyond our scope, but something we are considering for future work.}

The numerical experiments are important to show how the model works with real properties.
However, they seem to be an add-on and not very well developed and discussed. Something
needs to be done with this section to make it as important part of the rest of the paper.

\hl{More detail has been added to the numerical examples.}

The idea of putting the scale differences together in the model boundary seems to be a very
useful advance.

\hl{The new ``open'' boundary conditions are capable of utilizing a synthetic eddy method to superimpose turbulence length scales on the background mean flow profiles.  A small example case is provided as a proof of concept.  We plan to explore the affect of inflow turbulence in more depth as future work.}

\end{document}

### 2020-12-08 Response_to_Reviewer_3.tex
Update on Overleaf.

\documentclass[12pt]{article}

\usepackage{amsmath}

\setlength{\textwidth}{6.25in}
\setlength{\textheight}{8.75in}
\setlength{\topmargin}{0.in}
\setlength{\headheight}{0.in}
\setlength{\headsep}{0.in}
\setlength{\parindent}{0.in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\parskip}{\baselineskip}
\setlength{\leftmargini}{\parindent} % Controls the indenting of the "bullets" in a list

\usepackage{xcolor}
\newcommand\hl[1]{\textcolor{red}{#1}}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf Response to comments by Reviewer 3}
\end{center}

The study "Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over Complex Terrain" describe a method to consider the forcing induced by a complex topography in wildfire propagation within the framework of "physical models", combining different approach.

The work is undoubtedly well structured and seems the fruit of great number of years of serious work, so substantially I have nothing to point out. But from a formal point of view, in my opinion, it shows some aspect that should be improved. Here I suggest some observation to work upon.

1)	INTRODUCTION

Should be outlined where ``the state of the art'' and the bibliography analysis end, and where it starts the new work.

Given that some aims of this study are "fast simulation and outdoor fires" I would recommend to cite
the ``mass-consistent'' approach, ( for example the article Wagenbrenner, NS, Forthofer, JM, Lamb, BK, Shannon, KS, Butler, BW, (2016) Downscaling surface wind predictions from numerical weather prediction models in complex terrain with WindNinja. Atmos. Chem. Phys. 16:5229-5241, doi:10.5194/acp-16-5229-2016) and an "mass-consistent" application to a Rothermel/level set approach (Arca Bachisio, Ghisu Tiziano, Casula Marcello, Salis Michele, Duce Pierpaolo (2019) A web-based wildfire simulator for operational
applications. International Journal of Wildland Fire 28, 99-112. https://doi.org/10.1071/WF18078)

\hl{An additional paragraph has been added to the introduction that provides a reference to the review of models given in Arca et al. (2019) and provides some context on how the FDS suite of models differs from other approaches.}

2) MATHEMATICAL MODEL

The description of the mathematical model is too short. It is not possible to be clear in 30 lines for such a complex model. There is the risk of making a collection of citations and references to another series of works, while there should be an understandable basic outline.

Please provide at least a simple description of the phenomena that is going to be modelled, together with the model hypothesis that allow for the simplification of the general physical frame. For example, at least the kind of fuels and the transformation process involved in combustion should be briefly described. It would be important also to provide a short description of the single term of each equation and its derivation. For example, to understand the model and to have a ready legibility it is important to specify as (p-po)*g*l/p is the buoyancy term.

\hl{Defined explicitly the buoyancy term.}

The statement ``derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law'' is difficult to understand. Therefore a few words should be dedicated.

\hl{We include Equation (3) is derived by factoring the divergence85from the sensible enthalpy equation and applying the ideal gas law.}

63: ``realizability'' is an important topic that should be described a little.

\hl{We added after equations(1)-(2) : Mass fractions $Y_\alpha$, solution of equations (1), must obey realizability constraints $0<Y_\alpha<1$ and $\sum_\alpha Y_\alpha=1$. In equations (1) starting from a realizable solution, realizability implies $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$.}

64: Please describe the phenomena ``volumetric combustion'' and ``subgrid particle gassification''

\hl{We added : The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification. These correspond to the amount of mass per unit volume and time of species $\alpha$ being added or substracted in a given point $\mathbf{x}$ due to chemical reaction or gasification of solid particles (or droplets), respectively.}

66: ``an ambient density'' is too vague. Please specify. 
\hl{We added : $\rho_0(z)$ is the modeled height dependent background density of the atmosphere.}

67: Please supply a few words about the physical phenomena ``fb is the particle force term.''

\hl{We changed the statement to : $\mathbf{f}_{b}$ is the term contributed by modeled particle drag forces.}

Eq. (3): $w$ is not defined. Please supply the definition.

\hl{We added : Also, $\boldsymbol{\omega}$ refers to the vorticity field.}

76: What is the meaning of ``thermodynamic divergence'' in the superscript of the RHS of the energy conservation equation? Is the thermodynamic divergence different from a normal divergence?

\hl{A velocity divergence, $\nabla\cdot\mathbf{u}$, can be factored out of every transport equation.  Of course, in order for the fomulation to be consistent, each divergence must be the same.  In a typical low-speed flow algorithm for solving the Navier-Stokes equations, a target divergence is prescribed from thermodyanmics.  Often, for simplicity, the flow is considered ``divergence free'' if constant density is a reasonable approximation.  In that case, the target ``thermodynamic divergence'' is implied to be zero. The algorithm then uses a \emph{projection scheme} to enforce this constraint (this is the purpose of solving the Poisson equation for the pressure). With variable density combustion codes, the divergence is more complicated.  The target thermodynamic divergence is computed from the sensible enthalpy equation, utilizing mass conservation and the ideal gas equation of state. Again, the projection scheme enforces this constraint.  After projection, all the transport equations share the same divergence, as must be true for consistency.  Further, the sensible enthalpy equation has been satisfied, which accounts for the important effects of heat transfer and dilatation from the combustion.}

3) TERRAIN DESCRIPTION AN DISCRETIZATION

Other important concept are only cited but not described at all, please some word should be spend to describe :

2 lines before eq. (7) ``a normal probe approach''

\hl{We added : Also, a normal probe approach is employed to sample information from the fluid to define fluxes in boundary cut-faces. In a normal probe method, an external point is defined at a normal distance from the body of the order of the Cartesian grid size. Then, information from the fluid at this external point is obtained via interpolation. }

line 96: ``flux-limited interpolation'' 
line 96: ``Godunov interpolation''

\hl{We reorganized the paragraph : In the cut-cell region a Godunov flux limited interpolation for the advective term is used. Flux limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic systems, of which mass transport inherits its mathematical properties.}

line 97: ``Runge-Kutta method''

\hl{We added : Explicit Runge-Kutta time integration methods are single-step multistage schemes widely used for advancing ordinary and partial differential equations [Leveque2007].}

\hl{Marcos,Randy}

\end{document}

### 2020-12-08 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 
+\usepackage[export]{adjustbox}
+
 %=================================================================
@@
 Momentum transport and heat transfer close to the immersed terrain are handled using dynamic wall models and a direct forcing immersed boundary method.
-Further, atmospheric wind conditions are specified using a mean forcing concept.
+A new ``open'' convective inflow/outflow method for specifying atmospheric wind boundary conditions is presented.
 Additionally, three basic approaches have been explored to model fire spread: (1) representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set method in which the fire spreads as a function of terrain slope, vegetation type, and wind speed.
@@
 \newcommand{\dS}{{\rm d}S}
+\newcommand{\bx}{\mathbf{x}}
+\newcommand{\bs}{\mathbf{s}}
+\renewcommand{\C}{\mathrm{C}}
+\renewcommand{\H}{\mathrm{H}}
+\renewcommand{\O}{\mathrm{O}}
 
@@
 Fuel models have been cataloged~\cite{Anderson:1982}, and their parameters defined so that model users can select and combine the most appropriate listed fuels.
-Rothermel's rate of spread formula is still widely used, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,mcgratta_2013,Coen:2,Coen:2015,LAUTENBERGER_2013,Coen:2013,Mandel:2009,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
+Rothermel's rate of spread formula is still widely used, and has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,FDS_Users_Guide,Coen:2,Coen:2015,LAUTENBERGER_2013,Coen:2013,Mandel:2009,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting capabilities.
 
@@
 This scalar field is called the level set function and numerical methods that compute the function are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
-Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}.
+Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,FDS_Users_Guide,LAUTENBERGER_2013}.
 A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.
@@
 
+A number of other approaches to modeling fire spread over complex terrain exist.
+A recent review, with an emphasis on operational applications, is given in~\cite{Arca_2019}.
+A range of approximations to modeling both the wind and fire have been employed.
+The uniqueness of the approach presented here is that, within a single computational tool, a range of approximations to the wind and fire physics can be employed.
+This can support a direct comparison between models of varying physical fidelity.
+For example, one can investigate the difference in predicted fire behavior between a simulation that models fire spread and heat release via the explicit accounting of the thermal degradation of vegetation and one that approximates fire spread via an empirically based model.
+
 In the following section the mathematical model for fire spread over complex terrain is described.
@@
 \section{Mathematical Model} \label{sec:matmodel}
+
+\subsection{Governing Equations} \label{sec:goveqns}
 
@@
     \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
-    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau} \right] \label{eqn:momtran}
+    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau}^{dev} \right] \label{eqn:momtran}
 \end{eqnarray}
@@
 Fick's Law for binary diffusion with respect to a background species is assumed, and $D_\alpha$ is the diffusivity of $\alpha$ with respect to the background species.
-To enforce realizability, $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}.
+Mass fractions $Y_\alpha$, solution of equations~\eqref{eqn:spectran}, must obey realizability constraints $0<Y_\alpha<1$ and $\sum_\alpha Y_\alpha=1$.
+In equations~\eqref{eqn:spectran} starting from a realizable solution, realizability implies $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$.
+Therefore, to enforce realizability, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}.
 The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification.
-In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an ambient density of the atmosphere.
-The gravity vector is $\mathbf{g}=(0,0,g_z)$, $\mathbf{f}_{b}$ is the particle force term, and $\boldsymbol{\tau}$ is a tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model.
-This last model is employed by default in LES of outdoor fires.
-Details of the model equations and their derivation are found in Ref.~\cite{mcgratta_2013tr}.
+These correspond to the amount of mass per unit volume and time of species $\alpha$ being added or substracted in a given point $\mathbf{x}$ due to chemical reaction or gasification of solid particles (or droplets), respectively.
+In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is the modeled height dependent background density of the atmosphere.
+Also, $\boldsymbol{\omega}$ refers to the vorticity field.
+The gravity vector is $\mathbf{g}=(0,0,g_z)$, the buoyancy force term is $ (\rho-\rho_0) \mathbf{g}$, $\mathbf{f}_{b}$ is the term contributed by modeled particle drag forces, and $\boldsymbol{\tau}^{dev}$ is the deviatoric stress tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model, discussed below in Sec.~\ref{sec:subgrid}.
 
@@
     \frac{\partial \bar{p}}{\partial t} + \frac{w \rho_0 g_z}{\rho c_p T} \nonumber \\
-    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
+    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \nabla \cdot \dot{\mathbf{q}}_{\rm r}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
     &+& \frac{1}{\rho} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right) \left[ \dot{m}_\alpha'''  + \dot{m}_{b,\alpha}''' - \nabla \cdot (\mathbf{J_{d}}_\alpha) - \mathbf{u} \cdot \nabla (\rho Y_\alpha) \right] \label{eq:divth}
 \end{eqnarray}
-where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux vector (conduction, convection and radiation), and $T$ is the local gas temperature.
+where $h_{s,\alpha}$ is the sensible enthalpy of species $\alpha$ and $h_s$ is the sensible enthalpy of the mixture, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $T$ is the local gas temperature, and $w$ is the local vertical velocity.
 The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively.
@@
 
-\section{Terrain description and discretization} \label{sec:terraindisc}
+The term $\dot{\mathbf{q}}''$ in Eq.~(\ref{eq:divth}) represents the conductive and diffusive heat fluxes discussed in Sec~\ref{sec:subgrid}.
+The net contribution from thermal radiation in the energy equation is defined by:
+\begin{equation}
+   -\nabla \cdot \dot{\mathbf{q}}_{\rm r}''(\bx) =
+    \kappa(\bx) \, \left[ U(\bx) - 4 \pi \, I_{\rm b}(\bx) \right]  \quad ; \quad
+    U(\bx) = \int_{4\pi} \, I(\bx,\bs') \, d\bs'  \label{simple_rte}
+\end{equation}
+where $\kappa(\bx)$ is the absorption coefficient, $I_b(\bx)$ is the source term, and $I(\bx,\bs)$ is the solution of the radiation transport equation (RTE) for a non-scattering gray gas:
+\begin{equation}
+   \bs \cdot \nabla I(\bx,\bs) = \kappa(\bx) \; \left[ I_{\rm b}(\bx) - I(\bx,\bs) \right] \label{bandRTE1}
+\end{equation}
+The source term, $I_{\rm b}$, requires special treatment because of the limited resolution of the underlying numerical grid in the vicinity of flames.
+In large scale fire simulations, grid cells are typically on the order of
+tens of centimeters.
+Flame sheets cannot be resolved, meaning that the computed cell-average temperature can be significantly lower than temperatures one would expect to find in the reacting flame.
+Consequently, the
+source term is approximated in grid cells where fuel and oxygen react. Elsewhere, the subgrid temperature field is homogeneous and the source term can be computed directly:
+\begin{equation} \kappa I_{\rm b} = \left\{ \begin{array}{cll}
+    \kappa \sigma T^4/\pi      & \hbox{Outside flame zone}, & \dot{q}'''=0  \\ [0.1in]
+    C \kappa \sigma T^4/\pi  & \hbox{Inside flame zone}, & \dot{q}'''>0
+    \end{array} \right.  \label{radapprox1}
+\end{equation}
+The constant $C$ is computed at each time step so that the volume integral of Eq.~(\ref{simple_rte}) over the entire flaming region is approximately equal to the volume integral of $\chi_{\rm r} \,\dot{q}'''$ over that same region.
+Here, $\chi_{\rm r}$ is an empirical estimate of the {\em global} fraction of that energy emitted as thermal radiation.
+Typically, a sooty fire radiates approximately one-third of the total combustion energy.
+
+The radiation equation is solved using a technique similar to a finite volume method for convective transport, thus the name given to it is the Finite Volume Method (FVM).
+Using approximately 100 discrete angles which are updated over multiple time steps, the finite volume solver requires about 20\% of the total CPU time of a calculation, a modest cost given the complexity of radiation heat transfer.
+
+Section~\ref{particle_model} discusses how Lagrangian particles are used to represent vegetation like leaves, pine needles and so on.
+These particles absorb and emit thermal radiation, and the underlying assumptions are described in that section.
+
+\subsection{Subgrid Parameterizations} \label{sec:subgrid}
+
+In the formal derivation of the LES equations (see, e.g, \cite{Pope:2000,FDS_Math_Guide}) the filtered nonlinear terms (advection terms, mean chemical source term, and the unresolved boundary fluxes) require subgrid-scale parameterizations.
+In this section we will temporarily utilize the LES filter formalism as well as Cartesian tensor index notation to clarify the terms used in the governing equations.
+
+\subsubsection{Subgrid Advection}
+
+Unresolved turbulent eddies enhance the transport of mass, momentum, and energy.
+This added transport is accounted for with residual stresses and scalar fluxes.
+The subgrid-scale stress tensor is defined as
+\begin{equation}
+\tau_{ij}^{sgs} \equiv \bar{\rho}(\widetilde{u_i u_j} - \tilde{u}_i \tilde{u}_j)
+\end{equation}
+where an overline represents a volumetric filter and the tilde represents a Favre filter (see, e.g., \cite{Poinsot:TNC}).
+
+In FDS the deviatoric part of the subgrid stress tensor is modeled using the gradient diffusion hypothesis.  Thus, the combined deviatoric stress may be written as follows:
+\begin{equation}
+\tau_{ij}^{dev} = \overline{\tau}_{ij} + \tau_{ij}^{sgs} - \frac{1}{3} \tau_{kk}^{sgs} = -2 (\mu + \mu_t) \left[ \tilde{S}_{ij} - \frac{1}{3} \tilde{S}_{kk} \right]
+\end{equation}
+where the symmetric rate-of-strain tensor is $S_{ij} \equiv \frac{1}{2}(\partial u_i/\partial x_j + \partial u_j/\partial x_i)$ and $\mu_t$ is an isotropic eddy viscosity.
+Note that $\overline{\tau}_{ij}$ is already the deviatoric part of the molecular stress as the isotropic part is the pressure; $\mu$ is the molecular dynamic viscosity.
+
+The eddy viscosity $\mu_t$ is an important quantity in the simulation as it affects all the turbulent transport and subgrid mixing time scales, which, as we will see below, are crucial to the combustion model.
+In FDS, the eddy viscosity is modeled as \cite{Pope:2000,Deardorff:1980}
+\begin{equation}
+\mu_t = \bar{\rho} C_\nu \Delta \sqrt{k_{sgs}}
+\end{equation}
+with the constant taken as $C_\nu = 0.1$ (this value can be derived assuming production equals dissipation and a model Kolmogorov spectrum \cite{Pope:2000}; we also test the model against decaying isotropic turbulence \cite{CBC}).
+In practice, the filter scale $\Delta$ is taken as the cube root of the cell volume.
+Deardorff \cite{Deardorff:1980} solved a transport equation for the subgrid kinetic energy per unit mass, $k_{sgs}$, in order to account for subgrid buoyancy.
+In FDS, the density field is sufficiently resolved to account for buoyant plume accelerations, and therefore we find that an algebraic closure for $k_{sgs}$ is sufficient.
+We utilize a scale-similarity idea, following the work of Bardina \cite{Bardina:1980}, and model the subgrid kinetic energy per unit mass as follows:
+\begin{equation}
+k_{sgs} = \frac{1}{2}(\tilde{u}_i - \hat{\tilde{u}}_i)(\tilde{u}_i - \hat{\tilde{u}}_i)
+\end{equation}
+where the hat represents a test filter at a scale $2\Delta$.
+For the first Cartesian cell off the wall, where a test filter is not well defined, the eddy viscosity is taken from the WALE model \cite{Nicoud:1999}.
+
+Subgrid advection of mass and heat are also modeled using gradient diffusion.
+The turbulent diffusivities use constant turbulent Schmidt and Prandtl numbers, both set to 0.5; this simplification is justified based on high resolution simulations of fire plumes (see, e.g., \cite{Maragkos:2020}).
+For mass diffusion of species $\alpha$ in direction $i$ we have
+\begin{equation}
+J_{\alpha,i} = \bar{J}_{\alpha,i} + J_{\alpha,i}^{sgs} = -\bar{\rho}(D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t}) \frac{\partial \tilde{Y}_\alpha}{\partial x_i}
+\end{equation}
+We use effective thermal conductivity and mass diffusivity in the heat flux, which may be written as
+\begin{equation}
+{\dot{q}''}_i = -(\bar{k} + \frac{\mu_t \bar{c}_p}{\mathrm{Pr}_t}) \frac{\partial \widetilde{T}}{\partial x_i} - \sum_\alpha \tilde{h}_{{\rm s},\alpha} ( \bar{\rho} D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t} ) \frac{\partial \tilde{Y}_\alpha }{\partial x_i}
+\end{equation}
+
+\subsubsection{Filtered Chemical Source Term}
+
+FDS treats combustion chemistry using a simplified approach.
+We track only Fuel, Air, and Products, which we refer to as ``lumped species''.
+Lumped species are groups of species that transport and react together and are always found in the same proportion.
+For example, Air is a lumped species made up of 23 \% O$_2$ and 77 \% N$_2$ by mass.
+For vegetation, a surrogate hydrocarbon Fuel is usually assumed.
+For example, cellulose may be assumed to decompose to a volatile fuel gas with the formula $\C_{6} \H_{10} \O_{5}$.
+The Products are defined stoichiometrically, often including prescribed yields of soot (for smoke) and carbon monoxide that may be obtained from empirical relationships for the specific fuel or from other sources (e.g., \cite{SFPE:Mulholland}).
+Then, usually, a single reaction is tracked\footnote{More complicated reaction schemes are possible in the code, but for this work a single reaction is sufficient.}:
+\begin{equation}
+\mathrm{Fuel} + s \, \mathrm{Air} \rightarrow (1+s) \,\mathrm{Products}
+\end{equation}
+where $s$ is the mass stoichiometric coefficient.
+
+In LES of turbulent combustion the filtered chemical source term, $\overline{\dot{m}_\alpha'''}$, is unclosed because the reaction kinetics are nonlinear functions of species concentration and temperature.
+However, since the heat releasing reactions in a fire are extremely fast compared to the time scales for turbulent mixing, it is common practice in fire dynamics simulations to utilize the eddy dissipation model of Magnussen and Hjertager \cite{Magnussen:1} (see also \cite{Poinsot:TNC}).
+The mass production term for the Fuel species is written as
+\begin{equation}
+\overline{\dot{m}_{F}'''} = -\bar{\rho} \,\frac{\min(\tilde{Y}_F,\tilde{Y}_A/s)}{\tau_{mix}}
+\end{equation}
+This formula states that the fuel is consumed at a rate proportional to the local limiting reactant concentration and inversely proportional to the local mixing time scale.
+This is the so-called ``mixed is burnt'' approximation.
+The mixing time scale is given by $\tau_{mix} = \min(\tau_d, \tau_u,\tau_g)$ \cite{McDermott:2011}, where the time scales for diffusion, turbulent advection, and gravitational acceleration are, respectively,
+\begin{align}
+\tau_d &= \Delta^2/D_F \\
+\tau_u &= C_u \Delta / \sqrt{(2/3)k_{sgs}} \\
+\tau_g &= \sqrt{2\Delta/g}
+\end{align}
+Here, $D_F$ is the binary diffusivity of the Fuel species in Air, $k_{sgs}$ is the modeled subgrid kinetic energy per unit mass discussed above, and $g$ is the gravitational acceleration.
+The constant $C_u=0.4$ is determined by matching flame height correlations across a wide range of fire Froude numbers \cite{FDS_Validation_Guide}.
+
+The remaining mass production rates are found from stoichiometry.  The local heat release rate, $\dot{q}'''$, an important term in Eq.~(\ref{eq:divth}), is then determined from
+\begin{equation}
+\label{eq:qdotppp}
+\dot{q}''' = - \sum_\alpha \overline{\dot{m}_\alpha'''} \Delta h_{f,\alpha}
+\end{equation}
+where $\Delta h_{f,\alpha}$ are the heats of formation for each lumped species $\alpha$.
+
+\paragraph{Flame Extinction} One of the advantages of a physics-based tool like FDS is the possibility of building into the model the effects of defensive actions on the fire spread rate.
+FDS uses simple empirical rules to predict local extinction within a gas phase grid cell based on resolved species concentrations and the mean cell temperature.
+The rules are based on the concept of a critical flame temperature.
+The basic theory behind the critical flame temperature is described in \cite{SFPE:Beyler}.
+In brief, if the local heat release given by Eq.~(\ref{eq:qdotppp}) is not sufficient to raise the cell temperature above the CFT, then $\overline{\dot{m}_{F}'''}$ is set to zero for that cell for that time step.
+Note that the flame extinction model does not directly affect surface cooling or smothering of the vegetative fuel, which may happen with certain fire suppression activities.
+However, such tactics can be accounted for through modifications to the solid fuel thermal degradation model discussed below.
+
+\subsubsection{Unresolved Boundary Fluxes}
+
+In this section we discuss the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries.
+Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main point of this paper).
+
+The scaled streamwise velocity is denoted $u^+ = u/u_\tau$, where the friction velocity is $u_\tau \equiv \sqrt{\tau_w/\rho}$ and $\tau_w$ is the wall stress needed to close the momentum equation, Eq.~(\ref{eqn:momtran}).
+For rough walls FDS employs the log law presented in \cite{Pope:2000},
+\begin{equation}
+\label{eqn_roughwallloglaw}
+u^+ = \frac{1}{\kappa} \ln \left(\frac{y}{s}\right) + \tilde{B}(s^+)
+\end{equation}
+where $\kappa = 0.41$ is the von K\'arm\'an constant, $s^+ = s/\delta_\nu$ is the roughness length in viscous units, $s$ is the dimensional ``sand grain'' roughness, and $\delta_\nu = \mu/(\rho u_\tau)$ is the viscous length scale.
+The distance to the wall, $y$, is taken as $\delta y/2$ for the first off-wall grid cell.
+The parameter $\tilde{B}$ varies with $s^+$ but attains a constant value $B_2=8.5$ in the fully rough limit.
+In FDS, we implement $\tilde{B}$ as the following piece-wise function:
+\begin{equation}
+\tilde{B} = \left\{ \begin{array}{lll} B + (1/\kappa)\ln(s^+) & \mbox{for} & s^+ < 5.83 \\
+\tilde{B}_{\rm max} & \mbox{for} & 5.83 \le s^+ < 30.0 \\
+B_2 & \mbox{for} & s^+ \ge 30.0 \end{array} \right.
+\end{equation}
+where $\tilde{B}_{\rm max} = 9.5$.
+
+In the fully rough limit the sand grain roughness may be equated to the aerodynamic roughness, $z_0$, typically employed in atmospheric codes for a neutral boundary layer,
+\begin{equation}
+s = z_0 e^{8.5 \kappa} \approx 32.6 \, z_0
+\end{equation}
+It is important to appreciate that FDS was originally designed to handle compartment fires and must consider flows with smooth walls such as HVAC ducts.
+Further, FDS typically resolves the thermal motions that would necessitate the use of Monin-Obukhov stability corrections to the boundary profile.
+
+The unresolved heat flux at the surface, $\dot{q}_w''$, is determined using an engineering approach,
+\begin{equation}
+\dot{q}_w'' = h (T_g - T_w) 
+\end{equation}
+where $h$ is the convective heat transfer coefficient, $T_w$ is the local temperature of the surface, and $T_g$ is the temperature of the first gas phase cell off the wall.
+Empirical natural/forced convection correlations are used to determine $h$:
+\begin{equation}
+h = \max \left[C\, |T_g-T_w|^{1/3} \; , \;
+\frac{k}{L} \, \mathrm{Nu} \; , \;
+\frac{k}{\delta n/2} \right]
+\label{eq:qconv}
+\end{equation}
+where $C$ is a empirical coefficient for natural convection (1.52 for a horizontal plate and 1.31 for a vertical plane or cylinder)~\cite{Holman:1}, $L$ is a characteristic length related to the size of the physical obstruction, and $k$ is the thermal conductivity of the gas.
+The forced convection Nusselt number (Nu) depends on the geometry and flow characteristics \cite{Holman:1,Incropera:1}.
+Further details are provided in \cite{FDS_Math_Guide}.
+
+Again, this simple approach is viable because FDS usually resolves the thermal flow fields that would necessitate added complexity.
+However, in future work we plan to explore improvements to wall functions that may lead to better results at coarser grid resolution.
+
+\section{Terrain Description and Discretization} \label{sec:terraindisc}
 
@@
 $\Delta x_{\rm cc}$ in Fig.~\ref{Fig:figure_2}) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase.
-Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces.
+Also, a normal probe approach~\cite{balaras_2004}, is employed to sample information from the fluid to define fluxes in boundary cut-faces.
+In a normal probe method, an external point is defined at a normal distance from the body of the order of the cartesian grid size.
+Then, information from the fluid at this external point is obtained via interpolation.
 \begin{figure}[ht]
@@
 \end{equation}
-where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{mcgratta_2013}.
-In the cut-cell region we use Godunov interpolation for the advective term.
-As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of Eqs.~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known.
+where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{FDS_Users_Guide}.
+In the cut-cell region a Godunov flux limited interpolation for the advective term is used.
+Flux limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic equations, of which mass transport inherits its mathematical properties.
+The time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of Eqs.~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known.
+Explicit Runge-Kutta time integration methods are single-step multistage schemes widely used for advancing ordinary and partial differential equations~\cite{leveque2007finite}.
 The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:
@@
 Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid.
-Details of the FDS time integration scheme can be found in references~\cite{mcgratta_2013,mcdermo_2014}.
+Details of the FDS time integration scheme can be found in references~\cite{FDS_Users_Guide,mcdermo_2014}.
 For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region.
@@
 
-Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within Eq.~(\ref{eqn:momtran}), the model momentum transport problem can be written~\cite{mcgratta_2013}:
+Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within Eq.~(\ref{eqn:momtran}), the model momentum transport problem can be written~\cite{FDS_Users_Guide}:
 \begin{eqnarray}
@@
 Known velocities and fluid parameters are interpolated from the surrounding fluid points $e_1,\dots,e_4$ to $e_x$.
-This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution~\cite{mcgratta_2013}.
+This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution~\cite{FDS_Users_Guide}.
 The target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux-matched to the underlying Cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$.
@@
 The Level Set Method relies on a set of predetermined spread rates for different types of vegetation and wind speeds.
+The pyrolysis model, developed specifically for vegetation~\cite{Porterie:2006,Morvan:CF2004,Houssami:2016}, consists of three reactions:
+
+\vspace{\baselineskip}
+
+\noindent 1. Endothermic moisture evaporation
+\begin{equation}
+ {\rm Wet\ Vegetation} \rightarrow \nu_{\rm H_2O} \, {\rm H_2O} + (1-\nu_{\rm H_2O}) \, {\rm Dry\ Vegetation} \quad ; \quad \nu_{\rm H_2O} = \frac{M}{1+M}
+\end{equation}
+
+\noindent 2. Endothermic pyrolysis of dry vegetation
+\begin{equation}
+ {\rm Dry\ Vegetation} \rightarrow \nu_{\rm char} \, {\rm Char} + (1-\nu_{\rm char}) \, {\rm Fuel\ Gas}
+\end{equation}
+
+\noindent 3. Exothermic char oxidation
+\begin{equation}
+ \label{char_reaction}
+ {\rm Char} + \nu_{\rm O_2, char} \, {\rm O_2} \rightarrow (1+ \nu_{\rm O_2,char} - \nu_{\rm ash}) \, {\rm CO_2} + \nu_{\rm ash} \, {\rm Ash}
+\end{equation}
+$M$ is the vegetation {\em moisture content} or {\em moisture fraction} determined on a dry weight basis.
+$\nu_{\rm char}$ is the mass fraction of dry vegetation that is converted to char during pyrolysis.
+$\nu_{\rm O_2,char}$ is the mass of oxygen required per unit mass of char consumed.
+$\nu_{\rm ash}$  is the mass fraction of char that is converted to ash during char oxidation.
+Reaction rates and the various empirical kinetic constants are given in Refs.~\cite{Porterie:2006,Morvan:CF2004,Houssami:2016}.
 
 \subsection{Particle Model}
+\label{particle_model}
 
-Lagrangian particles can be used to represent different types of vegetation, like leaves, grass, and so on.
-In the model, each type of vegetation is represented by a single particle in each computational grid cell.
-Most often the particle is assumed to be cylindrical in shape.
-The particle is assigned a diameter based on the measured surface area to volume ratio, $\sigma$.
+Lagrangian particles represent different types of vegetation, like leaves, grass, and so on.
+Each type of vegetation is represented by a single particle in each computational grid cell.
+Most often the particle is assumed to be cylindrical in shape, but it can be spherical or planar.
+The particle is assigned a diameter based on the measured surface area to volume ratio, $\sigma'$.
 The length is relatively unimportant so long as it is assumed to be much greater than the diameter.
@@
 \begin{equation}
-   \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \mathbf{u} \, \|\mathbf{u}\|
+   \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma' \, \mathbf{u} \, \|\mathbf{u}\|
    \label{drag}
 \end{equation}
-where $\rho$ is the air density, $C_{\rm d}$ is the drag coefficient, $C_{\rm s}$ is the shape factor (ratio of projected area to surface area), $\beta$ is the packing ratio, $\sigma$ is the surface area to volume ratio, and $\mathbf{u}$ is the air velocity.
-The diameter is specified via the surface area to volume ratio, $\sigma=2/r$ for a cylinder.
+where $\rho$ is the air density, $C_{\rm d}$ is the drag coefficient, $C_{\rm s}$ is the shape factor (ratio of projected area to surface area), $\beta$ is the packing ratio, $\sigma'$ is the surface area to volume ratio, and $\mathbf{u}$ is the air velocity.
+The diameter is specified via the surface area to volume ratio, $\sigma'=2/r$ for a cylinder.
 The packing ratio is the volume of solid needles divided by the volume they occupy, typically denoted $\beta$.
@@
 
-A coniferous tree (e.g., pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.
+The drag coefficient for miscellaneous vegetation is taken to be 2.8, based on wind tunnel measurements~\cite{Falkenstein-Smith:2018}.
+The shape factor is assumed to be 0.25, which is the exact value for spheres and an approximate value for randomly oriented cylinders.
+In practice, the single term
+\begin{equation}
+   \kappa = C_{\rm s}  \, \beta \, \sigma'
+\end{equation}
+can be easier to determine than the three individual terms by measuring the relative amount of sunlight, $W$, that penetrates through a layer of vegetation of depth $\delta$:
+\begin{equation}
+   \kappa = -\frac{\ln \, W}{\delta}
+\end{equation}
+This value of $\kappa$ also serves as the absorption coefficient for thermal radiation for a collection of particles.
+
+The Lagrangian particles that represent vegetation are assumed to be thermally-thick, but for simplicity the heat conduction within the particle is assumed to be one-dimensional in either a cylindrical or spherical coordinate system.
+It is assumed that the particles interact with the surrounding gas via an additional source term in the energy conservation equation.
+For a grid cell with indices $ijk$, the source term is:
+\begin{equation} \label{eq:qr}
+   (-\nabla \cdot \dot{\mathbf{q}}_{\rm r}'')_{ijk} = \sum \kappa_{\rm p} \left( U_{ijk} - 4 \sigma \, T_{\rm p}^4 \right)
+\end{equation}
+where $T_{\rm p}$ is the particle surface temperature and the summation is over all the particles within the cell. The effective absorption coefficient for a single particle is given by
+\begin{equation}
+   \kappa_{\rm p} = \frac{f \, A_{\rm p}}{4 \, V} = C_{\rm s}  \, \beta \, \sigma'
+\end{equation}
+where $A_{\rm p}$ is the surface area of the particle, $V$ is the volume of the cell, and $f$ is the number of actual particles represented by the single modeled particle.
+Notice that the absorption coefficient plays a role for both drag and radiation heat transfer.
+
+The net radiative heat flux onto the surface of the particle is taken as:
+\begin{equation}
+   \dot{q}_{\rm r}'' = \epsilon \left( \frac{U_{ijk}}{4} - \sigma T_{\rm p}^4 \right)
+\end{equation}
+where $\epsilon$ is the emissivity of the particle surface, typically assumed to be about 0.9. The convective heat transfer to the surface is given by:
+\begin{equation}
+   \dot{q}_{\rm c}'' = h \, (T_{\rm g} - T_{\rm s})  \quad ; \quad  h = \frac{k}{L} \, \hbox{Nu}  \label{q_con}
+\end{equation} 
+$L$ is a characteristic length equal to the diameter of the sphere or cylinder. The Nusselt number is given by:
+\begin{equation}
+   \hbox{Nu} = C_1 + C_2 \, \hbox{Re}^n \, \hbox{Pr}^m  \quad ; \quad \hbox{Re} = \frac{\rho |\mathbf{u}| L}{\mu} \quad ; \quad \hbox{Pr} = 0.7 
+\end{equation}
+For cylindrical particles, the default values are $C_1=0$, $C_2=0.683$, $n=0.466$, $m=0.33$, and $L=D$, the diameter of the cylinder.
+For spherical particles, the default values are $C_1=2$, $C_2=0.6$, $n=0.5$, $m=0.33$, and $L=D$, the diameter of the sphere~\cite{Incropera:1}.
 
@@
 \begin{equation}
-    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma \, \frac{h_{\rm b}}{\delta z} \, \mathbf{u} \, \|\mathbf{u}\|
+    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma' \, \frac{h_{\rm b}}{\delta z} \, \mathbf{u} \, \|\mathbf{u}\|
 \end{equation}
@@
 \begin{equation}
-   \kappa = C_{\rm s} \, \sigma \, \beta
+   \kappa = C_{\rm s} \, \sigma' \, \beta
 \end{equation}
@@
 \begin{equation}
-   \langle \dot{q}_{\rm c,b}''' \rangle = \sigma \, \beta \, \dot{q}_{\rm c}'' \quad ; \quad \dot{q}_{\rm c}'' = h \, (T_{\rm g} - T_{\rm s}) \quad \hbox{W/m}^2 \quad ; \quad  h = \frac{k}{L} \, {\rm Nu}
+   \langle \dot{q}_{\rm c,b}''' \rangle = \sigma' \, \beta \, \dot{q}_{\rm c}'' 
 \end{equation}
-where $\dot{q}_{\rm c}''$ the convective heat transferred to a small, cylindrical solid under the assumption that the subgrid-scale vegetation is cylindrical and the gas velocity and temperature are taken from the first gas phase grid cell adjacent to the boundary.
+where $\dot{q}_{\rm c}''$ is given by Eq.~(\ref{q_con}) and the gas velocity, $|\mathbf{u}|$, and temperature, $T_{\rm g}$, are taken from the first gas phase grid cell adjacent to the boundary, and the particle surface temperature, $T_{\rm s}$, is extracted from the solution of the 1-D heat conduction equation.
 
@@
 
-\section{Defining atmospheric wind : The Mean Forcing Concept} \label{sec:wind}
+\section{Atmospheric Wind Boundary Conditions}
+\label{sec:wind}
 
-The process of steering the solution of the mass, momentum, and energy equations to match wind speed and direction gathered at random weather stations is known as \emph{data assimilation}~\cite{Kalnay:2003}.
-FDS uses a relatively simple data assimilation technique, a method known as \emph{nudging}, where a forcing term is added to the momentum equation to ``nudge'' the flow in the direction of the specified wind.
-FDS automatically drives the mean velocity components toward the desired values by adding a forcing term to the momentum equation.
-For example, the $u$ component equation is modified as follows:
+For outdoor flows FDS prescribes the wind field through special ``open'' boundary conditions on the exterior of the computational domain.  The boundary conditions consist of three components:
+\begin{enumerate}
+\item[(i)] the specified upstream wind field (vertical profile of streamwise velocity components) based on prescribed Monin-Obukhov parameters imposed on fluid elements \emph{entering} the domain,
+\item[(ii)] optional specification of upstream turbulence based on Jarrin's synthetic eddy method \cite{Jarrin:2008} (which is possible with the code, but not utilized in the test cases within this paper), and
+\item[(iii)] nonuniform and nonstationary Dirichlet pressure boundary values for the Poisson equation.
+\end{enumerate}
+
+\subsection{Velocity and Temperature Profiles}
+\label{sec:veltmpprof}
+
+While completely custom vertical boundary profiles of velocity and temperature are possible, usually the wind field is specified via Monin-Obukhov (MO) similarity parameters \cite{Dyer:1974}: Obukhov length scale, $L$, aerodynamic roughness height, $z_0$, wind speed, $u_r$, and temperature, $T_r$, at a reference height $z_r$.
+For a given set of MO parameters, FDS generates velocity and temperature profiles as functions of height as shown in Fig.~\ref{fig:MOprofs}.
+These values are specified for any fluid elements entering the domain at height $z$, including along the top of the domain.
+Convective outflow boundaries are applied locally for any fluid elements leaving the domain.
+The profiles can vary in time allowing one-way ``downscaling'' from either weather observations or output from a numerical weather prediction (NWP) model.
+However, two-way coupling with an NWP model (e.g., \cite{Coen:2013,CAWFE}) has not yet been implemented with FDS.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=.45\textwidth]{figures/vel_L=-500.pdf}
+\includegraphics[width=.45\textwidth]{figures/tmp_L=-500.pdf}
+\caption{Monin-Obukhov velocity and temperature profiles for inflow boundary conditions.
+Here parameters are $L=-500\;\mathrm{m}$, $z_0=0.03\;\mathrm{m}$, and $u_r=4.8\;\mathrm{m/s}$ and $T_r=34\;^\circ\mathrm{C}$.}
+\label{fig:MOprofs}
+\end{figure}
+
+\subsection{Turbulent Inflow Conditions}
+
+It is possible in FDS to superimpose synthetic turbulence on a specified inlet profile utilizing Jarrin's synthetic eddy method (SEM) \cite{Jarrin:2008}.
+An example verification case is presented in Fig.~\ref{fig:semprofs}.
+The left image of the figure shows contours of the normal component of velocity at the inlet.
+The plot on the right shows the mean and rms profiles compared to the specified inlet parameters.
+The domain is 20 m in the streamwise dimension by 12 m in the spanwise by 13 m in height.
+In The turbulence intensity is set to 10 \% and the turbulence integral length scale is set to 3 m.
+The specified MO parameters are $L=-667\;\mathrm{m}$, $z_0=0.022\;\mathrm{m}$, and $u_r=10\;\mathrm{m/s}$ and $T_r=20\;^\circ\mathrm{C}$ at reference height of 10 m.
+The simulation runs for 60 seconds with statistics collected over the last 30 seconds.
+The lateral, top, and downstream boundaries are set to ``open'' pressure boundaries as discussed in the next section.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.45\textwidth,valign=c]{figures/sem_open_wind_contours.pdf}
+\includegraphics[width=0.5\textwidth,valign=c]{figures/sem_open_wind_u_prof.pdf}
+\caption{Example of inflow synthetic turbulence superimposed on a Monin-Obukhov boundary layer profile.
+Contours of the normal component of velocity at the domain inlet taken at time midway through the simulation (left); mean and rms profiles along the inlet compared with specified profiles (right).
+The MO parameters are $L=-667\;\mathrm{m}$, $z_0=0.022\;\mathrm{m}$, and $u_r=10\;\mathrm{m/s}$ and $T_r=20\;^\circ\mathrm{C}$ at reference height of 10 m.
+SEM turbulence parameters were 10 \% turbulence intensity and an eddy integral length scale of 3 m.}
+\label{fig:semprofs}
+\end{figure}
+
+Other approaches to imposing boundary layer turbulence are possible (e.g., \cite{Munoz-Esparza:2014}).
+The cases presented later in this paper do not utilize SEM since detailed data on the turbulence statistics were not available.
+
+\subsection{Pressure Boundary Values}
+
+The Poisson equation for $H$, Eq.~(\ref{it:FSPoisson}), links the momentum equation with mass and energy in the low-Mach flow algorithm.
+The boundary conditions for this equation determine the inflow and outflow velocity component values.
+Let $\mathbf{n}$ denote a unit normal vector pointing \emph{into} the domain.
+Boundary values at an interface are denoted with a subscript ``$I$''.
+Then $\mathbf{u}_I\cdot\mathbf{n}>0$ is an inflow and $\mathbf{u}_I\cdot\mathbf{n}<0$ is an outflow condition.
+
+At an inflow, generally the mean viscous and convective forces are small and we make use of a simplified momentum equation that approximates $\partial u_i/\partial t = -\partial H/\partial x_i$.
+Let $u$ denote the $x$-component of velocity and consider an interface normal to $x$ pointing into the domain.
+The prescribed external wind field velocity component at height $z$ and time $t$ (discussed above in Sec.~\ref{sec:veltmpprof}) is denoted $u_{wind}(z,t)$.
+Using a one-sided difference for $H$ and an explicit Euler approximation of the velocity time derivative, the boundary value for $H$ may be written as
 \begin{equation}
-   \frac{\partial u}{\partial t} + \ldots = \frac{u_0(z,t)-\overline{u}(z,t)}{\tau} \label{mean_forcing_u}
+\label{eq:Hin}
+H_I = H_{I\pm\frac{1}{2}}^n \pm \frac{\Delta x}{2}\left[\frac{u_{wind,I}(z,t) - u_I^n}{\Delta t}\right] \quad \mbox{if} \quad \mathbf{u}_I\cdot\mathbf{n}>0
 \end{equation}
-Here, $u(\mathbf{x},t)$ is the computed velocity component, $u_0$ is the specified wind field component that can vary with height and time, and $\overline{u}$ is the average of the computed velocity component at the height $z$.
-The relaxation time scale, $\tau$, is an important parameter.
-The shorter this time scale, the faster the flow field will ``catch up'' to the mean wind, but at the expense of possibly washing out important flow structures.
+At an outflow boundary, $H$ is set equal to the local kinetic energy per unit mass from the cell just upwind of the boundary:
+\begin{equation}
+\label{eq:Hout}
+H_I = \frac{1}{2}(\bar{u}^n \bar{u}^n + \bar{v}^n \bar{v}^n + \bar{w}^n \bar{w}^n)_{I\pm\frac{1}{2}} \quad \mbox{if} \quad \mathbf{u}_I\cdot\mathbf{n}<=0
+\end{equation}
+The overbar on the velocity components denotes a linear interpolation of the primitive staggered component values to the cell center.
 
@@
 
-Two of these experiments were originally simulated with FDS by Mell~et~al.~\cite{Mell:IJWF2007}.
-These simulations modeled the grass as a collection of cylindrical Lagrangian particles.
-Now these two experiments are also simulated using the Boundary Fuel Model (BFM)~\cite{Perez-Ramirez:FT2017} and the Rothermel-Albini fire spread algorithm~\cite{Rothermel:1972,Albini:1976}.
-For the experiment labelled Case~C064, fuel index 1 (Short Grass) is used, with a modified moisture fraction of 0.063.
-For F19, fuel index 3 (Tall Grass) is used, with a modified moisture fraction of 0.058.
+Two of these experiments were originally simulated with FDS by Mell~et~al.~\cite{Mell:IJWF2007} using a form of the Boundary Fuel Model.
+Now these two experiments are also simulated using the Lagrangian Particle Model and the Level Set Model.
+The level set simulations of Case~C064 use fuel index 1 (Short Grass) and for Case~F19, fuel index 3 (Tall Grass)~\cite{Rothermel:1972,Albini:1976}.
 
-Measured properties for the specific types of grasses burned in the two experiments are listed in Table~\ref{Properties_Grasses}.
+Measured bulk properties of the grasses burned in the two experiments are listed in Table~\ref{Properties_Grasses}.
 Properties that were not measured are listed in Table~\ref{Assumed_Properties_Grasses}.
@@
 
-\begin{figure}[ht]
+\begin{figure}[p]
 \includegraphics[width=\textwidth]{figures/F19_collage.png}
@@
 \label{F19}
+\end{figure}
+
+\begin{figure}[ht]
+\includegraphics[width=0.49\textwidth]{figures/Case_F19_flame_position.pdf}
+\includegraphics[width=0.49\textwidth]{figures/Case_C064_flame_position.pdf}
+\caption{Contours of heat release rate in a plane near the surface compared to observations of CSIRO Grassland Fires F19 (left) and C064 (right).
+FDS contour times for Case C064 have been shifted uniformly by 10 s to line up the initial head fire observation and compare the subsequent spread rate.}
+\label{fig:CaseF19_contours}
 \end{figure}
@@
 The weighting factor is determined from the measured bulk mass per unit area.
-The fires in the experiments were ignited by two men carrying drip torches walking in opposite directions along the upwind boundary of the plot (the red strip in Fig.~\ref{F19}).
+The fires in the experiments were ignited by two field workers carrying drip torches walking in opposite directions along the upwind boundary of the plot (the red strip in Fig.~\ref{F19}).
 In FDS, this action was modeled using a specified spread rate along the strip.
@@
 Measured ROS                    & m/s           & 1.5           & 1.1           \\ \hline
-Calculated ROS, Particle Method & m/s           & 1.5           & 1.0           \\ \hline
-Calculated ROS, Boundary Fuel Method & m/s      & 3.7           & 2.5           \\ \hline
-Calculated ROS, Level Set Method & m/s          & 1.3           & 0.7           \\ \hline
+Calculated ROS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s           & 1.4, 1.3, 1.4           & 1.1, 1.2, 1.2          \\ \hline
+Calculated ROS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s      & 1.5, 1.4, 1.7           & 1.3, 1.3, 1.3           \\ \hline
+Calculated ROS, Level Set Method (5 m, 10 m, 20 m resolution) & m/s          & 1.1, 1.1, 1.2          & 0.4, 0.5, 0.8          \\ \hline
 \end{tabular}
@@
 
-The predicted rates of spread for the two cases are listed in Table~\ref{Properties_Grasses}.
-For these simulations, the particle method is the most accurate, and the boundary fuel method over-predicts by about a factor of two.
-The level set method under-predicts the rate of spread, but only because of the somewhat arbitrary choice of fuel type.
-For a simple flat terrain scenario like this, the choice of fuel model to accompany the level set calculation largely determines the spread rate.
-The original FDS simulations of this experiment ~\cite{Mell:IJWF2007} also used a boundary fuel approach, with better results, work is underway to reconcile the differences.
-It is not surprising that the particle method does well in this scenario because it is closest to the actual physics.
-However, for larger domains, the boundary fuel and level set methods are better able to approximate the fire spread while the particle method would require an excessive amount of computer resources.
+\begin{figure}[p]
+\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
+\includegraphics[height=2.2in]{figures/Case_C064} &
+\includegraphics[height=2.2in]{figures/Case_F19} \\
+\includegraphics[height=2.2in]{figures/Case_C064_BFM} &
+\includegraphics[height=2.2in]{figures/Case_F19_BFM} \\
+\includegraphics[height=2.2in]{figures/Case_C064_LS}  &
+\includegraphics[height=2.2in]{figures/Case_F19_LS}
+\end{tabular*}
+\caption{Comparison of the measured and predicted fire front position for the CSIRO Grassland Fires using three different methods of fire spread.}
+\label{CSIRO}
+\end{figure}
+
+The simulations have been conducted with the three different fire spread models, and each case has been run with three levels of spatial resolution.
+The simulations using the Particle and Boundary Fuel Models have been run with cubic grid cells that are 0.25~m, 0.5~m, and 1~m on a side.
+The simulations using the Level Set Model are run with cells of 5~m, 10~m, and 20~m.
+The predicted rates of spread for the two cases are listed in Table~\ref{Properties_Grasses} and shown graphically in Fig.~\ref{CSIRO}.
+For these simulations, the rate of spread predicted by the Particle and Boundary Fuel Models are comparable to the measured rate of spread while the level set method under-predicts the RoS.
+The under-prediction by the level set simulations is a consequence of the choice of fuel model, and one should not draw any conclusions as to the performance of level sets versus the other two rate of spread methods.
+
+It is difficult to compare the Particle and Boundary Fuel simulations with those conducted by Mell et al.
+in 2007~\cite{Mell:IJWF2007}.
+The underlying numerical model, FDS, has undergone a major revision over the past decade, changing the basic finite-difference scheme, boundary conditions, drag modeling, and so on, and the vegetation-specific models have changed as well, most notably the charring reaction.
+In addition, the amount of information about the grasses and field conditions is limited, and the thermal and kinetic parameters are gleaned from various sources and are generally average values over a wide range of vegetation types.
+A sensitivity study~\cite{McGrattan:CI2017} for the particle model simulations reveals that near-surface wind speed is the has the most direct effect on the RoS, and the combined effect of the dozens of thermal, kinetic, and numerical parameters would certainly explain the difference in RoS.
 
@@
 
+This section presents an example of the methodologies discussed in the paper.
+It is notable because it considers an actual fire that occurred in the urban-wildland interface, which presents a challenge for a fire spread model because the surface vegetation is mixed with the built environment.
+
 At approximately 23:00 (local time) on March 25, 2019, a wildfire started in the seaside town of Cogoleto, near Genova, Italy.
@@
 
-The simulation results were compared with those obtained from CIMA Propagator~\cite{Trucchia:2020}, an experimental propagation model that has provided several European civil protection organizations with real-time fire predictions since 2009.
+The simulation results have been compared with those obtained from CIMA Propagator~\cite{Trucchia:2020}, an experimental propagation model that has provided several European civil protection organizations with real-time fire predictions since 2009.
 The model has been implemented by the CIMA Foundation, its propagation model is based on stochastic cellular automata.
+
+The two models produce a qualitatively similar patterns of fire spread, mainly because the vicinity of the fire has numerous areas that are considered ``non-combustible,'' that is, devoid of vegetation or structures.
+This points out the problem of ``validating'' wildfire models.
+The simulations of the flat terrain grassland fire experiments described in the previous section match fairly well the observed fire behavior, but in these cases, the vegetation is uniform and fairly well described in terms of its mass loading and geometrical characteristics, the terrain is flat and of relatively small area, the wind is steady, and the experiment lasts a few minutes.
+In a real fire, none of this will be true.
+The detailed physics of the fire are not resolvable on a relatively coarse grid, the vegetation is inhomogeneous, sparse, and sometimes unknown, the weather conditions are varying, there are suppression efforts with water and the setting of backfires, and the fires can burn for days.
+All of which raise the question as to what level of physical fidelity ought to be required in these models.
+The strategy that has been adopted in FDS is to accommodate simulation options ranging from a level set calculation of fire spread over tens of kilometers that can be run in minutes, similar to models like BEHAVE developed by the U.S.
+Forest Service, all the way to detailed simulations of fire spread at sub-meter resolution for relatively short time periods (minutes to hours) and small areas (tens of hectares).
+As was the case for models developed for building fires thirty years ago, it is not clear at the moment exactly how these wildland fire models will be used in the future.
+The best strategy is to advance the various modeling options until a clear path forward becomes apparent.
 
@@
 \externalbibliography{yes}
-\bibliography{Vanella_Article_Atmosphere_2020,LOFM_Workshop_References}
+\bibliography{Vanella_Article_Atmosphere_2020,LOFM_Workshop_References,FDS_refs}
 

### 2020-12-09 Response_to_A_Maranghides.tex
Update on Overleaf.

@@
 
-   From origin to furthest point measured distance is ~ 1200 m.
+   \hl{This info was obtained from the simulation domain. From origin to furthest point measured distance is about 1200 m. The fire run overnight and onto the next day.}
 
@@
 
+   \hl{From the information provided from CIMA : Winds recorded at nearby sensor station were found predominantly coming from the north with magnitudes that ranged from 20 m/sec during the night and diminished to 5 m/sec towards the morning hours.}
+
    \item information on defensive actions.
@@
 I do not know what was used and when.
+
+   \hl{Information from https://www.thelocal.it/20190326/residents-evacuated-as-wildfires-take-hold-near-genoa : Up to 70 firefighters worked at the scene, with help from helicopters and fire hydrants.
+Suppression work was directed into stopping fire spread into the coastal town of Cogoleto.}
 
@@
 
+   \hl{Ema, private communication : There's been a lot of spotting.
+Added : Further, spot ignition was reported by first responders at the scene, a common fire spread mechanism that remains a major issue in wildfire modeling.}
+
    \item Explicit listing of the model limitations of dealing with fuel breaks both natural and manmade.
@@
 
+    \hl{Kevin : For the particle and boundary fuel models of flame spread, fuel breaks are handled directly by removing combustible vegetation from the model.
+For the level set method, a fire break is a region declared as "non-combustible." The 13 Rothermel fuel types have been supplemented by "water", "ice", "pavement" and various other non-combustible regions that would be tiled onto our terrain map.}
+
    \item I would also like to see some statement explicitly addressing the complexity of dealing with non-uniform fuels (not only vegetative but also WUI).
 Even using the level set approach one needs to be mindful of the potential change in fire spread as the fire migrates between different fuels.
+
+   \hl{Non-uniform fuels : Adressed in the statement on the complexity of simulation real WUI fires (last paragraph of cogoleto section.)}
 

### 2020-12-09 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 \usepackage[export]{adjustbox}
+\usepackage{url}
 
@@
 This scalar field is called the level set function and numerical methods that compute the function are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}.
-Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,FDS_Users_Guide,LAUTENBERGER_2013}.
-A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.
+Some wildfire solvers that use LSM to track the fire front are described in Refs.~\cite{coen_2013,Bova:IJWF2015,FDS_Users_Guide,LAUTENBERGER_2013}.
+A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in Ref.~\cite{Bova:IJWF2015}.
 
@@
 For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind.
-For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model.
+For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a level set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model.
 The wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution.
@@
 For fire modeling in the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}.
-Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used in the equation of state (ideal gas law), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.
+Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background hydrostatic pressure field $\bar{p}(z,t)$ used in the equation of state (ideal gas law) and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ driving fluid motion.
 
@@
 Therefore, to enforce realizability, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}.
-The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification.
+The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$;  the term $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification.
 These correspond to the amount of mass per unit volume and time of species $\alpha$ being added or substracted in a given point $\mathbf{x}$ due to chemical reaction or gasification of solid particles (or droplets), respectively.
+Details of the combustion model are provided below in Sec.~\ref{sec:subgrid}.
+
 In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is the modeled height dependent background density of the atmosphere.
@@
 
-Section~\ref{particle_model} discusses how Lagrangian particles are used to represent vegetation like leaves, pine needles and so on.
+Section~\ref{particle_model} discusses how Lagrangian particles are used to represent vegetation like leaves, pine needles and other subgrid objects.
 These particles absorb and emit thermal radiation, and the underlying assumptions are described in that section.
@@
 In the formal derivation of the LES equations (see, e.g, \cite{Pope:2000,FDS_Math_Guide}) the filtered nonlinear terms (advection terms, mean chemical source term, and the unresolved boundary fluxes) require subgrid-scale parameterizations.
-In this section we will temporarily utilize the LES filter formalism as well as Cartesian tensor index notation to clarify the terms used in the governing equations.
+In this section, LES filter notation as well as Cartesian tensor index notation are used to clarify the terms used in the governing equations.
 
@@
 
-The eddy viscosity $\mu_t$ is an important quantity in the simulation as it affects all the turbulent transport and subgrid mixing time scales, which, as we will see below, are crucial to the combustion model.
+The eddy viscosity $\mu_t$ is an important quantity in the simulation as it affects all the turbulent transport and subgrid mixing time scales, which, as shown below, are crucial to the combustion model.
 In FDS, the eddy viscosity is modeled as \cite{Pope:2000,Deardorff:1980}
@@
 \end{equation}
-with the constant taken as $C_\nu = 0.1$ (this value can be derived assuming production equals dissipation and a model Kolmogorov spectrum \cite{Pope:2000}; we also test the model against decaying isotropic turbulence \cite{CBC}).
+with the constant taken as $C_\nu = 0.1$ (this value can be derived assuming production equals dissipation and a model Kolmogorov spectrum \cite{Pope:2000}; the model is also tested against decaying isotropic turbulence \cite{CBC}).
 In practice, the filter scale $\Delta$ is taken as the cube root of the cell volume.
 Deardorff \cite{Deardorff:1980} solved a transport equation for the subgrid kinetic energy per unit mass, $k_{sgs}$, in order to account for subgrid buoyancy.
-In FDS, the density field is sufficiently resolved to account for buoyant plume accelerations, and therefore we find that an algebraic closure for $k_{sgs}$ is sufficient.
-We utilize a scale-similarity idea, following the work of Bardina \cite{Bardina:1980}, and model the subgrid kinetic energy per unit mass as follows:
+In FDS, the density field is sufficiently resolved to account for buoyant plume accelerations, and therefore it is found that an algebraic closure for $k_{sgs}$ is sufficient.
+Following the work of Bardina \cite{Bardina:1980}, a scale-similarity idea is used; the subgrid kinetic energy per unit mass is modeled as follows:
 \begin{equation}
@@
 The turbulent diffusivities use constant turbulent Schmidt and Prandtl numbers, both set to 0.5; this simplification is justified based on high resolution simulations of fire plumes (see, e.g., \cite{Maragkos:2020}).
-For mass diffusion of species $\alpha$ in direction $i$ we have
+For mass diffusion of species $\alpha$ in direction $i$ the flux is
 \begin{equation}
@@
 \end{equation}
-We use effective thermal conductivity and mass diffusivity in the heat flux, which may be written as
+Effective thermal conductivity and mass diffusivity are used in the heat flux, which may be written as
 \begin{equation}
@@
 FDS treats combustion chemistry using a simplified approach.
-We track only Fuel, Air, and Products, which we refer to as ``lumped species''.
+Only Fuel, Air, and Products, referred to as ``lumped species'', are tracked.
 Lumped species are groups of species that transport and react together and are always found in the same proportion.
@@
 \subsubsection{Unresolved Boundary Fluxes}
-
-In this section we discuss the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries.
+\label{sec:boundflx}
+In this section, the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries are discussed.
 Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main point of this paper).
@@
 The parameter $\tilde{B}$ varies with $s^+$ but attains a constant value $B_2=8.5$ in the fully rough limit.
-In FDS, we implement $\tilde{B}$ as the following piece-wise function:
+In FDS, $\tilde{B}$ is implemented as the following piece-wise function:
 \begin{equation}
@@
 Again, this simple approach is viable because FDS usually resolves the thermal flow fields that would necessitate added complexity.
-However, in future work we plan to explore improvements to wall functions that may lead to better results at coarser grid resolution.
+However, future work is planned to explore improvements to wall functions that may lead to better results at coarser grid resolution.
 
@@
 
-\subsection{Scalar transport and energy discretization in unstructured region}
+\subsection{Cut-cell Scalar Transport and Energy Discretization}
 
@@
 \end{equation}
-For a control volume that does not change with time, the time derivative and source terms are approximated by
+For a cut-cell control volume the time derivative and source terms are approximated by
 \begin{equation}
-  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \widetilde{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
-  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons}
+  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \overline{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
+  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\overline{ \dot{m}_\alpha''' }_{ii}+ \overline{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons}
 \end{equation}
-where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages.
-In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged.
+where $V_{ii}$ is the volume of cell $ii$ and the overlines imply cell averages.
+In the following the notation is simplified by dropping overlines, observing that quantities are cell or face averaged.
 Consider the FV discretization of the diffusive term of Eq.~(\ref{eq:intconvdiff}) on cut-cell $ii$ of Fig.~\ref{Fig:figure_2}:
@@
 In the cut-cell region a Godunov flux limited interpolation for the advective term is used.
-Flux limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic equations, of which mass transport inherits its mathematical properties.
+Flux limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic equations~\cite{leveque_2002}, of which mass transport inherits its mathematical properties.
 The time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of Eqs.~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known.
@@
 Details of the FDS time integration scheme can be found in references~\cite{FDS_Users_Guide,mcdermo_2014}.
-For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region.
+For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) is recomputed on the unstructured cut-cell region.
 A similar procedure is done for the thermodynamic divergence.
 % Small cell linking for explicit time integration.
-As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step.
+As an explicit time integrator is used, in general there will arise cut-cells whose small size will penalize severely the time step.
 These cut-cells are linked to larger surrounding cells.
@@
 
-\subsection{Immersed boundary method on cut-cell region and wall modeling}
+\subsection{Immersed Boundary Method and Wall Modeling}
 
@@
 As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$.
-Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th}$ known
+Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1}$ are known:
 \begin{eqnarray}
   \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \mathbf{F}^n -  \boldsymbol{\nabla} H^n \label{eq:LowMachMomEu}\\
-  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} \label{eq:LowMachDivEu}
+  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1} \label{eq:LowMachDivEu}
 \end{eqnarray}
@@
 This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator.
-As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible for enforcing the divergence condition and used on the projection step.
+The potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is responsible for enforcing the divergence condition and used on the projection step.
 Taking the divergence of Eq.~\eqref{eq:LowMachMomEu} and considering  the constraint, Eq.~\eqref{eq:LowMachDivEu}, the two steps of the method are
@@
 \begin{equation}
-   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
+   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
 \end{equation}
@@
 \end{enumerate}
-A consequence of the projection scheme is that boundary conditions are required on the Poisson equation, \eqref{it:FSPoisson}.
+A consequence of the projection scheme is that boundary conditions are required on the Poisson equation, Eq.~\eqref{it:FSPoisson}.
 For explicit methods and stationary \textit{solid} boundaries, the corresponding boundary condition is \textit{homogeneous} Neumann for $H^n$ in $\partial \Omega,\partial \Omega_1,...,\partial \Omega_{\rm nbods}$~\cite{perot_1993}.
@@
 Known velocities and fluid parameters are interpolated from the surrounding fluid points $e_1,\dots,e_4$ to $e_x$.
-This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution~\cite{FDS_Users_Guide}.
+This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution of section~\ref{sec:boundflx}.
 The target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux-matched to the underlying Cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$.
@@
 
-Lagrangian particles represent different types of vegetation, like leaves, grass, and so on.
-Each type of vegetation is represented by a single particle in each computational grid cell.
+Lagrangian particles represent different types of vegetation, like leaves, grass, and pine needles.
+Each type of vegetation is represented by a single particle in each computational grid cell\footnote{Multiple particles within a cell are permitted, but may not be the most computationally efficient approach.}.
 Most often the particle is assumed to be cylindrical in shape, but it can be spherical or planar.
@@
 The vectors $\boldsymbol{\phi}_W,\boldsymbol{\phi}_S \ge 0$ are computed using the local wind and slope through empirical rules~\cite{Wilson:1980}.
-See Appendix 1 of reference~\cite{Bova:IJWF2015} for details on these calculations.
+See Appendix 1 of Ref~\cite{Bova:IJWF2015} for details on these calculations.
 In FDS, the level set spread model can be applied in various ways:
@@
 \centering
-\includegraphics[width=0.45\textwidth,valign=c]{figures/sem_open_wind_contours.pdf}
-\includegraphics[width=0.5\textwidth,valign=c]{figures/sem_open_wind_u_prof.pdf}
+\includegraphics[width=0.5\textwidth,valign=c]{figures/sem_open_wind_contours.pdf}
+\includegraphics[width=0.48\textwidth,valign=c]{figures/sem_open_wind_u_prof.pdf}
 \caption{Example of inflow synthetic turbulence superimposed on a Monin-Obukhov boundary layer profile.
-Contours of the normal component of velocity at the domain inlet taken at time midway through the simulation (left); mean and rms profiles along the inlet compared with specified profiles (right).
-The MO parameters are $L=-667\;\mathrm{m}$, $z_0=0.022\;\mathrm{m}$, and $u_r=10\;\mathrm{m/s}$ and $T_r=20\;^\circ\mathrm{C}$ at reference height of 10 m.
-SEM turbulence parameters were 10 \% turbulence intensity and an eddy integral length scale of 3 m.}
+Contours of the normal component of velocity at the domain inlet taken at a time midway through the simulation (left); mean and rms profiles along the inlet compared with specified profiles (right).}
 \label{fig:semprofs}
@@
 
-At an inflow, generally the mean viscous and convective forces are small and we make use of a simplified momentum equation that approximates $\partial u_i/\partial t = -\partial H/\partial x_i$.
-Let $u$ denote the $x$-component of velocity and consider an interface normal to $x$ pointing into the domain.
+At an inflow, generally the mean viscous and convective forces are small and a simplified momentum equation   may be used to develop a boundary value for $H$, namely, $\partial u_i/\partial t = -\partial H/\partial x_i$.
+To develop the boundary condition, let $u$ denote the $x$-component of velocity and consider an interface normal to $x$ pointing into the domain.
 The prescribed external wind field velocity component at height $z$ and time $t$ (discussed above in Sec.~\ref{sec:veltmpprof}) is denoted $u_{wind}(z,t)$.
@@
 \includegraphics[width=\textwidth]{figures/F19_collage.png}
-\caption{Photographs of the experiment and snapshots of the simulation of CSIRO Grassland Fire F19, 56~s, 86~s, and 138~s following ignition.}
+\caption{Photographs of the experiment and snapshots of the simulation (medium resolution Particle Model) of CSIRO Grassland Fire F19, 56~s, 86~s, and 138~s following ignition.}
 \label{F19}
-\end{figure}
-
-\begin{figure}[ht]
-\includegraphics[width=0.49\textwidth]{figures/Case_F19_flame_position.pdf}
-\includegraphics[width=0.49\textwidth]{figures/Case_C064_flame_position.pdf}
-\caption{Contours of heat release rate in a plane near the surface compared to observations of CSIRO Grassland Fires F19 (left) and C064 (right).
-FDS contour times for Case C064 have been shifted uniformly by 10 s to line up the initial head fire observation and compare the subsequent spread rate.}
-\label{fig:CaseF19_contours}
 \end{figure}
@@
 \hline
-Property                        & Units         & Case F19      & Case C064     \\ \hline \hline
-Wind Speed                      & m/s           & 4.8           & 4.6           \\ \hline
-Ambient Temperature             & $^\circ$C     & 34            & 32            \\ \hline
-Surface Area to Volume Ratio    & m$^{-1}$      & 12240         & 9770          \\ \hline
-Grass Height                    & m             & 0.51          & 0.21          \\ \hline
-Bulk Mass per Unit Area         & kg/m$^2$      & 0.313         & 0.283         \\ \hline
-Moisture Fraction               & \%            & 5.8           & 6.3           \\ \hline
-Measured ROS                    & m/s           & 1.5           & 1.1           \\ \hline
-Calculated ROS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s           & 1.4, 1.3, 1.4           & 1.1, 1.2, 1.2          \\ \hline
-Calculated ROS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s      & 1.5, 1.4, 1.7           & 1.3, 1.3, 1.3           \\ \hline
-Calculated ROS, Level Set Method (5 m, 10 m, 20 m resolution) & m/s          & 1.1, 1.1, 1.2          & 0.4, 0.5, 0.8          \\ \hline
+Property                        & Units        & Case C064 & Case F19    \\ \hline \hline
+Wind Speed                      & m/s          & 4.6       & 4.8         \\ \hline
+Ambient Temperature             & $^\circ$C    & 32.       & 34          \\ \hline
+Surface Area to Volume Ratio    & m$^{-1}$     & 9770      & 12240       \\ \hline
+Grass Height                    & m            & 0.21      & 0.51        \\ \hline
+Bulk Mass per Unit Area         & kg/m$^2$     & 0.283     & 0.313       \\ \hline
+Moisture Fraction               & \%           & 6.3       & 5.8         \\ \hline
+Measured RoS                    & m/s          & 1.1       & 1.5         \\ \hline
+Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s           & 1.4, 1.3, 1.4           & 1.1, 1.2, 1.2          \\ \hline
+Calc'd RoS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s      & 1.5, 1.4, 1.7           & 1.3, 1.3, 1.3           \\ \hline
+Calc'd RoS, Level Set Method (5 m, 10 m, 20 m resolution) & m/s          & 1.1, 1.1, 1.2          & 0.4, 0.5, 0.8          \\ \hline
 \end{tabular}
@@
 
-It is difficult to compare the Particle and Boundary Fuel simulations with those conducted by Mell et al.
-in 2007~\cite{Mell:IJWF2007}.
+\begin{figure}[ht]
+\includegraphics[width=0.49\textwidth]{figures/Case_C064_flame_position.pdf}
+\includegraphics[width=0.49\textwidth]{figures/Case_F19_flame_position.pdf}
+\caption{Contours of heat release rate in a plane near the surface from the fine resolution (0.25 m) Particle Model compared to observations of CSIRO Grassland Fires C064 (left) and F19 (right).}
+\label{fig:CaseF19_contours}
+\end{figure}
+
+Figure \ref{fig:CaseF19_contours} shows contours of heat release rate near the surface compared to the observed flame front for the high resolution (0.25 m) Particle Model.
+These contours may be compared to the head fire spread rates at the top of Fig.~\ref{CSIRO}.
+Case F19 contours are shown on the right and may be compared to the flame images in Fig.~\ref{F19}.
+These results show some improvement to the flank fire spread rates compared to previous simulations with FDS conducted by Mell et al.~in 2007~\cite{Mell:IJWF2007}.
+
+However, it is difficult to compare the present results with those from \cite{Mell:IJWF2007}.
 The underlying numerical model, FDS, has undergone a major revision over the past decade, changing the basic finite-difference scheme, boundary conditions, drag modeling, and so on, and the vegetation-specific models have changed as well, most notably the charring reaction.
-In addition, the amount of information about the grasses and field conditions is limited, and the thermal and kinetic parameters are gleaned from various sources and are generally average values over a wide range of vegetation types.
-A sensitivity study~\cite{McGrattan:CI2017} for the particle model simulations reveals that near-surface wind speed is the has the most direct effect on the RoS, and the combined effect of the dozens of thermal, kinetic, and numerical parameters would certainly explain the difference in RoS.
+In addition, the amount of information about the grasses and field conditions is limited, and the thermal and kinetic parameters gleaned from various sources and are generally average values over a wide range of vegetation types.
+A sensitivity study~\cite{McGrattan:CI2017} for the particle model simulations reveals that near-surface wind speed has the most direct effect on the rate of spread (RoS), and the combined effect of the dozens of thermal, kinetic, and numerical parameters would certainly explain the difference in RoS between the 2007 results and those of this study.
 
@@
 This section presents an example of the methodologies discussed in the paper.
-It is notable because it considers an actual fire that occurred in the urban-wildland interface, which presents a challenge for a fire spread model because the surface vegetation is mixed with the built environment.
+It is notable because it considers an actual fire that occurred in the wildland-urban interface (WUI), which presents a challenge for a fire spread model because the surface vegetation is mixed with the built environment.
 
-At approximately 23:00 (local time) on March 25, 2019, a wildfire started in the seaside town of Cogoleto, near Genova, Italy.
-Strong winds, gusting up to 100~km/h from the North, spread the fire from its ignition point due to a faulty electric line on a ridge above the town towards the sea.
-Several houses were destroyed, hundreds of residents had to be evacuated, the firefighter crews worked for three days to tame the flames, no casualties were reported.
+At approximately 23:00 (local time) on March 25, 2019, a wildfire started in the seaside town of Cogoleto, near Genoa, Italy.
+Strong winds, gusting up to 100~km/h from the north, spread the fire from its ignition point towards the sea.
+The fire was caused by a faulty power line on a ridge above the town.
+Several houses were destroyed and hundreds of residents had to be evacuated.
+The burnt region measured over 1~km from origin to farthest point.
+The fire burned overnight and into the next day.
+Up to 70 firefighters worked at the scene, with help from helicopters and fire hydrants.
+Suppression efforts were directed towards stopping the fire spread into the coastal town of Cogoleto.
 
@@
 Shown in Fig.~\ref{Cogoleto_satellite} are the results of a level set simulation of the fire superimposed on a satellite photograph of the area and the outline of the extent of the actual fire.
-
 The computational domain is assembled using the open-source GIS (Geographic Information System) program called QGIS~\cite{QGIS}, extended by the qgis2fds~\cite{qgis2fds} plugin.
@@
 The wind speed and direction are based on a single weather station near the point of ignition.
+Winds recorded were found predominantly coming from the north with magnitudes that ranged from 20~m/s during the night diminishing to 5~m/s towards the morning hours.
 The vegetation type is provided by a database maintained at the CIMA\footnote{International Centre on Environmental Monitoring.} Research Foundation, derived from regional forest management maps.
-In the simulation, the single point wind data is taken as the prevailing wind, and the local wind field is obtained via a CFD computation.
+In the simulation, the single point wind data is taken as the prevailing wind, and the local wind field is obtained via the CFD computation.
 The fire is simulated based on the position of the level set front; that is, when the front arrives at a given location, a fire is ignited and burns for a duration of time consistent with the specified fuel loading of that particular point on the map.
 
-The extent of the simulated fire, shown in green in Fig.~\ref{Cogoleto_satellite}, is greater than the actual fire, shown in red.
-This is not surprising given that the simulation does not include fire suppression and the rate of spread of the level set front is determined empirically for vegetation that may not exactly resemble that of this particular region.
+The extent of the simulated fire, shown in green in Fig.~\ref{Cogoleto_satellite}, is greater than that of the actual fire, shown in red.
+This is not surprising given that the simulation does not include suppression and the rate of spread of the level set front is determined based on vegetation that may not exactly resemble that of this particular region.
+It is important to note that for the level set method, a fire break created by first responders can be modeled as a region declared as "non-combustible." The 13 Rothermel fuel types have been supplemented by areas such as water, pavement, and other non-combustibles.
+The fire front determined by the level set calculation is stopped by the non-combustible area, highlighting the challenge in simulating firebrand spot ignitions that were reported by first responders at the scene.
 
@@
 This points out the problem of ``validating'' wildfire models.
-The simulations of the flat terrain grassland fire experiments described in the previous section match fairly well the observed fire behavior, but in these cases, the vegetation is uniform and fairly well described in terms of its mass loading and geometrical characteristics, the terrain is flat and of relatively small area, the wind is steady, and the experiment lasts a few minutes.
+The simulations of the flat terrain grassland fire experiments described in the previous section match fairly well the observed fire behavior, but in those cases, the vegetation is uniform and fairly well described in terms of its mass loading and geometrical characteristics, the terrain is flat and of relatively small area, the wind is steady, and the experiment lasts a few minutes.
 In a real fire, none of this will be true.
-The detailed physics of the fire are not resolvable on a relatively coarse grid, the vegetation is inhomogeneous, sparse, and sometimes unknown, the weather conditions are varying, there are suppression efforts with water and the setting of backfires, and the fires can burn for days.
+The detailed physics of the fire are not resolvable on a relatively coarse grid, the vegetation is heterogeneous, sparse, and sometimes unknown, the weather conditions are varying, there are suppression efforts with water and the setting of backfires, and the fires can burn for days.
+
 All of which raise the question as to what level of physical fidelity ought to be required in these models.
-The strategy that has been adopted in FDS is to accommodate simulation options ranging from a level set calculation of fire spread over tens of kilometers that can be run in minutes, similar to models like BEHAVE developed by the U.S.
+The strategy that has been adopted in FDS is to accommodate simulation options ranging from a level set calculation of fire spread over tens of kilometers that can be run in minutes, similar to models like FARSITE developed by the U.S.
 Forest Service, all the way to detailed simulations of fire spread at sub-meter resolution for relatively short time periods (minutes to hours) and small areas (tens of hectares).
@@
 \end{document}
+
+https://www.overleaf.com/project/5f9825767d349c000144f040

### 2021-01-07 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 % Full title of the paper (Capitalized)
-\Title{Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over Complex Terrain}
+\Title{A Multi-Fidelity Framework for Wildland Fire Behavior Simulations over Complex Terrain}
 
@@
 
-There are various ways of simulating wildland fire spread depending on the resolution of the underlying numerical grid. 
+There are various ways of simulating fire in vegetation which can depend on the fire behavior of interest.
+For example, for a fire spreading through surface vegetation (e.g., grass) the fire behavior differs depending on the orientation of the local wind relative to the direction of fire spread.
+For head fires, the wind flow and fire spread are in the same direction.
+Back fires spread into the wind.
+Flank fires spread approximately orthogonal to wind.
+For a surface fire spreading from a point in a uniform vegetation, the head fire portion of the resulting fireline has the largest flame base (easiest to resolve) and is characterized by fastest spread and flames with larger lengths that tilt toward the unburned vegetation.
+Backfires have the slowest spread and smallest flame base (require finer grid cells than the head fire) and have flames with smaller lengths that tilt away from the unburned vegetation.
+The roles of radiant and convective heat transfer differ along the fireline.
+Heat gain b
 \begin{enumerate}
-\item {\bf Particle Model:} The vegetation is represented by a collection of Lagrangian particles that are heated via convection and radiation.
-This model is appropriate for grid resolution of the order of 1~m or less.
-\item {\bf Boundary Fuel Model:} Ground vegetation is modeled like a porous solid with a thickness equal to the height of the vegetation.
-This model is appropriate for grid resolution of the order of 1~m to 10~m.
-\item {\bf Level Set Method:} The fire front propagates using purely empirical rules. This model is appropriate for grid resolution of the order of roughly 10~m and greater. 
+\item {\bf Particle Model:} The vegetation (surface and raised) is represented by a collection of Lagrangian particles that are heated via convection and radiation.
+This model, with sufficient grid resolution, is appropriate for head, back, and flank surface fires, as well as fire through raised vegetation (e.g., trees).
+Heat transfer in the volume containing the vegetation is modeled in all three directions.
+This model is appropriate for grid resolutions of the order of 1~m or less, depending on the size of the flame base and properties of the vegetation (e.g.,~\cite{Perez-Ramirez:FT2017}).
+\item {\bf Boundary Fuel Model:} Surface vegetation has its own grid and is modeled like a porous solid with a thickness equal to the height of the vegetation.
+This model was designed for head fire spread in surface vegetation based on the assumption that heat transfer in the fuel bed is dominated by radiation from the overhead flame and therefore in the vertical direction.
+In the implementation here, the height of the surface vegetation is assumed to unresolved on the grid.
+The appropriate gas-phase grid resolution of the order of 1~m to 10~m.
+\item {\bf Level Set Method:} The fire front of a surface fire propagates using purely empirical rules in a level set method.
+Thermal degradation of the surface vegetation is not modeled.
+More than one implementation of this method is possible, largely differentiated by how the wind and fire-atmosphere interaction is modeled.
+The simpler implementations of this model can use grid resolutions coarser, 10 m and greater, than the more physics-based particle and boundary fuel models.
+The level set model can be used for firespread in the surface vegetation along with the particle model for fire behavior in raised vegetation.
 \end{enumerate}
@@
 
-In many simulations of wildland fire, the ground vegetation layer is too shallow to be resolved explicitly, as is done when using Lagrangian particles to represent the vegetation.
-In such cases, the ground vegetation can be modeled as a porous boundary consisting of a layer of dry vegetation, moisture, and air, underneath which is hard ground~\cite{Perez-Ramirez:FT2017}.
+In many simulations of wildland fire, the surface vegetation layer is too shallow to be resolved explicitly, as is done when using Lagrangian particles to represent the vegetation.
+In such cases, the surface vegetation can be modeled as a porous boundary consisting of a layer of dry vegetation, moisture, and air, underneath which is hard ground~\cite{Mell:IJWF2007}.
 The drag exerted by the vegetation is modeled using a special velocity boundary condition, and convective heat transfer is modeled via a source term in the one-dimensional heat conduction equation that is solved through the layer of vegetation and solid ground.
@@
 
-For simulations of wildland fires spanning large areas that cannot be gridded finely enough to predict fire spread using a physics-based model, fire spread via level sets is an option~\cite{Bova:IJWF2015}.
-This approach makes use of the Rothermel-Albini~\cite{Rothermel:1972,Albini:1976} surface fire spread rate formulae and the assumption that a surface fire spreading from a point under certain wind, slope and vegetation conditions does so with an ellipse-shaped fire front with, for a given wind speed, a fixed length-to-breadth ratio~\cite{Richards:1990}.
-The location of a fire front at time $t$ is defined by the set of points $x,y$ where the level set scalar function $\phi(x,y,t)=0$.
-The advection equation for the level set function is
+For simulations of wildland fires spanning large areas that cannot be gridded finely enough to predict fire spread using the particle or boundary fuel models for vegetation, fire spread via a level approach is used~\cite{Bova:IJWF2015}.
+It's assumed that a surface fire spreading from a point under certain wind, slope, and vegetation conditions does so with an ellipse-shaped fire front with, for an effective wind vector, a fixed length-to-breadth ratio~\cite{Finney:FARSITE,Bova:IJWF2015}.
+Determining the spread rate vector at a point on the simulated fire front requires a formula for the magnitude of head fire rate of spread which is adjusted based on the location of the point on the idealized ellipse (e.g., head versus flank fire).
+This location is determined from the orientation of the effective wind vector relative to the normal to the fire front represented by the level set.
+
+The location of a fire front at time $t$ is defined by the set of points $x,y$ where the level set scalar function $\phi(x,y,t)=0$. The advection equation for the level set function is
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{R}_{uv}=(R_u,R_v)$ is the spread rate vector. The magnitude of the vector is
+where $\mathbf{R}_{uv}=(R_u,R_v)$ is the spread rate vector.
+
+As stated above, the magnitude of the spread rate vector on a point on the level set fire front is determined by adjusting the magnitude of the head fire spread rate.
+The direction of the spread rate vector is assumed to be normal to the fire front.
+A formula for magnitude of the head fire spread rate can be obtained a number of ways, including empirical models~\cite{Cheney:IJWF1998} and physics-based simulations~\cite{Mell:FBFC2019}.
+A commonly used semi-empirical model is the Rothermel model~\cite{Rothermel:1972,Albini:1976} which can be implemented in FDS.
+Here, the magnitude of the head fire spread rate is
 \begin{equation}
@@
 See Appendix 1 of Ref~\cite{Bova:IJWF2015} for details on these calculations.
-In FDS, the level set spread model can be applied in various ways:
+The Rothermel model has been used in other CFD-based approaches including the atmospheric weighted models WRF-SFIRE~\cite{Mandel:2014} and CAWFE~\cite{Coen:2}.
+
+In FDS, the level set spread model can be implemented in the following ways:
 \begin{enumerate}
@@
 \item The wind field follows the terrain, but there is no actual fire in the simulation, just front-tracking.  The level set evolves continuously in time with the flow field.
-\item The wind and fire are fully-coupled. When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
+\item The wind and fire are fully-coupled and the resulting wind values are used in the head fire spread rate formula.
+When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
+\item The wind and fire are fully-coupled in the gas-phased, but the head fire spread rate is not influenced by the wind speed. 
 \end{enumerate}
@@
 \end{enumerate}
+The following sections provide more detail on the implementation of the boundary profiles and the pressure boundary conditions.
 
@@
 
-\subsection{Turbulent Inflow Conditions}
+% \subsection{Turbulent Inflow Conditions}
 
-It is possible in FDS to superimpose synthetic turbulence on a specified inlet profile utilizing Jarrin's synthetic eddy method (SEM) \cite{Jarrin:2008}.
+% It is possible in FDS to superimpose synthetic turbulence on a specified inlet profile utilizing Jarrin's synthetic eddy method (SEM) \cite{Jarrin:2008}.
 An example verification case is presented in Fig.~\ref{fig:semprofs}.
@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth,valign=c]{figures/sem_open_wind_contours.pdf}
-\includegraphics[width=0.48\textwidth,valign=c]{figures/sem_open_wind_u_prof.pdf}
-\caption{Example of inflow synthetic turbulence superimposed on a Monin-Obukhov boundary layer profile.
+% \begin{figure}[ht]
+% \centering
+% \includegraphics[width=0.5\textwidth,valign=c]{figures/sem_open_wind_contours.pdf}
+% \includegraphics[width=0.48\textwidth,valign=c]{figures/sem_open_wind_u_prof.pdf}
+% \caption{Example of inflow synthetic turbulence superimposed on a Monin-Obukhov boundary layer profile.
 Contours of the normal component of velocity at the domain inlet taken at a time midway through the simulation (left); mean and rms profiles along the inlet compared with specified profiles (right).}
-\label{fig:semprofs}
-\end{figure}
+% \label{fig:semprofs}
+% \end{figure}
 
-Other approaches to imposing boundary layer turbulence are possible (e.g., \cite{Munoz-Esparza:2014}).
+% Other approaches to imposing boundary layer turbulence are possible (e.g., \cite{Munoz-Esparza:2014}).
 The cases presented later in this paper do not utilize SEM since detailed data on the turbulence statistics were not available.
@@
 Moisture Fraction               & \%           & 6.3       & 5.8         \\ \hline
-Measured RoS                    & m/s          & 1.1       & 1.5         \\ \hline
+Measured RoS                    & m/s          & 1.0       & 1.5         \\ \hline
 Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s           & 1.4, 1.3, 1.4           & 1.1, 1.2, 1.2          \\ \hline
@@
 For these simulations, the rate of spread predicted by the Particle and Boundary Fuel Models are comparable to the measured rate of spread while the level set method under-predicts the RoS.
-The under-prediction by the level set simulations is a consequence of the choice of fuel model, and one should not draw any conclusions as to the performance of level sets versus the other two rate of spread methods.
+The under-prediction by the level set simulations is a consequence of the the limitations of the Rothermel model which is based on empirical relations largely derived from laboratory-scale experiments.
+It is well known that practical application of the Rothermel model often requires calibration for the case of interest~\cite{Arca_2007}.
+The use of the Rothermel for the head fire spread rate in surface vegetation in FDS is to be viewed a placeholder that is in-line with other CFD based models.
+If the observed head fire rate of spread for the F19 and C064 experiments is used in FDS, then good agreement with the observed fire perimeter is obtained (not shown).
+Further work is needed to develop rate of spread models appropriate for use with the Level Set Model (e.g.,~\cite{Mell:FBFC2019}).
 

### 2021-01-24 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@
 % Contact information of the corresponding author
-\corres{Correspondence: marcos.vanella@nist.gov}
+\corres{Correspondence: marcos.vanella@nist.gov, kevin.mcgrattan@nist.gov}
 
@@
 In this section, the wall functions used to close the mass, momentum, and energy fluxes on solid boundaries are discussed.
-Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main point of this paper).
+Mass transfer of moisture and fuel from the surface is discussed in more detail below in Sec.~\ref{sec:firespread} where the various flames spread models are compared (which is really the main focus of this paper).
 
@@
 
-There are various ways of simulating fire in vegetation which can depend on the fire behavior of interest.
-For example, for a fire spreading through surface vegetation (e.g., grass) the fire behavior differs depending on the orientation of the local wind relative to the direction of fire spread.
+There are various types of wildland fire spread models to choose from, and the choice depends on the desired level of physical fidelity and the available computing resources.
+The choice also depends on the nature of the fire.
+For example, for surface vegetation like dry grass, the fire behavior depends on the orientation of the local wind relative to the direction of fire spread.
 For head fires, the wind flow and fire spread are in the same direction.
-Back fires spread into the wind.
-Flank fires spread approximately orthogonal to wind.
-For a surface fire spreading from a point in a uniform vegetation, the head fire portion of the resulting fireline has the largest flame base (easiest to resolve) and is characterized by fastest spread and flames with larger lengths that tilt toward the unburned vegetation.
-Backfires have the slowest spread and smallest flame base (require finer grid cells than the head fire) and have flames with smaller lengths that tilt away from the unburned vegetation.
-The roles of radiant and convective heat transfer differ along the fireline.
-Heat gain b
+Back fires spread against the wind; flank fires spread laterally.
+Typically, the head fire exhibits the widest front width, most rapid spread, and longest flame lengths that tilt toward the virgin vegetation.
+Back fires exhibit the least rapid spread, most narrow front width, and shortest flame lengths that tilt away from the virgin vegetation.
+Of these three types of fire, the head fire is least demanding in terms of grid resolution; the back fire the most.
 \begin{enumerate}
@@
 This model was designed for head fire spread in surface vegetation based on the assumption that heat transfer in the fuel bed is dominated by radiation from the overhead flame and therefore in the vertical direction.
-In the implementation here, the height of the surface vegetation is assumed to unresolved on the grid.
+In the implementation here, the height of the surface vegetation is assumed to be unresolved on the grid.
 The appropriate gas-phase grid resolution of the order of 1~m to 10~m.
@@
 The simpler implementations of this model can use grid resolutions coarser, 10 m and greater, than the more physics-based particle and boundary fuel models.
-The level set model can be used for firespread in the surface vegetation along with the particle model for fire behavior in raised vegetation.
+The level set model can be used for fire spread in the surface vegetation along with the particle model for fire behavior in raised vegetation.
 \end{enumerate}
@@
 When the fire front arrives at a given surface cell, it burns for a finite duration and with a heat release per unit area provided as part of the fuel model.
-\item The wind and fire are fully-coupled in the gas-phased, but the head fire spread rate is not influenced by the wind speed. 
+\item The wind and fire are fully-coupled in the gas-phase, but the head fire spread rate is not influenced by the wind speed. 
 \end{enumerate}
@@
 Moisture Fraction               & \%           & 6.3       & 5.8         \\ \hline
-Measured RoS                    & m/s          & 1.0       & 1.5         \\ \hline
-Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s           & 1.4, 1.3, 1.4           & 1.1, 1.2, 1.2          \\ \hline
-Calc'd RoS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) & m/s      & 1.5, 1.4, 1.7           & 1.3, 1.3, 1.3           \\ \hline
-Calc'd RoS, Level Set Method (5 m, 10 m, 20 m resolution) & m/s          & 1.1, 1.1, 1.2          & 0.4, 0.5, 0.8          \\ \hline
+Measured RoS                    & m/s          & 1.2       & 1.5         \\ \hline
+Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) 
+                                & m/s          & 1.1, 1.2, 1.2   & 1.4, 1.3, 1.4          \\ \hline
+Calc'd RoS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) 
+                                & m/s          & 1.3, 1.3, 1.3   & 1.5, 1.4, 1.7          \\ \hline
+Calc'd RoS, Level Set Method (5 m, 10 m, 20 m resolution) 
+                                & m/s          & 0.5, 0.6, 0.6   & 1.0, 1.1, 1.2          \\ \hline
 \end{tabular}

### 2021-02-05 Review_Process/Response_to_Reviewer_1_Rev_1.tex
Update on Overleaf.

\documentclass[12pt]{article}

\setlength{\textwidth}{6.25in}
\setlength{\textheight}{8.75in}
\setlength{\topmargin}{0.in}
\setlength{\headheight}{0.in}
\setlength{\headsep}{0.in}
\setlength{\parindent}{0.in}
\setlength{\oddsidemargin}{0.0in}
\setlength{\evensidemargin}{0.0in}
\setlength{\parskip}{\baselineskip}
\setlength{\leftmargini}{\parindent} % Controls the indenting of the "bullets" in a list

\usepackage{xcolor}
\newcommand\hl[1]{\textcolor{red}{#1}}

\pagestyle{empty}

\begin{document}

\begin{center}
{\bf Response to comments by Reviewer of Revision 1}
\end{center}

Compared to the first review, the authors accounted for many raised comments. In particular, the used models are now well described. However, even the authors agree that this work allows drawing a clear conclusion, in particular on the section 6.1 on the observed poor relevance of the level set method.

\hl{How well the level set based predictions agree with an observed fire perimeter evolution depends mostly on the value of head fire spread rate model. The head fire spread rate can be obtained in a number of ways, such as empirical models, semi-empirical models, or simulations the are more physics-based. The simulations presented in this paper use the Rothermel model to obtain the head fire spread rate. This semi-empirical model is based mostly on measurements from laboratory-scale simulations. It's well known that, for many applications, the Rothermel model requires calibration (which we did not do). At this point in its development, FDS uses the Rothermel model as a placeholder for obtaining the head fire spread rate. This is in-line with other CFD based approaches that use a front-tracking approach. For this reason, it is not expected that the level set simulations presented here will agree well with observations of Australian grass fires. We have added this discussion and references to other CFD approaches that use the Rothermel model and a reference that provides an example of the need to calibrate the Rothermel model.}

Indeed, two main elements (among others) could explain the observed discrepancies with the more realistic solid particles or the boundary fuel method: the coarse discretization or the imposed (and Indeed difficult to calibrate) RoS value. Without providing more hints on this problem and complementary simulations to discriminate between these two effects, in my opinion this paper could hardly be published. I fully understand that the author's aim is difficult and corresponds to an open question, so I have a couple of suggestions:

- Has the level set method a strong dependence w.r.t. the grid cell size, including cell sizes tending towards the ones used in the two first methods ?

\hl{Sensitivity to the grid resolution is presented in Table 1. The level set model, when implemented with the head fire spread rate being dependent on the wind speed, can depend on the grid resolution because the wind can depend on the grid resolution. The extent of this dependence is something that a model user needs to assess through grid sensitivity tests. However, the under prediction of the head fire spread rate seen here are due to, as discussed above, error in the Rothermel model.}

- Is it possible to use  the well refined simulations to chacterize a RoS and the include it in the level set approach to see if the results are more relevant ?

\hl{If by ``well refined simulations" the reviewer is referring to using more physics-based simulations to determine a RoS for use in the level set approach then, yes, this is possible and work is being conducted along these lines. Text has been added to bring up this point.}

Another (less important) point regards the SEM: why presenting it without using it for comparison ?

\hl{OK, the SEM discussion has been removed.}

\end{document}

### 2021-02-05 Review_Process/Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

--- 2021-01-24 Vanella_Article_Atmosphere_2020.tex
@@
 %=================================================================
-\documentclass[journal,article,atmosphere,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[journal,article,atmosphere,submit,moreauthors,pdftex]{Definitions_Review_Process/mdpi}
 
@@
    \centering
-   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{./figures/SketchFig1.pdf}
+   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{../figures/SketchFig1.pdf}
    \put(-305,-5){(a)}
@@
    \centering
-   \includegraphics[trim = 30mm 42mm 35mm 45mm, clip,width=1\linewidth]{./figures/SketchFig2.pdf}
+   \includegraphics[trim = 30mm 42mm 35mm 45mm, clip,width=1\linewidth]{../figures/SketchFig2.pdf}
    \put(-335,-5){(a)}

### 2021-02-05 atmosphere-1087074-final.tex
Update on Overleaf.

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[atmosphere,article,accept,moreauthors,pdftex]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, animals, antibiotics, antibodies, antioxidants, applmech, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejbc, ejihpe, electrochem, electronics, endocrines, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, oceans, ohbm, optics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, sustainableworld, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1} 
\makeatletter 
\setcounter{page}{\@firstpage} 
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{1}
\pubyear{2021}
\copyrightyear{2021}
\externaleditor{Academic Editor: Nicholas Herold}
\datereceived{} 
\dateaccepted{} 
\datepublished{} 
\hreflink{doi:} % Please use \linebreak if need to do line break
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

\usepackage[export]{adjustbox}
\usepackage{url}

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{A Multi-Fidelity Framework for Wildland Fire Behavior Simulations over Complex Terrain}
%Attention AE/ME. The following layout issues have not been checked by the English Editing Department and must be carefully verified by the AE/Layout Department: All callout issues, bold usage of callouts, and references to callouts in the text. Correct callout usage in figures. Figure and Table layout issues. Footnote formatting and Glossaries have not been checked. En dash usage for negative values, en dash usage to indicate relationships, en dash usage to indicate bonds (especially in chemistry). The English Editing Department is not responsible for correct italic usage for genes, proteins and technical terminology. This responsibility belongs to the authors. The following are also not checked: spacing between numbers and units of measurement, ratios, en dashes for ranges, date and time formats, punctuation in equation lines, and less than/more than spacing (< >). Finally, capitalization and layout of titles/headings must be properly checked as well as ensuring 'Eq.' and 'Fig.' are properly spelled out, as these are layout issues.

\TitleCitation{A Multi-Fidelity Framework for Wildland Fire Behavior Simulations over Complex Terrain}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Marcos Vanella$^{1,}$*, Kevin McGrattan $^{1,}$*, Randall McDermott $^1$, Glenn Forney $^1$, William Mell $^2$, Emanuele Gissi $^3$, and Paolo Fiorucci $^4$}

% Authors, for metadata in PDF
\AuthorNames{Marcos Vanella, Kevin McGrattan, Randall McDermott, Glenn Forney, William Mell, Emanuele Gissi, and Paolo Fiorucci}
\AuthorCitation{Vanella, M.; McGrattan, K.; McDermott, R.; Forney, G.; Mell, W.; Gissi, E.; Fiorucci, P.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad National Institute of Standards and Technology, Gaithersburg, MD 20899, USA; kevin.mcgrattan@nist.gov (K.M.); randall.mcdermott@nist.gov (R.M.); glenn.forney@nist.gov (G.F.) \\
$^{2}$ \quad U.S. Forest Service, Seattle, Washington, WA 98103, USA; william.mell@usda.gov \\
$^{3}$ \quad Corpo nazionale dei Vigili del fuoco, Savona 17100, Italy; emanuele.gissi@vigilfuoco.it \\
$^{4}$ \quad CIMA Research Foundation, Savona 17100, Italy; paolo.fiorucci@cimafoundation.org}

% Contact information of the corresponding author
\corres{Correspondence: marcos.vanella@nist.gov, kevin.mcgrattan@nist.gov}

\abstract{A method for the large-eddy simulation (LES) of wildfire spread over complex terrain is presented. In this scheme, a cut-cell immersed boundary method (CC-IBM) is used to render the complex terrain, defined by a tessellation, on a rectilinear Cartesian grid. Discretization of scalar transport equations for chemical species is done via a finite volume scheme on cut-cells defined by the intersection of the terrain geometry and the Cartesian cells. Momentum transport and heat transfer close to the immersed terrain are handled using dynamic wall models and a direct forcing immersed boundary method. A new ``open'' convective inflow/outflow method for specifying atmospheric wind boundary conditions is presented. Additionally, three basic approaches have been explored to model fire spread: (1) Representing the vegetation as a collection of Lagrangian particles, (2) representing the vegetation as a semi-porous boundary, and (3) representing the fire spread using a level set method, in which the fire spreads as a function of terrain slope, vegetation type, and wind speed. Several test and validation cases are reported to demonstrate the capabilities of this novel wildfire simulation methodology.}

\keyword{complex terrain; fire spread; immersed boundary method; level sets}

\newcommand{\dV}{{\rm d}V}
\newcommand{\dS}{{\rm d}S}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bs}{\mathbf{s}}
\renewcommand{\C}{\mathrm{C}}
\renewcommand{\H}{\mathrm{H}}
\renewcommand{\O}{\mathrm{O}}

\begin{document}

\section{Introduction}

In the past few decades, wildland and wildland-urban interface (WUI) fires have received substantial attention due to their destructive nature and extensive cost to lives and property~\cite{thomas_2017,mcdermott_2019,richards_2020}. A~large body of literature can be found on the subject, as~seen in various review articles~\cite{Papadopoulos_2011,Bakhshaii_2019,mcdermott_2019}. Due to the complex nature of fire, limited computing resources, and~the needs of planners and first responders, most models of wildfires have historically relied on simplified field-tested rules and correlations. Among~these, the~rate of the spread model of Rothermel~\cite{Rothermel:1972} assumes quasi-steady-state surface fire conditions, and takes, as input, fuel properties, wind, terrain slope, and~moisture that can be measured in~situ. Fuel models have been cataloged~\cite{Anderson:1982}, and~their parameters defined so that model users can select and combine the most appropriate listed fuels. Rothermel's rate of spread formula is still widely used, and~has been implemented in several simulation tools~\cite{Finney:FARSITE,Bova:IJWF2015,FDS_Users_Guide,Coen:2,Coen:2015,LAUTENBERGER_2013,coen_2013,Mandel:2009,Mandel:2011,Mandel:2014,Kochanski:2016}, enabling a range of planning and operational forecasting~capabilities.

Wildland fire modeling typically focuses on tracking the fire front, which can have a complex shape that moves and deforms based on local conditions, like fuel loading, moisture content, wind, and~terrain. The~fire front requires a mathematical representation. The~Lagrangian approach is based on a polygonal mesh of control markers that represents the front~\cite{Finney:FARSITE,Bova:IJWF2015}. On~the other hand, an~Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field. This scalar field is called the level set function, and numerical methods that compute the function are called level set methods (LSM)~\cite{Sethian:1999,Osher:2006}. Some wildfire solvers that use LSM to track the fire front are described in the References~\cite{coen_2013,Bova:IJWF2015,FDS_Users_Guide,LAUTENBERGER_2013}. A~practical comparison of the two techniques, as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE}, is provided in Ref.~\cite{Bova:IJWF2015}.

Because the spread rate of the fire front depends in part on the local wind conditions, a~well-resolved wind field over complex terrain should improve the accuracy of the overall model. The~de facto physics-based technique for practical simulation of wind is the Large Eddy Simulation (LES)~\cite{Linn:2007,coen_2013,Mell:IJWF2007}. For~fire simulation over complex terrain, a~range of techniques are being developed, in~particular, in~the coupling of the fire and the wind. For~fast simulation of outdoor fires, an~immersed boundary method has been interfaced with a level set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model. The~wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution. These are, in turn, also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept. This is a first step at implementing a level set approach in an LES wind calculation with combustion. It is recognized that in the empirically based development of the Rothermel formula, the wind speed is not influenced by the local buoyancy-induced~flow.

A number of other approaches to modeling fire spread over complex terrain exist. A~recent review, with~an emphasis on operational applications, is given in~\cite{Arca_2019}. A~range of approximations to modeling both wind and fire have been employed. The~uniqueness of the approach presented here is that, within~a single computational tool, a~range of approximations to the wind and fire physics can be employed. This can support a direct comparison between models of varying physical fidelity. For~example, one can investigate the difference in predicted fire behavior between a simulation that models fire spread and heat release via the explicit accounting of the thermal degradation of vegetation and one that approximates fire spread via an empirically based~model.

In the following section, the mathematical model for fire spread over complex terrain is described. The~description of the terrain discretization is provided in Section~\ref{sec:terraindisc}. The~various fire spread methods are described in Section~\ref{sec:firespread}, and~the atmospheric boundary conditions in Section~\ref{sec:wind}. Some flat terrain simulations are compared to experimental data, and~a complex terrain simulation is compared to an actual wildfire in Section~\ref{sec:numexp}.

\section{Mathematical~Model} \label{sec:matmodel}
\unskip

\subsection{Governing~Equations} \label{sec:goveqns}

For fire modeling in the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}. Under~this assumption, the pressure field $p$ can be viewed as the summation of two components: a background hydrostatic pressure field $\bar{p}(z,t)$ used in the equation of state (ideal gas law), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ driving fluid~motion.

Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$. If~$\rho(\mathbf{x},t)$ is the mixture density, the~mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, where $\rho_\alpha(\mathbf{x},t)$ is the species mass density and $\rho = \sum \rho_\alpha$. The~scalar transport and momentum equations take the form:
\begin{eqnarray}
   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \rho Y_\alpha  \mathbf{u} ) &=& - \nabla \cdot \mathbf{J_{d}}_\alpha + \dot{m}_\alpha'''  +
    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau}^{dev} \right] \label{eqn:momtran},
\end{eqnarray}
where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ is the diffusive flux for component $\alpha$. Fick's Law for binary diffusion with respect to a background species is assumed, and~$D_\alpha$ is the diffusivity of $\alpha$ with respect to the background species.  Mass fractions $Y_\alpha$, solution of equations~\eqref{eqn:spectran}, must obey realizability constraints $0<Y_\alpha<1$ and $\sum_\alpha Y_\alpha=1$. In~equations~\eqref{eqn:spectran} starting from a realizable solution, realizability implies $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$. Therefore, to~enforce realizability, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}. The~volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$;  the term $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification. These correspond to the amount of mass per unit volume and time of species $\alpha$ being added or subtracted in a given point $\mathbf{x}$ due to chemical reaction or gasification of solid particles (or droplets), respectively.  Details of the combustion model are provided below in Section~\ref{sec:subgrid}.

In the momentum equation, the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and~$\rho_0(z)$ is the modeled height-dependent background density of the atmosphere.  \mbox{Additionally, $\boldsymbol{\omega}$ refers} to the vorticity field. The~gravity vector is $\mathbf{g}=(0,0,g_z)$, the~buoyancy force term is $ (\rho-\rho_0) \mathbf{g}$, $\mathbf{f}_{b}$ is the term contributed by modeled particle drag forces, and~$\boldsymbol{\tau}^{dev}$ is the deviatoric stress tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model, discussed below in Section~\ref{sec:subgrid}.

Conservation of energy is achieved by forcing the flow field to obey the following thermodynamic divergence constraint:
\begin{eqnarray}
    ( \nabla \cdot \mathbf{u} )^{\rm th} &=&
    \left[ \frac{1}{\rho c_p T} - \frac{1}{\bar{p}} \right]
    \frac{\partial \bar{p}}{\partial t} + \frac{w \rho_0 g_z}{\rho c_p T} \nonumber \\
    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \nabla \cdot \dot{\mathbf{q}}_{\rm r}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
    &+& \frac{1}{\rho} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right) \left[ \dot{m}_\alpha'''  + \dot{m}_{b,\alpha}''' - \nabla \cdot (\mathbf{J_{d}}_\alpha) - \mathbf{u} \cdot \nabla (\rho Y_\alpha) \right] \label{eq:divth},
\end{eqnarray}
where $h_{s,\alpha}$ is the sensible enthalpy of species $\alpha$, and $h_s$ is the sensible enthalpy of the mixture, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $T$ is the local gas temperature, and~$w$ is the local vertical velocity. The~mixture-specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$, respectively. This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}. Equation~(\ref{eq:divth}) is derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law.  Given a background pressure and a local mass density, the~local temperature is computed from the ideal gas law as $T = \bar{p} \overline{W} / (\rho R) $.

The term $\dot{\mathbf{q}}''$ in Equation~(\ref{eq:divth}) represents the conductive and diffusive heat fluxes discussed in Section~\ref{sec:subgrid}. The~net contribution from thermal radiation in the energy equation is defined by:
\begin{equation}
   -\nabla \cdot \dot{\mathbf{q}}_{\rm r}''(\bx) =
    \kappa(\bx) \, \left[ U(\bx) - 4 \pi \, I_{\rm b}(\bx) \right]  \quad ; \quad
    U(\bx) = \int_{4\pi} \, I(\bx,\bs') \, d\bs'  \label{simple_rte},
\end{equation}
where $\kappa(\bx)$ is the absorption coefficient, $I_b(\bx)$ is the source term, and~$I(\bx,\bs)$ is the solution of the radiation transport equation (RTE) for a non-scattering gray gas:
\begin{equation}
   \bs \cdot \nabla I(\bx,\bs) = \kappa(\bx) \; \left[ I_{\rm b}(\bx) - I(\bx,\bs) \right] \label{bandRTE1}.
\end{equation}

The source term, $I_{\rm b}$, requires special treatment because of the limited resolution of the underlying numerical grid in the vicinity of flames. In~large-scale fire simulations, \mbox{grid cells} are typically on the order of
tens of centimeters. Flame sheets cannot be resolved, meaning that the computed cell-average temperature can be significantly lower than temperatures one would expect to find in the reacting flame. Consequently, the~source term is approximated in grid cells where fuel and oxygen react. Elsewhere, the~subgrid temperature field is homogeneous, and the source term can be computed directly:
\begin{equation} \kappa I_{\rm b} = \left\{ \begin{array}{cll}
    \kappa \sigma T^4/\pi      & \hbox{Outside flame zone}, & \dot{q}'''=0  \\ [0.1in]
    C \kappa \sigma T^4/\pi  & \hbox{Inside flame zone}, & \dot{q}'''>0.
    \end{array} \right.  \label{radapprox1}
\end{equation}

The constant $C$ is computed at each time step so that the volume integral of \mbox{Equation~(\ref{simple_rte})} over the entire flaming region is approximately equal to the volume integral of $\chi_{\rm r} \,\dot{q}'''$ over that same region. Here, $\chi_{\rm r}$ is an empirical estimate of the {\em global} fraction of that energy emitted as thermal radiation. Typically, a~sooty fire radiates approximately one-third of the total combustion~energy.

The radiation equation is solved using a technique similar to a finite volume method for convective transport, thus the name given to it is the Finite Volume Method (FVM). Using approximately 100 discrete angles which are updated over multiple time-steps, the~finite volume solver requires about 20\% of the total CPU time of a calculation, a~modest cost given the complexity of radiation heat~transfer.

Section~\ref{particle_model} discusses how Lagrangian particles are used to represent vegetation like leaves, pine needles, and other subgrid objects. These particles absorb and emit thermal radiation, and~the underlying assumptions are described in that~section.

\subsection{Subgrid~Parameterizations} \label{sec:subgrid}

In the formal derivation of the LES equations (see, e.g,~\cite{Pope:2000,FDS_MathVal_Guides}) the filtered nonlinear terms (advection terms, mean chemical source term, and~the unresolved boundary fluxes) require subgrid-scale parameterizations.  In~this section, LES filter notation, as well as Cartesian tensor index notation, are used to clarify the terms used in the governing~equations.

\subsubsection{Subgrid~Advection}

Unresolved turbulent eddies enhance the transport of mass, momentum, and~energy.  This added transport is accounted for with residual stresses and scalar fluxes.  The~subgrid-scale stress tensor is defined as
\begin{equation}
\tau_{ij}^{sgs} \equiv \bar{\rho}(\widetilde{u_i u_j} - \tilde{u}_i \tilde{u}_j),
\end{equation}
where an overline represents a volumetric filter and the tilde represents a Favre filter (\mbox{see, e.g.,~\cite{Poinsot:TNC}}).

In FDS, the deviatoric part of the subgrid stress tensor is modeled using the gradient diffusion hypothesis.  Thus, the~combined deviatoric stress may be written as follows:
\begin{equation}
\tau_{ij}^{dev} = \overline{\tau}_{ij} + \tau_{ij}^{sgs} - \frac{1}{3} \tau_{kk}^{sgs} = -2 (\mu + \mu_t) \left[ \tilde{S}_{ij} - \frac{1}{3} \tilde{S}_{kk} \right],
\end{equation}
where the symmetric rate-of-strain tensor is $S_{ij} \equiv \frac{1}{2}(\partial u_i/\partial x_j + \partial u_j/\partial x_i)$, and $\mu_t$ is an isotropic eddy viscosity.  Note that $\overline{\tau}_{ij}$ is already the deviatoric part of the molecular stress as the isotropic part is the pressure; $\mu$ is the molecular dynamic~viscosity.

The eddy viscosity $\mu_t$ is an important quantity in the simulation, as it affects all the turbulent transport and subgrid mixing time-scales, which, as~shown below, are crucial to the combustion model.  In~FDS, the~eddy viscosity is modeled as~\cite{Pope:2000,Deardorff:1980}
\begin{equation}
\mu_t = \bar{\rho} C_\nu \Delta \sqrt{k_{sgs}},
\end{equation}
with the constant taken as $C_\nu = 0.1$ (this value can be derived, assuming production equals dissipation and a model Kolmogorov spectrum~\cite{Pope:2000}; the model is also tested against decaying isotropic turbulence~\cite{CBC}).  In~practice, the~filter scale $\Delta$ is taken as the cube root of the cell volume.  Deardorff~\cite{Deardorff:1980} solved a transport equation for the subgrid kinetic energy per unit mass, $k_{sgs}$, in~order to account for subgrid buoyancy. In~FDS, the~density field is sufficiently resolved to account for buoyant plume accelerations, and~therefore it is found that an algebraic closure for $k_{sgs}$ is sufficient.  Following the work of Bardina~\cite{Bardina:1980}, a~scale-similarity idea is used; the subgrid kinetic energy per unit mass is modeled \mbox{as follows}:
\begin{equation}
k_{sgs} = \frac{1}{2}(\tilde{u}_i - \hat{\tilde{u}}_i)(\tilde{u}_i - \hat{\tilde{u}}_i),
\end{equation}
where the hat represents a test filter at a scale $2\Delta$.  For~the first Cartesian cell off the wall, where a test filter is not well-defined, the~eddy viscosity is taken from the WALE model~\cite{Nicoud:1999}.

Subgrid advection of mass and heat are also modeled using gradient diffusion.  The~turbulent diffusivities use constant turbulent Schmidt and Prandtl numbers, both set to 0.5; this simplification is justified based on high-resolution simulations of fire plumes (\mbox{see, e.g.,~\cite{Maragkos:2020}}). For~mass diffusion of species $\alpha$ in direction $i$, the flux is
\begin{equation}
J_{\alpha,i} = \bar{J}_{\alpha,i} + J_{\alpha,i}^{sgs} = -\bar{\rho}(D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t}) \frac{\partial \tilde{Y}_\alpha}{\partial x_i}.
\end{equation}
Effective thermal conductivity and mass diffusivity are used in the heat flux, which may be written as
\begin{equation}
{\dot{q}''}_i = -(\bar{k} + \frac{\mu_t \bar{c}_p}{\mathrm{Pr}_t}) \frac{\partial \widetilde{T}}{\partial x_i} - \sum_\alpha \tilde{h}_{{\rm s},\alpha} ( \bar{\rho} D_\alpha + \frac{\mu_t}{\mathrm{Sc}_t} ) \frac{\partial \tilde{Y}_\alpha }{\partial x_i}.
\end{equation}

\subsubsection{Filtered Chemical Source~Term}

FDS treats combustion chemistry using a simplified approach.  Only fuel, air, and~products, referred to as ``lumped species'', are tracked.  Lumped species are groups of species that transport and react together and are always found in the same proportion.  For~example, air is a lumped species made up of 23 \% O$_2$ and 77 \% N$_2$ by mass.  For~vegetation, a~surrogate hydrocarbon, fuel, is usually assumed.  For~example, cellulose may be assumed to decompose to a volatile fuel gas with the formula $\C_{6} \H_{10} \O_{5}$.  The~products are defined stoichiometrically, often including prescribed yields of soot (for smoke) and carbon monoxide that may be obtained from empirical relationships for the specific fuel or from other sources (e.g., \cite{SFPE:Mulholland}).  Then, usually, a~single reaction is tracked (more complicated reaction schemes are possible in the code, but~for this work, a single reaction is sufficient):
\begin{equation}
\mathrm{Fuel} + s \, \mathrm{Air} \rightarrow (1+s) \,\mathrm{Products},
\end{equation}
where $s$ is the mass stoichiometric~coefficient.

In LES of turbulent combustion, the filtered chemical source term, $\overline{\dot{m}_\alpha'''}$, is unclosed because the reaction kinetics are nonlinear functions of species concentration and temperature.  However, since the heat-releasing reactions in a fire are extremely fast compared to the time scales for turbulent mixing, it is common practice in fire dynamics simulations to utilize the eddy dissipation model of Magnussen and Hjertager~\cite{Magnussen:1} (see also~\cite{Poinsot:TNC}).  The~mass production term for the fuel species is written as
\begin{equation}
\overline{\dot{m}_{F}'''} = -\bar{\rho} \,\frac{\min(\tilde{Y}_F,\tilde{Y}_A/s)}{\tau_{mix}}.
\end{equation}

This formula states that the fuel is consumed at a rate proportional to the local limiting reactant concentration, and inversely proportional to the local mixing time-scale.  This is the so-called ``mixed is burnt'' approximation.  The~mixing time-scale is given by $\tau_{mix} = \min(\tau_d, \tau_u,\tau_g)$ \cite{McDermott:2011}, where the time-scales for diffusion, turbulent advection, and~gravitational acceleration are, respectively,
\begin{align}
\tau_d &= \Delta^2/D_F \\
\tau_u &= C_u \Delta / \sqrt{(2/3)k_{sgs}} \\
\tau_g &= \sqrt{2\Delta/g}.
\end{align}

Here, $D_F$ is the binary diffusivity of the fuel species in air, $k_{sgs}$ is the modeled subgrid kinetic energy per unit mass discussed above, and~$g$ is the gravitational acceleration. The~constant $C_u=0.4$ is determined by matching flame height correlations across a wide range of fire Froude numbers~\cite{FDS_MathVal_Guides}.

The remaining mass production rates are found from stoichiometry.  The~local heat release rate, $\dot{q}'''$, an~important term in Equation~(\ref{eq:divth}), is then determined from
\begin{equation}
\label{eq:qdotppp}
\dot{q}''' = - \sum_\alpha \overline{\dot{m}_\alpha'''} \Delta h_{f,\alpha},
\end{equation}
where $\Delta h_{f,\alpha}$ are the heats of formation for each lumped species $\alpha$.

\paragraph{Flame Extinction} One of the advantages of a physics-based tool like FDS is the possibility of building into the model the effects of defensive actions on the fire spread rate.  FDS uses simple empirical rules to predict local extinction within a gas phase grid cell based on resolved species concentrations and the mean cell temperature. The~rules are based on the concept of a critical flame temperature. The~basic theory behind the critical flame temperature is described in~\cite{SFPE:Beyler}.  In~brief, if~the local heat release given by Equation~(\ref{eq:qdotppp}) is not sufficient to raise the cell temperature above the CFT, then $\overline{\dot{m}_{F}'''}$ is set to zero for that cell for that time-step.  Note that the flame extinction model does not directly affect surface cooling or smothering of the vegetative fuel, which may happen with certain fire suppression activities. However, such tactics can be accounted for through modifications to the solid fuel thermal degradation model discussed~below.

\subsubsection{Unresolved Boundary~Fluxes}
\label{sec:boundflx}
In this section, the~wall functions used to close the mass, momentum, and~energy fluxes on solid boundaries are discussed.  Mass transfer of moisture and fuel from the surface is discussed in more detail below in Section~\ref{sec:firespread}, where the various flames spread models are compared (which is really the main focus of this paper).

The scaled streamwise velocity is denoted $u^+ = u/u_\tau$, where the friction velocity is $u_\tau \equiv \sqrt{\tau_w/\rho}$ and $\tau_w$ is the wall stress needed to close the momentum equation, \mbox{Equation~(\ref{eqn:momtran})}.  For~rough walls, FDS employs the log law presented in~\cite{Pope:2000},
\begin{equation}
\label{eqn_roughwallloglaw}
u^+ = \frac{1}{\kappa} \ln \left(\frac{y}{s}\right) + \tilde{B}(s^+),
\end{equation}
where $\kappa = 0.41$ is the von K\'arm\'an constant, $s^+ = s/\delta_\nu$ is the roughness length in viscous units, $s$ is the dimensional ``sand grain'' roughness, and~$\delta_\nu = \mu/(\rho u_\tau)$ is the viscous length scale. The~distance to the wall, $y$, is taken as $\delta y/2$ for the first off-wall grid cell.  The~parameter $\tilde{B}$ varies with $s^+$ but attains a constant value $B_2=8.5$ in the fully rough limit.  In~FDS, $\tilde{B}$ is implemented as the following piece-wise function:
\begin{equation}
\tilde{B} = \left\{ \begin{array}{lll} B + (1/\kappa)\ln(s^+) & \mbox{for} & s^+ < 5.83 \\
\tilde{B}_{\rm max} & \mbox{for} & 5.83 \le s^+ < 30.0 \\
B_2 & \mbox{for} & s^+ \ge 30.0 \end{array} \right.,
\end{equation}
where $\tilde{B}_{\rm max} = 9.5$.

In the fully rough limit, the sand grain roughness may be equated to the aerodynamic roughness, $z_0$, typically employed in atmospheric codes for a neutral boundary layer,
\begin{equation}
s = z_0 e^{8.5 \kappa} \approx 32.6 \, z_0.
\end{equation}

It is important to appreciate that FDS was originally designed to handle compartment fires and must consider flows with smooth walls, such as HVAC ducts.  Further, FDS typically resolves the thermal motions that would necessitate the use of Monin–Obukhov stability corrections to the boundary~profile.

\textls[-20]{The unresolved heat flux at the surface, $\dot{q}_w''$, is determined using an \mbox{engineering approach}},
\begin{equation}
\dot{q}_w'' = h (T_g - T_w),
\end{equation}
where $h$ is the convective heat transfer coefficient, $T_w$ is the local temperature of the surface, and~$T_g$ is the temperature of the first gas phase cell off the wall.  Empirical natural/forced convection correlations are used to determine $h$:
\begin{equation}
h = \max \left[C\, |T_g-T_w|^{1/3} \; , \;
\frac{k}{L} \, \mathrm{Nu} \; , \;
\frac{k}{\delta n/2} \right],
\label{eq:qconv}
\end{equation}
where $C$ is a empirical coefficient for natural convection (1.52 for a horizontal plate and 1.31 for a vertical plane or cylinder)~\cite{Holman:1}, $L$ is a characteristic length related to the size of the physical obstruction, and~$k$ is the thermal conductivity of the gas. The~forced convection Nusselt number (Nu) depends on the geometry and flow characteristics~\cite{Holman:1,Incropera:1}.  Further details are provided in~\cite{FDS_MathVal_Guides}.

Again, this simple approach is viable because FDS usually resolves the thermal flow fields that would necessitate added complexity.  However, future work is planned to explore improvements to wall functions that may lead to better results at coarser grid~resolution.

\section{Terrain Description and~Discretization} \label{sec:terraindisc}

In Figure~\ref{Fig:figure_1}, the~terrain is represented by its surface triangulation within an FDS Cartesian mesh. Some Cartesian cells and faces belonging to this mesh are transversed by the terrain surface. The~remaining polyhedra and polygons that lie on the gas side of these intersected geometries are called cut-cells and cut-faces. They discretize the fluid domain in the region surrounding the solid surface. A~detail of this unstructured mesh is shown in Figure~\ref{Fig:figure_1}. The~resulting cut-cells and cut-faces are of various sizes and shapes, and~situations where there are more than one cut-cell per Cartesian cell are common. As~described later, small cells impose a temporal stability constraint on explicit time integration schemes. A~two-level grid refinement hierarchy emerges. The~coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured components. Methods that solve discrete model equations on these grids are called cut-cell or embedded boundary methods~\cite{berger_2016}. Reliably defining the cut-cell mesh geometric properties and topology is a complex problem in and of itself, but~beyond the scope of this~paper.
\begin{figure}[H]
   %\centering
   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{./figures/SketchFig1.pdf}
   \put(-305,-5){(\textbf{a})}
   \put(-95,-4){(\textbf{b})}
   \caption{Sketch of regular Cartesian and cut-cell grids around terrain: (\textbf{a}) An unstructured cut-cell region is defined on the gas phase side around a terrain $\mathbf{T}$. (\textbf{b}) Detail showing a slice representative of cut-cells and regular gas cells, terrain triangulation, and two level refinement interpretations.}
   \label{Fig:figure_1}
\end{figure}
\unskip

\subsection{Cut-Cell Scalar Transport and Energy~Discretization}

The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}, as~outlined in the diagram of Figure~\ref{Fig:figure_2}. Integrating Equation~(\ref{eqn:spectran}) over a cut-cell $ii$ with control volume $\Omega_{ii}$ gives:
\begin{equation}
 \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
      } \, \dV  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \mathbf{J_{d}}_\alpha   } \, \dV + \int_{\Omega_{ii}} { \left( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''' \right) } \, \dV \label{eq:intconvdiff}.
\end{equation}

For a cut-cell control volume, the time derivative and source terms are \mbox{approximated by}
\begin{equation}
  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \overline{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\overline{ \dot{m}_\alpha''' }_{ii}+ \overline{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons},
\end{equation}
where $V_{ii}$ is the volume of cell $ii$ and the overlines imply cell averages. In~the following, the notation is simplified by dropping overlines, observing that quantities are cell- or face-averaged. Consider the FV discretization of the diffusive term of Equation~(\ref{eq:intconvdiff}) on cut-cell $ii$ of Figure~\ref{Fig:figure_2}:
\begin{equation}
\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot   \mathbf{J_{d}}_\alpha   } \, \dV =
    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \, \dS = \sum^{n_{\rm f}}_{k=1}
    \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvdiffcc}.
\end{equation}

The integral over the cut-cell volume has been transformed in an area integral on its $n_{\rm f}=5$ faces using the divergence theorem. As~these $k$-faces with areas $A_k$ and outward normals $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the~method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization. Centroid to centroid (i.e., $\Delta x_{\rm cc}$ in Figure~\ref{Fig:figure_2}) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase. Additionally, a~normal probe approach~\cite{balaras_2004}, is employed to sample information from the fluid to define fluxes in boundary cut-faces. In~a normal probe method, an~external point is defined at a normal distance from the body of the order of the cartesian grid size. Then, information from the fluid at this external point is obtained via~interpolation. 
\begin{figure}[H]
   %\centering
   \includegraphics[trim = 30mm 42mm 35mm 45mm, clip,width=0.98\linewidth]{./figures/SketchFig2.pdf}
   \put(-335,-5){(\textbf{a})}
   \put(-95,-4){(\textbf{b})}
   \caption{Sketch of cut-cell and faces: (\textbf{a}) Cut-cell $ii$ surrounded by gas phase regular and cut-faces (3--5), and~boundary cut-faces (1--2).  (\textbf{b}) Interpolation sketch for wall-modeled, immersed boundary reconstruction of cut-face~velocities.}
   \label{Fig:figure_2}
\end{figure}
Similarly, the~discretization of the advective term is:
\begin{equation}
    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} \, \dV =
     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \, \dS \approx
     \sum^{n_{\rm f}}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \, A_k \label{eq:discfvadvcc},
\end{equation}
where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and~the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{FDS_Users_Guide}. In~the cut-cell region, a Godunov flux limited interpolation for the advective term is used. Flux-limited interpolation is one of the fundamental components of stable numerical schemes for hyperbolic equations~\cite{leveque_2002}, of~which mass transport inherits its mathematical properties.
The time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right-hand side of Equations~\eqref{eq:discfvdiffcc} and \eqref{eq:discfvadvcc} are assumed to be known. Explicit Runge–Kutta time integration methods are single-step multistage schemes widely used for advancing ordinary and partial differential equations~\cite{leveque2007finite}. The~FV counterpart of the thermodynamic divergence expression for \mbox{cut-cell $ii$ is}:
% start a new page without indent 4.6cm
%\clearpage
\end{paracol}
\nointerlineskip
\begin{eqnarray}
    ( \nabla \cdot \mathbf{u} )_{ii}^{\rm th} \; V_{ii} &=&
    \left[ \frac{1}{(\rho c_p T)_{ii}} - \frac{1}{\bar{p}_{ii}} \right]
    \frac{\partial \bar{p}_{ii}}{\partial t} \; V_{ii} +
    \frac{w_{ii} \: \rho_0 g_z}{(\rho c_p T)_{ii}} V_{ii} \nonumber \\
    &+& \frac{1}{(\rho c_p T)_{ii}} \left[ (\dot{q}'''+ \dot{q}_b''')_{ii} V_{ii} -
    \sum_{k=1}^{n_{\rm f}} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
    - \overline{\mathbf{u} \cdot \nabla (\rho h_s)} V_{ii} \right]  \\
    &+& \frac{1}{\rho_{ii}} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right)_{ii} \left[ (\dot{m}_\alpha'''+\dot{m}_{b,\alpha}''')_{ii} V_{ii} -
    \sum_{k=1}^{n_{\rm f}} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k
    - \overline{\mathbf{u} \cdot \nabla (\rho Y_\alpha)} V_{ii} \right] \nonumber \label{eq:divth2},
\end{eqnarray}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

\noindent where the over-line terms refer to flux-limited interpolation of corresponding scalars, and~terms defined with subscript $ii$ refer to cell-defined quantities. All heat and mass fluxes and scalars are assumed to be known. Additionally, the~vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid. Details of the FDS time-integration scheme can be found in the References~\cite{FDS_Users_Guide,mcdermo_2014}. For~each RK2 substep, the~species transport equations are advanced in all FDS Cartesian cells, and~then the solution (explicit fluxes and scalar densities) is recomputed on the unstructured cut-cell region. A~similar procedure is done for the thermodynamic divergence.
% Small cell linking for explicit time integration.
As an explicit time-integrator is used, in~general, there will arise cut-cells whose small size will severely penalize the time-step. These cut-cells are linked to larger surrounding cells. Cell-linking, in the context of momentum equations, can be found in the Reference~\cite{kirk_2003}. In~FDS, the~momentum equations and pressure Poisson equation are solved on the Cartesian mesh. Therefore, an~immersed boundary method is used to reconstruct velocities in the cut-cell region. A~divergence integral equivalence argument is used to transfer the divergence from cut-cells to the underlying Cartesian cells in order to build the source term of the Poisson equation for the~pressure.

\subsection{Immersed Boundary Method and Wall~Modeling}

Collecting advective, shear stress, and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within Equation~(\ref{eqn:momtran}), the~model momentum transport problem can be written as~\cite{FDS_Users_Guide}:
\begin{eqnarray}
  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \mathbf{F}(\mathbf{u},\mathbf{x},t) - \boldsymbol{\nabla} H(\mathbf{x},t)  \label{eq:LowMachMom} \\
         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{\rm th} \label{eq:LowMachDiv},
\end{eqnarray}
where Equation~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by Equation~\eqref{eq:divth}.
Boundary conditions are prescribed for $\mathbf{u}(\mathbf{x},t)$ on boundaries, including the immersed~terrain.

The previous equations are advanced in time using a fractional step method. As an~illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$. Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1}$ are known:
\begin{eqnarray}
  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \mathbf{F}^n -  \boldsymbol{\nabla} H^n \label{eq:LowMachMomEu}\\
  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1} \label{eq:LowMachDivEu},
\end{eqnarray}
where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$. This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator. The~potential field $H(\mathbf{x},t)$ does not have a time evolution equation, and it is responsible for enforcing the divergence condition and is used on the projection step. Taking the divergence of Equation~\eqref{eq:LowMachMomEu} and considering  the constraint, Equation~\eqref{eq:LowMachDivEu}, the~two steps of the method~are:
\begin{enumerate}
  \item Solve Poisson equation for $H^n$:
\begin{equation}
   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u} \right)^{\mathrm{th},\,n+1} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n. \label{it:FSPoisson}
\end{equation}

  \item Obtain final velocity for step:
\begin{equation}
     \mathbf{u}^{n+1} = \mathbf{u}^{n} - \Delta t \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{it:FSProject}.
   \end{equation}

\end{enumerate}

A consequence of the projection scheme is that boundary conditions are required on the Poisson equation, Equation~\eqref{it:FSPoisson}. For~explicit methods and stationary \textit{solid} boundaries, the~corresponding boundary condition is \textit{homogeneous} Neumann for $H^n$ in $\partial \Omega,\partial \Omega_1,...,$ $\partial \Omega_{\rm nbods}$~\cite{perot_1993}.
Next, an~approximation to the no-slip boundary condition at the immersed terrain surface is needed. To~this end, a~direct forcing immersed boundary method (IBM) for the momentum equations~\cite{fadlun_2000} is employed. A~force field is computed on the discrete momentum equations on grid faces crossed by the immersed surfaces to approximate the no-slip boundary condition on these. In~LES, the~surrounding velocity field is modeled using an equilibrium boundary layer~solution.

In Figure~\ref{Fig:figure_2}, the~velocity update in each of the gas phase cut-face centroids $d$ is done by individualizing point $B$ on the boundary and the normal direction through these. \mbox{Additionally, an~external} point $e_x$ through the normal $\mathbf{\hat{n}}$ is defined at a distance $\delta_{ex}$, of~the order of the Cartesian cells' size. Known velocities and fluid parameters are interpolated from the surrounding fluid points $e_1,\dots,e_4$ to $e_x$.
This information is used to estimate a target velocity at step $n+1$, $\mathbf{\hat{u}}_d^{n+1,k-1}$ at point $d$, assuming the log law equilibrium boundary layer solution of Section~\ref{sec:boundflx}. The~target velocity $\hat{u}_d^{n+1,k-1}$ component on the cut-face centroid is flux-matched to the underlying Cartesian face $E$ velocity component $\hat{u}_E^{n+1,k-1}$. \mbox{Finally, an~immersed} boundary force
\begin{equation}
F_E^k=-\frac{\hat{u}_E^{n+1,k-1}-u_E^{n}}{\Delta t} - \frac{\partial H^{n,k-1}}{\partial x}
\label{eq:FIBM}
\end{equation}
can be computed and used in Equation~\eqref{it:FSPoisson} to take into account the presence of the body. The~index $k$ refers, in this context, to the sub-iteration that can be performed in the IB force, Equation~\eqref{eq:FIBM}, and~projection, Equations~\eqref{it:FSPoisson} and \eqref{it:FSProject}, to~match final velocities with the wall-modeled targets. This velocity reconstruction procedure has been combined with the level set method for fire spread described in the next~section.

\section{Wildland Fire~Spread} \label{sec:firespread}

There are various types of wildland fire spread models to choose from, and~the choice depends on the desired level of physical fidelity and the available computing resources. The~choice also depends on the nature of the fire. For~example, for~surface vegetation like dry grass, the~fire behavior depends on the orientation of the local wind relative to the direction of fire spread. For~head fires, the~wind flow and fire spread are in the same direction. Back fires spread against the wind; flank fires spread laterally. Typically, the~head fire exhibits the widest front width, most rapid spread, and~longest flame lengths that tilt toward the virgin vegetation. Back fires exhibit the least rapid spread, most narrow front width, and~shortest flame lengths that tilt away from the virgin vegetation. Of~these three types of fire, the~head fire is least demanding in terms of grid resolution, and the back fire, the~most.
\begin{enumerate}
\item {\bf Particle Model:} % The bold makes the three fire spread methods names stand out in this enummeration. We like it.
The vegetation (surface and raised) is represented by a collection of Lagrangian particles that are heated via convection and radiation. This model, with~sufficient grid resolution, is appropriate for head, back, and~flank surface fires, as~well as fire through raised vegetation (e.g., trees). Heat transfer in the volume containing the vegetation is modeled in all three directions. This model is appropriate for grid resolutions of the order of 1~m or less, depending on the size of the flame base and properties of the vegetation (e.g.,~\cite{Perez-Ramirez:FT2017}).
\item {\bf Boundary Fuel Model:} Surface vegetation has its own grid and is modeled like a porous solid with a thickness equal to the height of the vegetation. This model was designed for head fire spread in surface vegetation based on the assumption that heat transfer in the fuel bed is dominated by radiation from the overhead flame, and therefore in the vertical direction.  In~the implementation here, the~height of the surface vegetation is assumed to be unresolved on the grid. The~appropriate gas-phase grid resolution of the order is of 1~m to 10~m.
\item {\bf Level Set Method:} The fire-front of a surface fire propagates using purely empirical rules in a level set method. Thermal degradation of the surface vegetation is not modeled. More than one implementation of this method is possible, largely differentiated by how the wind and fire-atmosphere interaction is modeled. The~simpler implementations of this model can use grid resolutions that are coarser and 10 m or greater, than~the more physics-based particle and boundary fuel models. The~level set model can be used for fire spread in the surface vegetation, along with the particle model for fire behavior in raised vegetation. 
\end{enumerate}

The Particle Method and Boundary Fuel Model require thermo-physical properties of the vegetative fuels, and the fire spread rate is {\em predicted} by the model. The~Level Set Method relies on a set of predetermined spread rates for different types of vegetation and wind speeds. The~pyrolysis model, developed specifically for vegetation~\cite{Porterie:2006,Morvan:CF2004,Houssami:2016}, \mbox{consists of} three reactions:

\vspace{\baselineskip}

\noindent 1. Endothermic moisture evaporation
\begin{equation}
 {\rm Wet\ Vegetation} \rightarrow \nu_{\rm H_2O} \, {\rm H_2O} + (1-\nu_{\rm H_2O}) \, {\rm Dry\ Vegetation} \quad ; \quad \nu_{\rm H_2O} = \frac{M}{1+M};
\end{equation}

\noindent 2. Endothermic pyrolysis of dry vegetation
\begin{equation}
 {\rm Dry\ Vegetation} \rightarrow \nu_{\rm char} \, {\rm Char} + (1-\nu_{\rm char}) \, {\rm Fuel\ Gas};
\end{equation}

\noindent 3. Exothermic char oxidation
\begin{equation}
 \label{char_reaction}
 {\rm Char} + \nu_{\rm O_2, char} \, {\rm O_2} \rightarrow (1+ \nu_{\rm O_2,char} - \nu_{\rm ash}) \, {\rm CO_2} + \nu_{\rm ash} \, {\rm Ash}.
\end{equation}

$M$ is the vegetation {\em moisture content} or {\em moisture fraction} determined on a dry weight basis. $\nu_{\rm char}$ is the mass fraction of dry vegetation that is converted to char during pyrolysis.   $\nu_{\rm O_2,char}$ is the mass of oxygen required per unit mass of char consumed. $\nu_{\rm ash}$  is the mass fraction of char that is converted to ash during char oxidation. Reaction rates and the various empirical kinetic constants are given in the Reference~\cite{Porterie:2006,Morvan:CF2004,Houssami:2016}.

\subsection{Particle~Model}
\label{particle_model}

Lagrangian particles represent different types of vegetation, like leaves, grass, and~pine needles. Each type of vegetation is represented by a single particle in each computational grid cell (multiple particles within a cell are permitted, but~may not be the most computationally efficient approach).
Most often, the particle is assumed to be cylindrical in shape, but~it can be spherical or planar. The~particle is assigned a diameter based on the measured surface area to volume ratio, $\sigma'$. The~length is relatively unimportant, so long as it is assumed to be much greater than the diameter. Material properties are assigned for the wet vegetation, which is assumed to dry out, decompose to char, and~then exothermally oxidize when exposed to heat from an oncoming fire. The~drag force exerted by the collection of particles is \mbox{given by}:
\begin{equation}
   \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma' \, \mathbf{u} \, \|\mathbf{u}\|,
   \label{drag}
\end{equation}
where $\rho$ is the air density, $C_{\rm d}$ is the drag coefficient, $C_{\rm s}$ is the shape factor (ratio of projected area to surface area), $\beta$ is the packing ratio, $\sigma'$ is the surface area to volume ratio, and~$\mathbf{u}$ is the air velocity. The~diameter is specified via the surface area to volume ratio, and $\sigma'=2/r$ for a cylinder. The~packing ratio is the volume of solid needles divided by the volume they occupy, typically denoted $\beta$. It is calculated by dividing the dry mass of vegetation per unit volume, the~so-called ``bulk density'', by the density of the dry vegetation, $\rho_{\rm d}$.

The drag coefficient for miscellaneous vegetation is taken to be 2.8, based on wind tunnel measurements~\cite{Falkenstein-Smith:2018}. The~shape factor is assumed to be 0.25, which is the exact value for spheres and an approximate value for randomly oriented cylinders. In~practice, the~single term
\begin{equation}
   \kappa = C_{\rm s}  \, \beta \, \sigma'
\end{equation}
can be easier to determine than the three individual terms by measuring the relative amount of sunlight, $W$, that penetrates through a layer of vegetation of depth $\delta$:
\begin{equation}
   \kappa = -\frac{\ln \, W}{\delta}.
\end{equation}

This value of $\kappa$ also serves as the absorption coefficient for thermal radiation for a collection of~particles.

The Lagrangian particles that represent vegetation are assumed to be thermally thick, but~for simplicity, the heat conduction within the particle is assumed to be one-dimensional in either a cylindrical or spherical coordinate system. It is assumed that the particles interact with the surrounding gas via an additional source term in the energy conservation equation. For~a grid cell with indices $ijk$, the~source term is:
\begin{equation} \label{eq:qr}
   (-\nabla \cdot \dot{\mathbf{q}}_{\rm r}'')_{ijk} = \sum \kappa_{\rm p} \left( U_{ijk} - 4 \sigma \, T_{\rm p}^4 \right),
\end{equation}
where $T_{\rm p}$ is the particle surface temperature and the summation is over all the particles within the cell. The~effective absorption coefficient for a single particle is given by
\begin{equation}
   \kappa_{\rm p} = \frac{f \, A_{\rm p}}{4 \, V} = C_{\rm s}  \, \beta \, \sigma',
\end{equation}
where $A_{\rm p}$ is the surface area of the particle, $V$ is the volume of the cell, and~$f$ is the number of actual particles represented by the single modeled particle. Notice that the absorption coefficient plays a role for both drag and radiation heat~transfer.

The net radiative heat flux onto the surface of the particle is taken as:
\begin{equation}
   \dot{q}_{\rm r}'' = \epsilon \left( \frac{U_{ijk}}{4} - \sigma T_{\rm p}^4 \right),
\end{equation}
where $\epsilon$ is the emissivity of the particle surface, typically assumed to be about 0.9. The~convective heat transfer to the surface is given by:
\begin{equation}
   \dot{q}_{\rm c}'' = h \, (T_{\rm g} - T_{\rm s})  \quad ; \quad  h = \frac{k}{L} \, \hbox{Nu}  \label{q_con}.
\end{equation} 
$L$ is a characteristic length equal to the diameter of the sphere or cylinder. The~Nusselt number is given by:
\begin{equation}
   \hbox{Nu} = C_1 + C_2 \, \hbox{Re}^n \, \hbox{Pr}^m  \quad ; \quad \hbox{Re} = \frac{\rho |\mathbf{u}| L}{\mu} \quad ; \quad \hbox{Pr} = 0.7.
\end{equation}

For cylindrical particles, the~default values are $C_1=0$, $C_2=0.683$, $n=0.466$, $m=0.33$, and~$L=D$, the~diameter of the cylinder. For~spherical particles, the~default values are $C_1=2$, $C_2=0.6$, $n=0.5$, $m=0.33$, and~$L=D$, the~diameter of the sphere~\cite{Incropera:1}.

\subsection{Boundary Fuel~Model}

In many simulations of wildland fire, the~surface vegetation layer is too shallow to be resolved explicitly, as~is done when using Lagrangian particles to represent the vegetation. In~such cases, the~surface vegetation can be modeled as a porous boundary consisting of a layer of dry vegetation, moisture, and~air, underneath which is hard ground~\cite{Mell:IJWF2007}. The~drag exerted by the vegetation is modeled using a special velocity boundary condition, and~convective heat transfer is modeled via a source term in the one-dimensional heat conduction equation that is solved through the layer of vegetation and solid ground. Thermal radiation penetrates the vegetation layer via a 1D radiative transport equation that is used for semi-transparent~solids.

The Boundary Fuel and Particle Models share the same basic input parameters and pyrolysis model. The~drag exerted on the wind flowing through the vegetation is imposed as a force term in the gas phase grid cell adjacent to the boundary:
\begin{equation}
    \mathbf{f}_{\rm b} = \frac{\rho}{2} \, C_{\rm d} \, C_{\rm s} \, \beta \, \sigma' \, \frac{h_{\rm b}}{\delta z} \, \mathbf{u} \, \|\mathbf{u}\|,
\end{equation}
where the parameters are the same as in Equation~(\ref{drag}) except for the additional depth of vegetation, $h_{\rm b}$, height of a grid cell, $\delta z$, and~$\mathbf{u}$ is the gas velocity in the first grid~cell.

Thermal radiation is absorbed in depth according to a 1D radiative transport solver. The~absorption coefficient is given by:
\begin{equation}
   \kappa = C_{\rm s} \, \sigma' \, \beta.
\end{equation}

Thermal convection is not imposed at the interface between the gas phase and the vegetation layer, but~rather imposed via a source term in the 1D heat conduction solver:
\begin{equation}
   \langle \dot{q}_{\rm c,b}''' \rangle = \sigma' \, \beta \, \dot{q}_{\rm c}'', 
\end{equation}
where $\dot{q}_{\rm c}''$ is given by Equation~(\ref{q_con}) and the gas velocity, $|\mathbf{u}|$, and~temperature, \mbox{$T_{\rm g}$, are taken} from the first gas phase grid cell adjacent to the boundary, and~the particle surface temperature, $T_{\rm s}$, is extracted from the solution of the 1D heat conduction~equation.

\subsection{Level Set~Model}

For simulations of wildland fires spanning large areas that cannot be gridded finely enough to predict fire spread using the particle or boundary fuel models for vegetation, fire spread via a level approach is used~\cite{Bova:IJWF2015}. It is assumed that a surface fire spreading from a point under certain wind, slope, and~vegetation conditions does so with an ellipse-shaped fire front with, for~an effective wind vector, a~fixed length-to-breadth ratio~\cite{Finney:FARSITE,Bova:IJWF2015}. Determining the spread rate vector at a point on the simulated fire front requires a formula for the magnitude of head fire rate of spread which is adjusted based on the location of the point on the idealized ellipse (e.g., head versus flank fire). This location is determined from the orientation of the effective wind vector relative to the normal to fire front represented by the level~set.

The location of a fire front at time $t$ is defined by the set of points $x,y$, where the level set scalar function $\phi(x,y,t)=0$. The~advection equation for the level set function is
\begin{equation}
   \frac{D \phi(x,y,t)}{D t} =  \frac{\partial \phi}{\partial t} + \mathbf{R}_{uv} \cdot \nabla (\phi) = 0 \label{eqn:lset},
\end{equation}
where $\mathbf{R}_{uv}=(R_u,R_v)$ is the spread rate~vector.

As stated above, the~magnitude of the spread rate vector on a point on the level set fire front is determined by adjusting the magnitude of the head fire spread rate. The~direction of the spread rate vector is assumed to be normal to the fire front. A~formula for magnitude of the head fire spread rate can be obtained a number of ways, including empirical models~\cite{Cheney:IJWF1998} and physics-based simulations~\cite{Mell:FBFC2019}. A~commonly used semi-empirical model is the Rothermel model~\cite{Rothermel:1972,Albini:1976}, which can be implemented in FDS. Here, the~magnitude of the head fire spread rate is
\begin{equation}
  \|\mathbf{R}_{uv}\|=R_0 \left(1 + \sqrt{(\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) \cdot (\boldsymbol{\phi}_W+\boldsymbol{\phi}_S) } \right),
\end{equation}
where $R_0$ is the zero-velocity, zero-slope rate of spread that depends on the fuel properties. The~vectors $\boldsymbol{\phi}_W,\boldsymbol{\phi}_S \ge 0$ are computed using the local wind and slope through empirical rules~\cite{Wilson:1980}. See Appendix 1 of Reference~\cite{Bova:IJWF2015} for details on these calculations. The~Rothermel model has been used in other CFD-based approaches, including the atmospheric weighted models WRF-SFIRE~\cite{Mandel:2014} and CAWFE~\cite{Coen:2}.

In FDS, the~level set spread model can be implemented in the following~ways:
\begin{enumerate}
\item Only the level set simulation is performed, with~a constant and uniform specified wind and slope. The~wind is not affected by the terrain, and~there is no fire.
\item The wind field is established over the terrain, but~it is ``frozen'' when the fire ignites.
\item The wind field follows the terrain, but~there is no actual fire in the simulation, \mbox{just front-tracking}.  The~level set evolves continuously in time with the flow field.
\item The wind and fire are fully coupled, and the resulting wind values are used in the head fire spread-rate formula. When the fire-front arrives at a given surface cell, \mbox{it burns} for a finite duration and with a heat release per unit area provided as part of the fuel model.
\item The wind and fire are fully coupled in the gas phase, but~the head fire spread-rate is not influenced by the wind speed. 
\end{enumerate}

When fully coupled to the CFD model, the~level set function acts as an igniter, as it reaches a new surface cell. The~cell burns for a pre-determined amount of time. For~cases where the depth of the fireline is less than the size of a grid cell, the~burning time is extended, while the burning rate is decreased so as to maintain conservation of fuel~mass.

\section{Atmospheric Wind Boundary~Conditions}
\label{sec:wind}

For outdoor flows, FDS prescribes the wind field through special ``open'' boundary conditions on the exterior of the computational domain.  The~boundary conditions consist of three~components:
\begin{enumerate}
\item[(i)] The specified upstream wind field (vertical profile of streamwise velocity components) based on prescribed Monin-Obukhov parameters imposed on fluid elements \emph{entering} the domain;
\item[(ii)] \textls[-20]{Optional specification of upstream turbulence based on Jarrin's synthetic eddy method~\cite{Jarrin:2008}} \textls[-20]{(which is possible with the code, but~not utilized in the test cases within this \mbox{paper); and}}
\item[(iii)] \textls[-30]{Nonuniform and nonstationary Dirichlet pressure boundary values for the \mbox{Poisson equation}.}
\end{enumerate}

The following sections provide more detail on the implementation of the boundary profiles and the pressure boundary~conditions.

\subsection{Velocity and Temperature~Profiles}
\label{sec:veltmpprof}

While completely custom vertical boundary profiles of velocity and temperature are possible, the wind field is usually specified via Monin–Obukhov (MO) similarity parameters~\cite{Dyer:1974}: Obukhov length scale, $L$, aerodynamic roughness height, $z_0$, wind speed, $u_r$, and~temperature, $T_r$, at~a reference height $z_r$.  For~a given set of MO parameters, \mbox{FDS generates} velocity and temperature profiles as functions of height, as shown in \mbox{Figure~\ref{fig:MOprofs}}. These values are specified for any fluid elements entering the domain at height $z$, including along the top of the domain.  Convective outflow boundaries are applied locally for any fluid elements leaving the domain.  The~profiles can vary in time allowing one-way ``downscaling'' from either weather observations or output from a numerical weather prediction (NWP) model.  However, two-way coupling with an NWP model (e.g., \cite{coen_2013,CAWFE}) has not yet been implemented with~FDS.

\begin{figure}[H]
%\centering
\includegraphics[width=.45\textwidth]{figures/vel_L=-500.pdf}

\includegraphics[width=.45\textwidth]{figures/tmp_L=-500.pdf}
\caption{Monin--Obukhov velocity and temperature profiles for inflow boundary conditions.  Here, parameters are $L=-500\;\mathrm{m}$, $z_0=0.03\;\mathrm{m}$, and~$u_r=4.8\;\mathrm{m/s}$ and $T_r=34\;^\circ\mathrm{C}$.} 
\label{fig:MOprofs}
\end{figure}

% \subsection{Turbulent Inflow Conditions}

% It is possible in FDS to superimpose synthetic turbulence on a specified inlet profile utilizing Jarrin's synthetic eddy method (SEM) \cite{Jarrin:2008}.  An example verification case is presented in Figure~\ref{fig:semprofs}.  The left image of the figure shows contours of the normal component of velocity at the inlet.  The plot on the right shows the mean and rms profiles compared to the specified inlet parameters.  The domain is 20 m in the streamwise dimension by 12 m in the spanwise by 13 m in height.  In The turbulence intensity is set to 10 \% and the turbulence integral length scale is set to 3 m.  The specified MO parameters are $L=-667\;\mathrm{m}$, $z_0=0.022\;\mathrm{m}$, and $u_r=10\;\mathrm{m/s}$ and $T_r=20\;^\circ\mathrm{C}$ at reference height of 10 m.  The simulation runs for 60 seconds with statistics collected over the last 30 seconds.  The lateral, top, and downstream boundaries are set to ``open'' pressure boundaries as discussed in the next section.

% \begin{figure}[ht]
% %\centering
% \includegraphics[width=0.5\textwidth,valign=c]{figures/sem_open_wind_contours.pdf}
% \includegraphics[width=0.48\textwidth,valign=c]{figures/sem_open_wind_u_prof.pdf}
% \caption{Example of inflow synthetic turbulence superimposed on a Monin-Obukhov boundary layer profile. Contours of the normal component of velocity at the domain inlet taken at a time midway through the simulation (left); mean and rms profiles along the inlet compared with specified profiles (right).}
% \label{fig:semprofs}
% \end{figure}

% Other approaches to imposing boundary layer turbulence are possible (e.g., \cite{Munoz-Esparza:2014}).  The cases presented later in this paper do not utilize SEM since detailed data on the turbulence statistics were not available.

\subsection{Pressure Boundary~Values}

The Poisson equation for $H$, Equation~(\ref{it:FSPoisson}), links the momentum equation with mass and energy in the low-Mach flow algorithm.  The~boundary conditions for this equation determine the inflow and outflow velocity component values.  Let $\mathbf{n}$ denote a unit normal vector pointing \emph{into} the domain.  Boundary values at an interface are denoted with a subscript ``$I$''.  Then, $\mathbf{u}_I\cdot\mathbf{n}>0$ is an inflow, and $\mathbf{u}_I\cdot\mathbf{n}<0$ is an outflow~condition.

At an inflow, the mean viscous and convective forces are generally small, and a simplified momentum equation   may be used to develop a boundary value for $H$, namely, $\partial u_i/\partial t = -\partial H/\partial x_i$. To~develop the boundary condition, let $u$ denote the $x$-component of velocity and consider an interface normal to $x$ pointing into the domain.  The~prescribed external wind field velocity component at height $z$ and time $t$ (discussed above in \mbox{Section~\ref{sec:veltmpprof}}) is denoted $u_{wind}(z,t)$.  Using a one-sided difference for $H$ and an explicit Euler approximation of the velocity time derivative, the~boundary value for $H$ may be written as
\begin{equation}
\label{eq:Hin}
H_I = H_{I\pm\frac{1}{2}}^n \pm \frac{\Delta x}{2}\left[\frac{u_{wind,I}(z,t) - u_I^n}{\Delta t}\right] \quad \mbox{if} \quad \mathbf{u}_I\cdot\mathbf{n}>0.
\end{equation}

At an outflow boundary, $H$ is set equal to the local kinetic energy per unit mass from the cell just upwind of the boundary:
\begin{equation}
\label{eq:Hout}
H_I = \frac{1}{2}(\bar{u}^n \bar{u}^n + \bar{v}^n \bar{v}^n + \bar{w}^n \bar{w}^n)_{I\pm\frac{1}{2}} \quad \mbox{if} \quad \mathbf{u}_I\cdot\mathbf{n}<=0.
\end{equation}

The overbar on the velocity components denotes a linear interpolation of the primitive staggered component values to the cell~center.

\section{Numerical~Experiments} \label{sec:numexp}
\unskip

\subsection{Flat Terrain Fire~Spread}  \label{sec:simexp}

In July and August of 1986, the~Commonwealth Scientific and Industrial Research Organisation (CSIRO) of Australia conducted controlled grassland fire experiments near Darwin, Northern Territory~\cite{Cheney:IJWF1993}. July and August are in the middle of the dry season when the grasses are fully cured (dried) and the weather is warm and dry. The~experiments were conducted on flat plots measuring 100~m by 100~m, 200~m by 200~m, or~200~m by 300~m. Two cases have been simulated. Case~C064 was conducted on a 100~m by 100~m plot of kerosene grass ({\it Eriachne burkittii}); Case~F19 was conducted on a 200~m by 200~m plot of kangaroo grass ({\it Themeda australis}).

Two of these experiments were originally simulated with FDS by Mell~et~al.~\cite{Mell:IJWF2007} using a form of the Boundary Fuel Model. Now these two experiments are also simulated using the Lagrangian Particle Model and the Level Set Model. The~level set simulations of Case~C064 use fuel index 1 (Short Grass) and for Case~F19, fuel index 3 (Tall Grass)~\cite{Rothermel:1972,Albini:1976}.

Measured bulk properties of the grasses burned in the two experiments are listed in Table~\ref{Properties_Grasses}. Properties that were not measured are listed in Table~\ref{Assumed_Properties_Grasses}. These assumed properties are typically for wood or cellulosic fuels. The~moisture is modeled as water. The~grass is assumed to be composed primarily of~cellulose.

\begin{specialtable}[H] 
\caption[Measured properties for the CSIRO Grassland Fire cases]{Measured properties for the CSIRO Grassland Fire cases~\cite{Cheney:IJWF1993}.}
\label{Properties_Grasses}
\setlength{\cellWidtha}{\columnwidth/4-2\tabcolsep+0.0in}
\setlength{\cellWidthb}{\columnwidth/4-2\tabcolsep+0.0in}
\setlength{\cellWidthc}{\columnwidth/4-2\tabcolsep+0.0in}
\setlength{\cellWidthc}{\columnwidth/4-2\tabcolsep+0.0in}
\scalebox{1}[1]{\begin{tabularx}{\columnwidth}{>{\PreserveBackslash\raggedright}m{5cm}>{\PreserveBackslash\centering}m{2cm}>{\PreserveBackslash\centering}m{2.5cm}>{\PreserveBackslash\centering}m{3cm}}
\toprule
\textbf{Property}                        & \textbf{Units}        & \textbf{Case C064} & \textbf{Case F19}    \\ \midrule 
Wind Speed                      & m/s          & 4.6       & 4.8         \\ \midrule
Ambient Temperature             & $^\circ$C    & 32.       & 34          \\ \midrule
Surface Area to Volume Ratio    & m$^{-1}$     & 9770      & 12240       \\ \midrule
Grass Height                    & m            & 0.21      & 0.51        \\ \midrule
Bulk Mass per Unit Area         & kg/m$^2$     & 0.283     & 0.313       \\ \midrule
Moisture Fraction               & \%           & 6.3       & 5.8         \\ \midrule
Measured RoS                    & m/s          & 1.2       & 1.5         \\ \midrule
Calc'd RoS, Particle Method (0.25 m, 0.5 m, 1.0 m resolution) 
                                & m/s          & 1.1, 1.2, 1.2   & 1.4, 1.3, 1.4          \\ \midrule
Calc'd RoS, Boundary Fuel Method (0.25 m, 0.5 m, 1.0 m resolution) 
                                & m/s          & 1.3, 1.3, 1.3   & 1.5, 1.4, 1.7          \\ \midrule
Calc'd RoS, Level Set Method (5 m, 10 m, 20 m resolution) 
                                & m/s          & 0.5, 0.6, 0.6   & 1.0, 1.1, 1.2          \\ \bottomrule
\end{tabularx}}
\end{specialtable}

Snapshots of the Lagrangian particle simulation of Case~F19 compared to photographs of the experiment are shown in Figure~\ref{F19}. The~plot of grass is 200~m by 200~m, and the computational domain is 240~m by 240~m by 20~m high. The~grid cells are 0.5~m cubes. The~domain is subdivided into 36 individual meshes and run in parallel. A~blade of grass is represented by a single cylindrically shaped particle within a grid cell. The~radius of the cylinder is derived from the measured surface area to volume ratio of the grass. Each simulated blade of grass represents many more actual blades of grass. The~weighting factor is determined from the measured bulk mass per unit area. The~fires in the experiments were ignited by two field workers carrying drip torches walking in opposite directions along the upwind boundary of the plot (the red strip in Figure~\ref{F19}). In~FDS, this action was modeled using a specified spread rate along the~strip.

% start a new page without indent 4.6cm
%\clearpage
\end{paracol}
\nointerlineskip
\begin{figure}[H]
\widefigure
\includegraphics[width=0.9\textwidth]{figures/F19_collage.png}
\caption{Photographs of the experiment and snapshots of the simulation (medium resolution Particle Model) of CSIRO Grassland Fire F19, 56~s, 86~s, and~138~s following~ignition.}
\label{F19}
\end{figure}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

The simulations were conducted with the three different fire spread models, and~each case was run with three levels of spatial resolution. The~simulations using the Particle and Boundary Fuel Models were run with cubic grid cells that were 0.25~m, 0.5~m, and~1~m on a side. The~simulations using the Level Set Model were run with cells of 5~m, 10~m, and~20~m. The~predicted rates of spread for the two cases are listed in \mbox{Table~\ref{Properties_Grasses}} and shown graphically in Figure~\ref{CSIRO}. For~these simulations, the~rate of spread predicted by the Particle and Boundary Fuel Models are comparable to the measured rate of spread, while the level set method under-predicts the RoS. The~under-prediction by the level set simulations is a consequence of the limitations of the Rothermel model, which is based on empirical relations largely derived from laboratory-scale experiments. It is well-known that practical application of the Rothermel model often requires calibration for the case of interest~\cite{Arca_2007}. The~use of the Rothermel for the head fire spread rate in surface vegetation in FDS is to be viewed as a placeholder that is in-line with other CFD-based models. If~the observed head fire rate of spread for the F19 and C064 experiments is used in FDS, \mbox{then good} agreement with the observed fire perimeter is obtained (not shown). \mbox{Further work} is needed to develop rate-of-spread models appropriate for use with the Level Set Model (e.g.,~\cite{Mell:FBFC2019}).

 \begin{specialtable}[H]
\caption[Assumed properties for dry grass and soil]{Assumed properties for various types of dried grass and soil. Note that the Pyrolysis Temperature is taken to be the temperature at which the mass loss rate peaks in the TGA experiments of Morvan and Dupuy~\cite{Morvan:CF2004}.}
\setlength{\tabcolsep}{4.1mm} 
\label{Assumed_Properties_Grasses}
\begin{tabular}{lccc}
\toprule
\textbf{Property}                        & \textbf{Units}                 & \textbf{Value}                     & \textbf{Reference}                             \\ \midrule 
Chemical Composition            & --                    & C$_6$H$_{10}$O$_5$        & Assumption                            \\ \midrule
Heat of Combustion              & kJ/kg                 & 15600                     & \cite{Susott:FS1982}                  \\ \midrule
Soot Yield                      & kg/kg                 & 0.015                     & \cite{SFPE:Tewarson}                  \\ \midrule
Char Yield                      & kg/kg                 & 0.2                       & \cite{Susott:FS1982}                  \\ \midrule
Specific Heat                   & kJ/(kg$\cdot$K)       & 1.5                       & Various sources                       \\ \midrule
Conductivity                    & W/(m$\cdot$K)         & 0.1                       & Assumption                            \\ \midrule
Density                         & kg/m$^3$              & 512                       & \cite{Rothermel:1972}                 \\ \midrule
Heat of Pyrolysis               & kJ/kg                 & 418                       & \cite{Morvan:CF2004}                  \\ \midrule
Pyrolyis Temperature            & $^\circ$C             & 200                       & \cite{Morvan:CF2004}                  \\ \midrule 
Obukhov Length                  & m                     & -500                      & Assumption                            \\ \midrule
Aerodynamic Roughness Length    & m                     & 0.03                      & Assumption                            \\ \midrule
Drag Coefficient                & --                    & 2.8                       & \cite{Falkenstein-Smith:2018}         \\ \midrule 
Soil Specific Heat              & kJ/(kg$\cdot$K)       & 2.0                       & \cite{Farouki:1981}                   \\ \midrule
Soil Conductivity               & W/(m$\cdot$K)         & 0.25                      & \cite{Farouki:1981}                   \\ \midrule
Soil Density                    & kg/m$^3$              & 1300                      & \cite{Farouki:1981}                   \\ \bottomrule

\end{tabular}
\end{specialtable}

Figure~\ref{fig:CaseF19_contours} shows contours of heat release rates near the surface compared to the observed flame front for the high-resolution (0.25 m) Particle Model.  These contours may be compared to the head fire spread rates at the top of Figure~\ref{CSIRO}.  Case F19 contours are shown on the right, and may be compared to the flame images in Figure~\ref{F19}.  These results show some improvement to the flank fire spread rates compared to previous simulations with FDS conducted by Mell~et~al.~in 2007~\cite{Mell:IJWF2007}.

However, it is difficult to compare the present results with those from~\cite{Mell:IJWF2007}. The~underlying numerical model, FDS, has undergone a major revision over the past decade, changing the basic finite-difference scheme, boundary conditions, drag modeling, and~so on, and~the vegetation-specific models have changed as well, most notably the charring reaction. In~addition, the~amount of information about the grasses and field conditions is limited, and~the thermal and kinetic parameters gleaned from various sources are generally average values over a wide range of vegetation types. A~sensitivity study~\cite{McGrattan:CI2017} for the particle model simulations reveals that near-surface wind speed has the most direct effect on the rate of spread (RoS), and~the combined effect of the dozens of thermal, \mbox{kinetic, and~numerical} parameters would certainly explain the difference in RoS between the 2007 results and those of this~study.

% start a new page without indent 4.6cm
%\clearpage
\end{paracol}
\nointerlineskip
\begin{figure}[H]
\widefigure
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
\includegraphics[height=2.2in]{figures/Case_C064} &
\includegraphics[height=2.2in]{figures/Case_F19} \\
\includegraphics[height=2.2in]{figures/Case_C064_BFM} &
\includegraphics[height=2.2in]{figures/Case_F19_BFM} \\
\includegraphics[height=2.2in]{figures/Case_C064_LS}  &
\includegraphics[height=2.2in]{figures/Case_F19_LS}
\end{tabular*}
\caption{Comparison of the measured and predicted fire front position for the CSIRO Grassland Fires using three different methods of fire~spread.}
\label{CSIRO}
\end{figure}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

% start a new page without indent 4.6cm
\clearpage
\end{paracol}
\nointerlineskip
\begin{figure}[H]
\vspace{-18pt}
\widefigure
\includegraphics[width=0.49\textwidth]{figures/Case_C064_flame_position.pdf}
\includegraphics[width=0.49\textwidth]{figures/Case_F19_flame_position.pdf}
\vspace{-6pt}
\caption{Contours of heat release rate in a plane near the surface from the fine resolution (0.25 m) Particle Model compared to observations of CSIRO Grassland Fires C064 (\textbf{left}) and F19 (\textbf{right}).}
\label{fig:CaseF19_contours}
\end{figure}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

\subsection{Complex Terrain Fire~Spread}  
\label{sec:cogo}

This section presents an example of the methodologies discussed in the paper. It is notable because it considers an actual fire that occurred in the wildland–urban interface (WUI), which presents a challenge for a fire spread model because the surface vegetation is mixed with the built~environment.

At approximately 23:00 (local time) on  25 March 2019, a~wildfire started in the seaside town of Cogoleto, near~Genoa, Italy. Strong winds, gusting up to 100~km/h from the north, spread the fire from its ignition point towards the sea. The~fire was caused by a faulty power line on a ridge above the town. Several houses were destroyed, and hundreds of residents had to be evacuated. The~burnt region measured over 1~km from the origin to the farthest point. The~fire burned overnight and into the next day. Up~to 70 firefighters worked at the scene, with~help from helicopters and fire hydrants. Suppression efforts were directed towards stopping the fire spread into the coastal town of~Cogoleto.

Shown in Figure~\ref{Cogoleto_satellite} are the results of a level set simulation of the fire superimposed on a satellite photograph of the area and the outline of the extent of the actual fire. The~computational domain was assembled using an open-source, Geographic Information System (GIS) program called QGIS~\cite{QGIS}, extended by the qgis2fds~\cite{qgis2fds} plugin. The~open-source qgis2fds plugin was specifically developed in the framework of the WUIFI-21 project for facilitating the use of geographical data for forest fire simulation and smoke pollutants dispersion in~FDS (WUIFI-21: {\it High-fidelity computational fluid dynamics modeling of forest fires for wildland–urban Interface communities resilience and protection} is an~Italy–US research project financed by the Italian Ministry of Foreign Affairs and International Cooperation).

The wind speed and direction are based on a single weather station near the point of ignition. Winds recorded were found to be predominantly coming from the north, with magnitudes that ranged from 20~m/s during the night and diminishing to 5~m/s towards the morning hours. The~vegetation type was provided by a database maintained at the International Centre on Environmental Monitoring (CIMA) %The journal does not allow footnotes, and the footnotes are unified into the main text. Please confirm.
Research Foundation, derived from regional forest management maps. In~the simulation, the~single-point wind data were taken as the prevailing wind, and~the local wind field was obtained via  CFD computation. The~fire was simulated based on the position of the level set front; that is, when the front arrived at a given location, a~fire was ignited and burned for a duration of time consistent with the specified fuel loading of that particular point on the~map.

% start a new page without indent 4.6cm
%\clearpage
\end{paracol}
\nointerlineskip
\begin{figure}[H]
\vspace{-12pt}
\widefigure
\includegraphics[width=0.9\textwidth]{figures/cogoleto_fire_2019_ls4_1000.png}
\caption{Satellite photograph of the region where the Cogoleto Fire occurred. The~area outlined in red is the extent of the actual fire; green is the~simulation.}
\label{Cogoleto_satellite}
\end{figure}
\begin{paracol}{2}
%\linenumbers
\switchcolumn

The extent of the simulated fire, shown in green in Figure~\ref{Cogoleto_satellite}, is greater than that of the actual fire, shown in red. This is not surprising, given that the simulation does not include suppression and the rate of spread of the level set front is determined based on vegetation that may not exactly resemble that of this particular region. It is important to note that for the level set method, a~fire break created by first responders can be modeled as a region declared as "non-combustible". The 13 Rothermel fuel types have been supplemented by areas such as water, pavement, and~other non-combustibles. The~fire-front determined by the level set calculation is stopped by the non-combustible area, highlighting the challenge in simulating firebrand spot ignitions that were reported by first responders at the~scene.

The simulation results have been compared with those obtained from the CIMA Propagator~\cite{Trucchia:2020}, an~experimental propagation model that has provided several European civil protection organizations with real-time fire predictions since 2009. The~model has been implemented by the CIMA Foundation, and its propagation model is based on stochastic cellular~automata.

The two models produce a qualitatively similar patterns of fire spread, mainly because the vicinity of the fire has numerous areas that are considered ``non-combustible',' that is, devoid of vegetation or structures. This points out the problem of ``validating'' wildfire models. The~simulations of the flat terrain grassland fire experiments described in the previous section match the observed fire behavior fairly well, but~in those cases, the~vegetation is uniform and fairly well-described in terms of its mass loading and geometrical characteristics, the~terrain is flat and of relatively small area, the~wind is steady, and~the experiment lasts a few minutes. In~a real fire, none of this will be true. The~detailed physics of the fire are not resolvable on a relatively coarse grid, the~vegetation is heterogeneous, sparse, and~sometimes unknown, the~weather conditions are varying, there are suppression efforts with water and the setting of backfires, and~the fires can burn for~days.

All of this raises the question as to what level of physical fidelity ought to be required in these models. The~strategy that has been adopted in FDS is to accommodate simulation options ranging from a level set calculation of fire spread over tens of kilometers that can be run in minutes, similar to models like FARSITE developed by the US Forest Service, \mbox{all the} way to detailed simulations of fire spread at sub-meter resolution for relatively short time-periods (minutes to hours) and small areas (tens of hectares). As~was the case for models developed for building fires 30 years ago, it is not clear at the moment exactly how these wildland fire models will be used in the future. The~best strategy is to advance the various modeling options until a clear path forward becomes~apparent.

\section{Summary} \label{sec:summary}

The numerical methods described in this paper span the range of length scales from tens of centimeters to tens of kilometers. Obviously, these techniques support differing descriptions of the underlying fire physics, but~all can be embedded within the same code base. This allows researchers to combine the different techniques in ways that were not possible when separate codes were maintained by separate organizations. With~FDS, \mbox{one can} simulate 24 hours of fire spread in minutes of computation time using the level set function and a wind field that varies only temporally, not spatially, with~no fire–atmospheric coupling. At~the same time, using the same set of input parameters, one can simulate the fire at far greater resolutions with much greater physical fidelity, albeit for shorter time-periods and spatial~extent.

\vspace{6pt}

\authorcontributions{Conceptualization, M.V.; Data curation, R.M.; Investigation, K.M. and W.M.; Software, E.G. and P.F.; Visualization, G.F.}

\funding{This work was partially funded by the Risk Reduction in Communities Program at NIST. W.M. acknowledges support from the Strategic Environmental Research and Development Program (SERDP) Grant RC19-1132.}

\dataavailability{The Fire Dynamics Simulator is an open-source computational fluids dynamics code available at \href{https://github.com/firemodels/fds}{https://github.com/firemodels/fds}. The input files for the CSIRO wildfire spread calculations are also kept within this repository.}

\conflictsofinterest{The authors declare no conflict of interest.}

\end{paracol}
\reftitle{References}

\begin{thebibliography}{999}

\bibitem[Thomas \em{et~al.}(2017)Thomas, Butry, Gilbert, Webb, and
  Fung]{thomas_2017}
Thomas, D.; Butry, D.; Gilbert, S.; Webb, D.; Fung, J.
\newblock  {The Costs and Losses of Wildfires: A Literature Review}.
\newblock In \emph{Special Publication NIST SP-1215}; National Institute of Standards
  and Technology: Gaithersburg, MD, USA,  2017.
\newblock
   {Available online: \url{https://www.nist.gov/publications/costs-and-losses-wildfires} (accessed on 05/02/2021).}

\bibitem[McDermott \em{et~al.}(2019)McDermott, Bryner, and
  Heintz]{mcdermott_2019}
McDermott, R.; Bryner, N.; Heintz, J.
\newblock  {Large Outdoor Fire Modeling (LOFM) Workshop Summary Report}.
\newblock In \emph{Special Publication NIST SP-1245}; National Institute of Standards
  and Technology: Gaithersburg, MD, USA,  2019.
\newblock
    Available online: \url{https://www.nist.gov/publications/large-outdoor-fire-modeling-lofm-workshop-summary-report} (accessed on 05/02/2021).

\bibitem[Richards \em{et~al.}(2020)Richards, Brew, and Smith]{richards_2020}
Richards, L.; Brew, N.; Smith, L.
\newblock  {2019–20 Australian bushfires—frequently asked questions: A Quick
  Guide}.
\newblock In \emph{Research Paper Series 2019-20}; Parliament of
  Australia: Canberra, Australia,  2020. ISSN 2203-5249. 
\newblock
   {Available online: \url{https://parlinfo.aph.gov.au/parlInfo/download/library/prspub/7234762/upload\_binary/7234762.pdf} (accessed on 05/02/2021).}

\bibitem[Papadopoulos and Niovi~Pavlidou(2011)]{Papadopoulos_2011}
Papadopoulos, G.D.; Niovi~Pavlidou, F.
\newblock A review of a new generation of wildfire–atmosphere modeling.
\newblock {\em IEEE Syst. J.} {\bf 2011}, {\em 5},~233--243.

\bibitem[Bakhshaii and Johnson(2019)]{Bakhshaii_2019}
Bakhshaii, A.; Johnson, E.
\newblock A review of a new generation of wildfire–atmosphere modeling.
\newblock {\em Can. J. For. Res.} {\bf 2019}, {\em
  49},~565--574.

\bibitem[Rothermel(1972)]{Rothermel:1972}
Rothermel, R.C.
\newblock A mathematical model for predicting fire spread in wildland fuels.
\newblock In \emph{Technical Report INT-115}; USDA Forest Service: Ogden, Utah,  1972.
\newblock
   {Available online: \url{http://www.treesearch.fs.fed.us/pubs/32533} (accessed on 05/02/2021).}

\bibitem[Anderson(1982)]{Anderson:1982}
Anderson, H.
\newblock {Aids to Determining Fuel Models for Estimating Fire Behavior}.
\newblock In \emph{General Technical Report INT-122}; Intermountain Forest and Range
  Experiment Station, USDA Forest Service: Ogden, Utah,  1982.
\newblock
   {Available online: \url{https://www.fs.usda.gov/treesearch/pubs/6447} (accessed on 05/02/2021).}

\bibitem[Finney(2004)]{Finney:FARSITE}
Finney, M.
\newblock {Fire Area Simulator–Model, Development and Evaluation}.
\newblock In \emph{Research Paper RMRS-RP-4 Revised}; United States Forest Service, Rocky
  Mountain Research Station, Missoula, Montana,  2004.
\newblock
   {Available online: \url{http://www.fs.fed.us/rm/pubs/rmrs\_rp004.pdf} (accessed on 05/02/2021).}

\bibitem[Bova \em{et~al.}(2015)Bova, Mell, and Hoffman]{Bova:IJWF2015}
Bova, A.; Mell, W.; Hoffman, C.
\newblock {A comparison of level set and marker methods for the simulation of
  wildland fire front propagation}.
\newblock {\em Int. J. Wildland Fire} {\bf 2015}, {\em
  25},~229--241.
\newblock{doi:10.1071/WF13178}.

\bibitem[McGrattan \em{et~al.}(2013)McGrattan, Hostikka, McDermott, Floyd,
  Weinschenk, and Overholt]{FDS_Users_Guide}
McGrattan, K.; Hostikka, S.; McDermott, R.; Floyd, J.; Weinschenk, C.;
  Overholt, K.
\newblock {\em {Fire Dynamics Simulator, User's Guide}}, 6th ed.;
\newblock National Institute of Standards and Technology: Gaithersburg,
  MD, USA; VTT Technical Research Centre of Finland: Espoo, Finland, 2013.

\bibitem[Coen(2013)]{Coen:2}
Coen, J.L.
\newblock {Modeling Wildland Fires: A Description of the Coupled
  Atmosphere-Wildland Fire Environment Model (CAWFE)}.
\newblock In~\emph{Technical Report NCAR/TN500+STR}; National Center for Atmospheric
  Research: Boulder, CO, USA, 2013.

\bibitem[Coen and Schroeder(2015)]{Coen:2015}
Coen, J.L.; Schroeder, W.
\newblock {The High Park Fire: Coupled weather-wildland fire model simulation
  of a windstorm-driven wildfire in Colorado's Front Range}.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 2015}, {\em 120},~131--146.

\bibitem[Lautenberger(2013)]{LAUTENBERGER_2013}
Lautenberger, C.
\newblock Wildland fire modeling with an Eulerian level set method and
  automated calibration.
\newblock {\em Fire Saf. J.} {\bf 2013}, {\em 62},~\mbox{289--298}.
\newblock {doi:10.1016/j.firesaf.2013.08.014}.

\bibitem[Mandel \em{et~al.}(2009)Mandel, Beezley, Coen, and Kim]{Mandel:2009}
Mandel, J.; Beezley, J.D.; Coen, J.L.; Kim, M.
\newblock {Data assimilation for wildland fires: Ensemble Kalman filters in
  coupled atmosphere-surface models}.
\newblock {\em IEEE Control  Syst. Mag.} {\bf 2009}, {\em 9},~47–65.
\newblock {doi:10.1109/MCS.2009.932224}.

\bibitem[Mandel \em{et~al.}(2011)Mandel, Beezley, and Kochanski]{Mandel:2011}
Mandel, J.; Beezley, J.D.; Kochanski, A.K.
\newblock {Coupled atmosphere-wildland fire modeling with WRF 3.3 and SFIRE
  2011}.
\newblock {\em \mbox{Geosci. Model} Dev.} {\bf 2011}, {\em 4},~591--610.
\newblock {doi:10.5194/gmd-4-591-2011}.

\bibitem[Mandel \em{et~al.}(2015)Mandel, Amram, Beezley, Kelman, Kochanski,
  Kondratenko, Lynn, Regev, and Vejmelka]{Mandel:2014}
Mandel, J.; Amram, S.; Beezley, J.D.; Kelman, G.; Kochanski, A.K.; Kondratenko,
  V.Y.; Lynn, B.H.; Regev, B.; Vejmelka, M.
\newblock {Recent advances and applications of WRF-SFIRE}.
\newblock {\em Nat. Hazards Earth Syst. Sci.} {\bf 2015}, {\em 14},~2829--2845.
\newblock {doi:10.5194/nhess-14-2829-2014}.

\bibitem[Kochanski \em{et~al.}(2016)Kochanski, Jenkins, Yedinak, Mandel,
  Beezley, and Lamb]{Kochanski:2016}
Kochanski, A.K.; Jenkins, M.A.; Yedinak, K.; Mandel, J.; Beezley, J.; Lamb, B.
\newblock {Toward an integrated system for fire, smoke, and air quality
  simulations}.
\newblock {\em Int. J. Wildland Fire} {\bf 2016}, {\em 25},~534–546.
\newblock {doi:10.1071/WF14074}.

\bibitem[Sethian(1999)]{Sethian:1999}
Sethian, J.A.
\newblock {\em Level Set Methods and Fast Marching Methods: Evolving Interfaces
  in Computational Geometry, Fluid Mechanics, Computer Vision, and Materials
  Science}; Cambridge University Press: Cambridge, United Kingdom, 1999.

\bibitem[Osher and Fedkiw(2006)]{Osher:2006}
Osher, S.; Fedkiw, R.
\newblock {\em Level Set Methods and Dynamic Implicit Surfaces}; Springer:  New York, United States, 2006.

\bibitem[Coen \em{et~al.}(2013)Coen, Cameron, Michalakes, Patton, Riggan, and
  Yedinak]{coen_2013}
Coen, J.L.; Cameron, M.; Michalakes, J.; Patton, E.G.; Riggan, P.J.; Yedinak,
  K.M.
\newblock WRF-Fire: Coupled Weather–Wildland Fire Modeling with the Weather
  Research and Forecasting Model.
\newblock {\em J. Appl. Meteorol. Climatol.} {\bf 2013}, {\em
  52},~16--38.
\newblock {doi:10.1175/JAMC-D-12-023.1}.

\bibitem[Mell \em{et~al.}(2007)Mell, Jenkins, Gould, and Cheney]{Mell:IJWF2007}
Mell, W.; Jenkins, M.; Gould, J.; Cheney, P.
\newblock {A physics-based approach to modelling grassland fires}.
\newblock {\em Int. J. Wildland Fire} {\bf 2007}, {\em
  16},~1--22.

\bibitem[Linn \em{et~al.}(2007)Linn, Winterkamp, Edminster, Colman, and
  Smith]{Linn:2007}
Linn, R.; Winterkamp, J.; Edminster, C.; Colman, J.J.; Smith, W.S.
\newblock {Coupled influences of topography and wind on wildland fire
  behaviour}.
\newblock {\em Int. J. Wildland Fire} {\bf 2007}, {\em 16},~183--195.

\bibitem[Arca \em{et~al.}(2019)Arca, G., M., M., and P.]{Arca_2019}
Arca, B.; Ghisu, T.; Casula, M.; Salis, M.; Duce, P.
\newblock A web-based wildfire simulator for operational applications.
\newblock {\em Int. J. Wildland Fire} {\bf 2019}, {\em
  28},~99--112.

\bibitem[Rehm and Baum(1978)]{Rehm:1}
Rehm, R.; Baum, H.
\newblock {The Equations of Motion for Thermally Driven, Buoyant Flows}.
\newblock {\em J. Res.  NBS} {\bf 1978}, {\em 83},~297--308.

\bibitem[McDermott and Floyd(2015)]{McDermott:2015}
McDermott, R.; Floyd, J.
\newblock {Enforcing realizability in explicit multi-component species
  transport}.
\newblock {\em Fire Saf. J.} {\bf 2015}, {\em 78},~180--187.

\bibitem[McDermott(2014)]{mcdermo_2014}
McDermott, R.J.
\newblock A velocity divergence constraint for large-eddy simulation of
  low-Mach flows.
\newblock {\em J. Comput. Phys.} {\bf 2014}, {\em
  274},~413--431.
\newblock {doi:10.1016/j.jcp.2014.06.019}.

\bibitem[Pope(2000)]{Pope:2000}
Pope, S.B.
\newblock {\em Turbulent Flows}; Cambridge University Press: Cambridge, United Kingdom, 2000.

\bibitem[Poinsot and Veynante(2005)]{Poinsot:TNC}
Poinsot, T.; Veynante, D.
\newblock {\em {Theoretical and Numerical Combustion}}, 2nd ed.; R.T. Edwards,
  Inc.: Philadelphia, PA, USA,  2005.

\bibitem[Deardorff(1980)]{Deardorff:1980}
Deardorff, J.
\newblock {Stratocumulus-capped mixed layers derived from a three-dimensional
  model}.
\newblock {\em Bound.-Layer Meteorol.} {\bf 1980}, {\em 18},~495--527.

\bibitem[Comte-{B}ellot and Corrsin(1971)]{CBC}
Comte-{B}ellot, G.; Corrsin, S.
\newblock Simple {Eulerian} time correlation of full- and narrow-band velocity
  signals in grid-generated, `isotropic' turbulence.
\newblock {\em J. Fluid Mech.} {\bf 1971}, {\em 48},~273--337.

\bibitem[Bardina \em{et~al.}(1980)Bardina, Ferziger, and
  Reynolds]{Bardina:1980}
Bardina, J.; Ferziger, J.H.; Reynolds, W.C.
\newblock {Improved Subgrid Scale Models for Large Eddy Simulation}.
\newblock   In Proceedings of the  AIAA 13th Fluid \& Plasma Dynamics Conference; American Institute of
  Aeronautics and Astronautics,  Snowmass, CO, USA,   14--16 July 1980. doi:10.2514/6.1980-1357.

\bibitem[Nicoud and Ducros(1999)]{Nicoud:1999}
Nicoud, F.; Ducros, F.
\newblock {Subgrid-Scale Stress Modelling Based on the Square of the Velocity
  Gradient Tensor}.
\newblock {\em Flow Turbul.  Combust.} {\bf 1999}, {\em
  62},~183--200.

\bibitem[Maragkos and Merci(2020)]{Maragkos:2020}
Maragkos, G.; Merci, B.
\newblock On the use of dynamic turbulence modelling in fire applications.
\newblock {\em Comb.  Flame} {\bf 2020}, {\em 216},~9--23.

\bibitem[Mulholland(2002)]{SFPE:Mulholland}
Mulholland, G.
\newblock Smoke Production and Properties. In {\em SFPE Handbook of Fire
  Protection Engineering}, 3rd ed.; National Fire Protection Association:
  Quincy, Massachusetts,  2002.

\bibitem[Magnussen and Hjertager(1977)]{Magnussen:1}
Magnussen, B.; Hjertager, B.
\newblock {On Mathematical Modeling of Turbulent Combustion with Special
  Emphasis on Soot Formation and Combustion}.
\newblock {\em Symposium (International) on Combustion} {\bf 1977}, {\em
  16:1},~719--729.

\bibitem[McDermott \em{et~al.}(2011)McDermott, McGrattan, and
  Floyd]{McDermott:2011}
McDermott, R.; McGrattan, K.; Floyd, J.
\newblock A Simple Reaction Time Scale for Under-Resolved Fire Dynamics.
\newblock  In Proceedings of the Fire Safety Science---10th International
  Symposium,  College Park, MD, USA,  19--24 June 2011; pp. 809--820.

\bibitem[McGrattan \em{et~al.}(2013)McGrattan, Hostikka, McDermott, Floyd,
  Weinschenk, and Overholt]{FDS_MathVal_Guides}
McGrattan, K.; Hostikka, S.; McDermott, R.; Floyd, J.; Weinschenk, C.;
  Overholt, K.
\newblock {\em {Fire Dynamics Simulator, Technical Reference Guide, Volume 1: Mathematical Model, Volume 3: Validation}}, 6th ed.;
\newblock National Institute of Standards and Technology: Gaithersburg,
  MD, USA;  VTT Technical Research Centre of Finland: Espoo, Finland, 2013.

\bibitem[Beyler(2016)]{SFPE:Beyler}
Beyler, C.
\newblock Flammability Limits of Premixed and Diffusion Flames. In {\em SFPE
  Handbook of Fire Protection Engineering}, 5th ed.;  \mbox{Hurley, M., Ed.;}
  Springer: New York, NY, USA, 2016.

\bibitem[Holman(1990)]{Holman:1}
Holman, J.
\newblock {\em Heat Transfer}, 7th ed.; McGraw-Hill: New York, NY, USA, 1990.

\bibitem[Incropera and De~Witt(1996)]{Incropera:1}
Incropera, F.P.; De~Witt, D.P.
\newblock {\em Fundamentals of Heat and Mass Transfer}, 4th ed.; John Wiley and
  Sons: New York,  NY, USA, 1996.

\bibitem[Berger(2017)]{berger_2016}
Berger, M.
\newblock Cut Cells: Meshes and Solvers.
\newblock {\em Handb. Numer. Anal.} {\bf 2017}, {\em 18},~1--22.
\newblock {doi:10.1016/bs.hna.2016.10.008}.

\bibitem[Eymard \em{et~al.}(2000)Eymard, Gallouet, and Herbin]{eymard_2000}
Eymard, R.; Gallouet, T.; Herbin, R.
\newblock Finite volume methods. In {\em Solution of Equation in Rn (Part 3),
  Techniques of Scientific Computing (Part 3)}; Handbook of Numerical Analysis;  Elsevier: Amsterdam, The Netherlands, 2000; Volume~7,  pp. 713 -- 1018.
\newblock {doi:10.1016/S1570-8659(00)07005-8}.

\bibitem[LeVeque(2002)]{leveque_2002}
\textls[-20]{LeVeque, R.
\newblock {\em Finite Volume Methods for Hyperbolic Problems}; Cambridge Texts
  in Applied Mathematics, Cambridge University Press: Cambridge, United Kingdom, 2002.}

\bibitem[Balaras(2004)]{balaras_2004}
Balaras, E.
\newblock Modeling complex boundaries using an external force field on fixed
  Cartesian grids in large-eddy simulations.
\newblock {\em Comput. Fluids} {\bf 2004}, {\em 33},~375--404.

\bibitem[LeVeque(2007)]{leveque2007finite}
LeVeque, R.
\newblock {\em Finite Difference Methods for Ordinary and Partial Differential
  Equations: Steady-State and Time-Dependent Problems}; \mbox{Other Titles} in Applied
  Mathematics, Society for Industrial and Applied Mathematics: Philadelphia, United States, 2007.

\bibitem[Kirkpatrick \em{et~al.}(2003)Kirkpatrick, Armfield, and
  Kent]{kirk_2003}
Kirkpatrick, M.; Armfield, S.; Kent, J.
\newblock A representation of curved boundaries for the solution of the
  Navier--Stokes equations on a staggered three-dimensional Cartesian grid.
\newblock {\em J. Comput. Phys.} {\bf 2003}, {\em 184},~1--36.
\newblock {doi:10.1016/S0021-9991(02)00013-X}.

\bibitem[Perot(1993)]{perot_1993}
Perot, J.
\newblock An Analysis of the Fractional Step Method.
\newblock {\em J. Comput. Phys.} {\bf 1993}, {\em 108},~51--58.

\bibitem[Fadlun \em{et~al.}(2000)Fadlun, Verzicco, Orlandi, and
  Mohd-Yusof]{fadlun_2000}
Fadlun, E.; Verzicco, R.; Orlandi, P.; Mohd-Yusof, J.
\newblock {Combined Immersed-Boundary Finite-Difference Methods for
  Three-Dimensional Complex Flow Simulations}.
\newblock {\em J. Comput. Phys.} {\bf 2000}, {\em 161},~35--60.

\bibitem[Perez-Ramirez \em{et~al.}(2017)Perez-Ramirez, Santoni, Tramoni,
  Bosseur, and Mell]{Perez-Ramirez:FT2017}
Perez-Ramirez, Y.; Santoni, P.; Tramoni, J.; Bosseur, F.; Mell, W.
\newblock {Examination of WFDS in Modeling Spreading Fires in a Furniture
  Calorimeter}.
\newblock {\em Fire Technol.} {\bf 2017}, {\em 53},~1795--1832.

\bibitem[Porterie \em{et~al.}(2005)Porterie, Consalvi, Kaiss, and
  Loraud]{Porterie:2006}
Porterie, B.; Consalvi, J.; Kaiss, A.; Loraud, J.
\newblock {Predicting Wildland Fire Behavior and Emissions Using a Fine-Scale
  Physical Model}.
\newblock {\em Numer. Heat Transf. Part A} {\bf 2005}, {\em 47},~571--591.

\bibitem[Morvan and Dupuy(2004)]{Morvan:CF2004}
Morvan, D.; Dupuy, J.
\newblock {Modeling the propagation of a wildfire through a Mediterranean shrub
  using a multiphase formulation}.
\newblock {\em Combust. Flame} {\bf 2004}, {\em 138},~199--210.

\bibitem[Houssami \em{et~al.}(2016)Houssami, Mueller, Filkov, Thomas,
  Skowronski, Gallagher, Clark, Kremens, and Simeoni]{Houssami:2016}
Houssami, M.; Mueller, E.; Filkov, A.; Thomas, J.; Skowronski, N.; Gallagher,
  M.; Clark, K.; Kremens, R.; Simeoni, A.
\newblock {Experimental procedures characterizing firebrand generation in
  wildland fires}.
\newblock {\em Fire Technol.} {\bf 2016}, {\em 52},~731--751.

\bibitem[Falkenstein-Smith \em{et~al.}(2019)Falkenstein-Smith, McGrattan,
  Toman, and Fernandez]{Falkenstein-Smith:2018}
Falkenstein-Smith, R.; McGrattan, K.; Toman, B.; Fernandez, M.
\newblock {Measurement of the Flow Resistance of Vegetation}.
\newblock In {\em NIST Technical Note 2039}; National Institute of Standards and
  Technology: Gaithersburg, MD, USA,  2019.
\newblock
  \href{doi:10.6028/NIST.TN.2039}{doi:10.6028/NIST.TN.2039}.

\bibitem[Cheney \em{et~al.}(1993)Cheney, Gould, and Catchpole]{Cheney:IJWF1998}
Cheney, N.; Gould, J.; Catchpole, W.
\newblock {Predition of Fire-Spread in Grasslands}.
\newblock {\em Int. J. Wildland Fire} {\bf 1993}, {\em
  8},~1--13.

\bibitem[Mell \em{et~al.}(2019)Mell, Hoffman, and Ziegler]{Mell:FBFC2019}
Mell, W.; Hoffman, C.; Ziegler, J.
\newblock The state of physics-based fire modeling and an example of a reduced
  fire-physics model.
\newblock   In~Proceedings of the 6th International Fire Behavior and Fuels Conference; International
  Association of Wildland Fire, Missoula, MT, USA, 29 April--3 May  2019.

\bibitem[Albini(1976)]{Albini:1976}
Albini, F.
\newblock {Estimating Wildfire Behavior and Effects}.
\newblock In \emph{Research Paper INT-30};  Intermountain Forest and Range Experiment
  Station, USDA Forest Service: Ogden, Utah,  1976.
\newblock
   {Available online: \url{https://www.fs.fed.us/rm/pubs_int/int_gtr030.pdf} (accessed on 05/02/2021).}

\bibitem[Wilson(1980)]{Wilson:1980}
Wilson, R.
\newblock {Reformulation of forest fire spread equations in SI units}.
\newblock In \emph{Research Note INT-292}; Intermountain Forest and Range Experiment
  Station, USDA Forest Service: Ogden, Utah,  1980.
\newblock
   {Available online: \url{https://www.fs.usda.gov/treesearch/pubs/33592} (accessed on 05/02/2021).}

\bibitem[Jarrin(2008)]{Jarrin:2008}
Jarrin, N.
\newblock Synthetic Inflow Boundary Conditions for the Numerical Simulation of
  Turbulence.
\newblock Ph.D. Thesis, The University of Manchester, Manchester, UK,  2008.

\bibitem[Dyer(1974)]{Dyer:1974}
Dyer, A.
\newblock {A review of flux profile relationships}.
\newblock {\em Bound.-Layer Meteorol.} {\bf 1974}, {\em 7},~363--372.

\bibitem[CAW()]{CAWFE}
{\em CAWFE: Coupled Atmosphere-Wildland Fire Environment};
\newblock National Center for Atmospheric Research. Available Online: \url{https://www.openwfm.org/wiki/CAWFE}  (accessed on 05/02/2021).%please add publisher location and year.

\bibitem[Cheney \em{et~al.}(1993)Cheney, Gould, and Catchpole]{Cheney:IJWF1993}
Cheney, N.; Gould, J.; Catchpole, W.
\newblock {The Influence of Fuel, Weather and Fire Shape Variables on
  Fire-Spread in Grasslands}.
\newblock {\em Int. J. Wildland Fire} {\bf 1993}, {\em
  3},~31--44.

\bibitem[Arca \em{et~al.}(2007)Arca, P., M., G., M., and D.]{Arca_2007}
Arca, B.; Duce, P.; Laconi, M.; Pellizzaro, G.; Salis, M.; Spano D.
\newblock Evaluation of FARSITE simulator Mediterranean marquis.
\newblock {\em Int. J. Wildland Fire} {\bf 2007}, {\em
  16},~563--573.

\bibitem[Susott(1982)]{Susott:FS1982}
Susott, R.
\newblock {Characterization of the thermal properties of forest fuels by
  combustible gas analysis}.
\newblock {\em For. Sci.} {\bf 1982}, {\em 2},~404--420.

\bibitem[Khan \em{et~al.}(2016)Khan, Tewarson, and Chaos]{SFPE:Tewarson}
Khan, M.; Tewarson, A.; Chaos, M., {Combustion Characteristics of Materials and
  Generation of Fire Products}.
\newblock In {\em SFPE Handbook of Fire Protection Engineering}, 5th ed.;
  Springer: New York, NY, USA,  2016; Chapter~36, pp. 1143--1232.

\bibitem[Farouki(1981)]{Farouki:1981}
Farouki, O.
\newblock {Thermal properties of soils}.
\newblock In \emph{CRREL Monograph 81-1}; U.S. Army Corps of Engineers, Cold Regions
  Research and Engineering Laboratory: Hanover, New Hampshire,  1981.

\bibitem[McGrattan(2017)]{McGrattan:CI2017}
McGrattan, K.
\newblock {Progress in Modeling Wildland Fires using Computational Fluid
  Dynamics}.
\newblock  In Proceedings of the 10th U.S. National Combustion Meeting. Combustion
  Institute, Pittsburgh, Pennsylvania,  College Park, MD, USA,  23--26 April 2017.

\bibitem[QGI()]{QGIS}
 {Available online: \url{https://qgis.org/en/site/} (accessed on 05/02/2021).} %Please add accessed date and title

\bibitem[qgi()]{qgis2fds}
 {Available online: \url{https://github.com/firetools/qgis2fds/wiki/} (accessed on 05/02/2021).}

\bibitem[Trucchia \em{et~al.}(2020)Trucchia, D’Andrea, Baghino, Fiorucci,
  Ferraris, Negro, Gollini, and Severino]{Trucchia:2020}
Trucchia, A.; D’Andrea, M.; Baghino, F.; Fiorucci, P.; Ferraris, L.; Negro,
  D.; Gollini, A.; Severino, M.
\newblock PROPAGATOR: An Operational Cellular-Automata Based Wildfire
  Simulator.
\newblock {\em Fire} {\bf 2020}, {\em 3},~26.
\newblock {doi:10.3390/fire3030026}.

\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

https://www.overleaf.com/project/5f9825767d349c000144f040

### 2021-02-11 atmosphere-1087074-final.tex
Reorder refs

@@
 
-In the formal derivation of the LES equations (see, e.g,~\cite{Pope:2000,FDS_MathVal_Guides}) the filtered nonlinear terms (advection terms, mean chemical source term, and~the unresolved boundary fluxes) require subgrid-scale parameterizations.
+In the formal derivation of the LES equations (see, e.g,~\cite{Pope:2000,FDS_Tech_Guide}) the filtered nonlinear terms (advection terms, mean chemical source term, and~the unresolved boundary fluxes) require subgrid-scale parameterizations.
 In~this section, LES filter notation, as well as Cartesian tensor index notation, are used to clarify the terms used in the governing~equations.
@@
 Here, $D_F$ is the binary diffusivity of the fuel species in air, $k_{sgs}$ is the modeled subgrid kinetic energy per unit mass discussed above, and~$g$ is the gravitational acceleration.
-The~constant $C_u=0.4$ is determined by matching flame height correlations across a wide range of fire Froude numbers~\cite{FDS_MathVal_Guides}.
+The~constant $C_u=0.4$ is determined by matching flame height correlations across a wide range of fire Froude numbers~\cite{FDS_Tech_Guide}.
 
@@
 The~forced convection Nusselt number (Nu) depends on the geometry and flow characteristics~\cite{Holman:1,Incropera:1}.
-Further details are provided in~\cite{FDS_MathVal_Guides}.
+Further details are provided in~\cite{FDS_Tech_Guide}.
 
@@
 
+%\bibliography{Review_Process/FDS_refs,Review_Process/LOFM_Workshop_References,Review_Process/Vanella_Article_Atmosphere_2020}
+
 \begin{thebibliography}{999}
@@
 Thomas, D.; Butry, D.; Gilbert, S.; Webb, D.; Fung, J.
-\newblock  {The Costs and Losses of Wildfires: A Literature Review}.
-\newblock In \emph{Special Publication NIST SP-1215}; National Institute of Standards
-  and Technology: Gaithersburg, MD, USA,  2017.
-\newblock
-   {Available online: \url{https://www.nist.gov/publications/costs-and-losses-wildfires} (accessed on 05/02/2021).}
+\newblock The Costs and Losses of Wildfires: A Literature Review.
+\newblock Special Publication NIST SP - 1215, National Institute of Standards
+  and Technology, Gaithersburg, Mayland,  2017.
+\newblock
+  \href{https://www.nist.gov/publications/costs-and-losses-wildfires}{https://www.nist.gov/publications/costs-and-losses-wildfires} (accessed on 05/02/2021).
 
@@
 McDermott, R.; Bryner, N.; Heintz, J.
-\newblock  {Large Outdoor Fire Modeling (LOFM) Workshop Summary Report}.
-\newblock In \emph{Special Publication NIST SP-1245}; National Institute of Standards
-  and Technology: Gaithersburg, MD, USA,  2019.
-\newblock
-    Available online: \url{https://www.nist.gov/publications/large-outdoor-fire-modeling-lofm-workshop-summary-report} (accessed on 05/02/2021).
+\newblock Large Outdoor Fire Modeling (LOFM) Workshop Summary Report.
+\newblock Special Publication NIST SP - 1245, National Institute of Standards
+  and Technology, Gaithersburg, Mayland,  2019.
+\newblock
+  \href{https://www.nist.gov/publications/large-outdoor-fire-modeling-lofm-workshop-summary-report}{https://www.nist.gov/publications/large-outdoor-fire-modeling-lofm-workshop-summary-report} (accessed on 05/02/2021).
 
@@
 Richards, L.; Brew, N.; Smith, L.
-\newblock  {2019–20 Australian bushfires—frequently asked questions: A Quick
-  Guide}.
-\newblock In \emph{Research Paper Series 2019-20}; Parliament of
-  Australia: Canberra, Australia,  2020. ISSN 2203-5249. 
-\newblock
-   {Available online: \url{https://parlinfo.aph.gov.au/parlInfo/download/library/prspub/7234762/upload\_binary/7234762.pdf} (accessed on 05/02/2021).}
+\newblock 2019–20 Australian bushfires—frequently asked questions: a quick
+  guide.
+\newblock Research Paper Series 2019-20, ISSN 2203-5249, Parliament of
+  Australia, Canberra, Australia,  2020.
+\newblock
+  \href{https://parlinfo.aph.gov.au/parlInfo/download/library/prspub/7234762/upload\_binary/7234762.pdf}{https://parlinfo.aph.gov.au/parlInfo/download/library/prspub/7234762/upload\_binary/7234762.pdf} (accessed on 05/02/2021).
 
@@
 \newblock A review of a new generation of wildfire–atmosphere modeling.
-\newblock {\em IEEE Syst. J.} {\bf 2011}, {\em 5},~233--243.
+\newblock {\em IEEE Systems Journal} {\bf 2011}, {\em 5},~233--243.
 
@@
 \newblock A review of a new generation of wildfire–atmosphere modeling.
-\newblock {\em Can. J. For. Res.} {\bf 2019}, {\em
+\newblock {\em Canadian Journal of Forest Research} {\bf 2019}, {\em
   49},~565--574.
@@
 \newblock A mathematical model for predicting fire spread in wildland fuels.
-\newblock In \emph{Technical Report INT-115}; USDA Forest Service: Ogden, Utah,  1972.
-\newblock
-   {Available online: \url{http://www.treesearch.fs.fed.us/pubs/32533} (accessed on 05/02/2021).}
+\newblock Technical Report INT-115, USDA Forest Service, Ogden, Utah,  1972.
+\newblock
+  \href{http://www.treesearch.fs.fed.us/pubs/32533}{http://www.treesearch.fs.fed.us/pubs/32533} (accessed on 05/02/2021).
 
@@
 \newblock {Aids to Determining Fuel Models for Estimating Fire Behavior}.
-\newblock In \emph{General Technical Report INT-122}; Intermountain Forest and Range
-  Experiment Station, USDA Forest Service: Ogden, Utah,  1982.
-\newblock
-   {Available online: \url{https://www.fs.usda.gov/treesearch/pubs/6447} (accessed on 05/02/2021).}
+\newblock General Technical Report INT-122, Intermountain Forest and Range
+  Experiment Station, USDA Forest Service, Ogden, Utah,  1982.
+\newblock
+  \href{https://www.fs.usda.gov/treesearch/pubs/6447}{https://www.fs.usda.gov/treesearch/pubs/6447} (accessed on 05/02/2021).
 
@@
 Finney, M.
-\newblock {Fire Area Simulator–Model, Development and Evaluation}.
-\newblock In \emph{Research Paper RMRS-RP-4 Revised}; United States Forest Service, Rocky
+\newblock {Fire Area Simulator – Model, Development and Evaluation}.
+\newblock Research Paper RMRS-RP-4 Revised, United States Forest Service, Rocky
   Mountain Research Station, Missoula, Montana,  2004.
 \newblock
-   {Available online: \url{http://www.fs.fed.us/rm/pubs/rmrs\_rp004.pdf} (accessed on 05/02/2021).}
+  \href{http://www.fs.fed.us/rm/pubs/rmrs\_rp004.pdf}{http://www.fs.fed.us/rm/pubs/rmrs\_rp004.pdf} (accessed on 05/02/2021).
 
@@
   wildland fire front propagation}.
-\newblock {\em Int. J. Wildland Fire} {\bf 2015}, {\em
+\newblock {\em International Journal of Wildland Fire} {\bf 2015}, {\em
   25},~229--241.
-\newblock{doi:10.1071/WF13178}.
+\newblock
+  \href{http://dx.doi.org/10.1071/WF13178}{http://dx.doi.org/10.1071/WF13178} (accessed on 05/02/2021).
 
@@
   Overholt, K.
-\newblock {\em {Fire Dynamics Simulator, User's Guide}}, 6th ed.;
-\newblock National Institute of Standards and Technology: Gaithersburg,
-  MD, USA; VTT Technical Research Centre of Finland: Espoo, Finland, 2013.
+\newblock {\em {Fire Dynamics Simulator, User's Guide}}.
+\newblock National Institute of Standards and Technology, Gaithersburg,
+  Maryland, USA, and VTT Technical Research Centre of Finland, Espoo, Finland,
+  sixth ed.,  2013.
 
@@
   Atmosphere-Wildland Fire Environment Model (CAWFE)}.
-\newblock In~\emph{Technical Report NCAR/TN500+STR}; National Center for Atmospheric
-  Research: Boulder, CO, USA, 2013.
+\newblock Technical Report NCAR/TN500+STR, National Center for Atmospheric
+  Research, Boulder, CO,  2013.
 
@@
   automated calibration.
-\newblock {\em Fire Saf. J.} {\bf 2013}, {\em 62},~\mbox{289--298}.
-\newblock {doi:10.1016/j.firesaf.2013.08.014}.
+\newblock {\em Fire Safety Journal} {\bf 2013}, {\em 62},~289 -- 298.
+\newblock
+  \href{https://doi.org/10.1016/j.firesaf.2013.08.014}{https://doi.org/10.1016/j.firesaf.2013.08.014} (accessed on 05/02/2021).
+
+\bibitem[Coen \em{et~al.}(2013)Coen, Cameron, Michalakes, Patton, Riggan, and
+  Yedinak]{coen_2013}
+Coen, J.L.; Cameron, M.; Michalakes, J.; Patton, E.G.; Riggan, P.J.; Yedinak,
+  K.M.
+\newblock WRF-Fire: Coupled Weather–Wildland Fire Modeling with the Weather
+  Research and Forecasting Model.
+\newblock {\em Journal of Applied Meteorology and Climatology} {\bf 2013}, {\em
+  52},~16--38.
+\newblock
+  \href{https://doi.org/10.1175/JAMC-D-12-023.1}{https://doi.org/10.1175/JAMC-D-12-023.1} (accessed on 05/02/2021).
 
@@
   coupled atmosphere-surface models}.
-\newblock {\em IEEE Control  Syst. Mag.} {\bf 2009}, {\em 9},~47–65.
-\newblock {doi:10.1109/MCS.2009.932224}.
+\newblock {\em IEEE Control Systems Magazine} {\bf 2009}, {\em 9},~47–65.
+\newblock
+  \href{https://doi.org/10.1109/MCS.2009.932224}{https://doi.org/10.1109/MCS.2009.932224} (accessed on 05/02/2021).
 
@@
   2011}.
-\newblock {\em \mbox{Geosci. Model} Dev.} {\bf 2011}, {\em 4},~591--610.
-\newblock {doi:10.5194/gmd-4-591-2011}.
+\newblock {\em Geosci. Model Dev.} {\bf 2011}, {\em 4},~591--610.
+\newblock
+  \href{https://doi.org/10.5194/gmd-4-591-2011}{https://doi.org/10.5194/gmd-4-591-2011} (accessed on 05/02/2021).
 
@@
 \newblock {\em Nat. Hazards Earth Syst. Sci.} {\bf 2015}, {\em 14},~2829--2845.
-\newblock {doi:10.5194/nhess-14-2829-2014}.
+\newblock
+  \href{https://doi.org/10.5194/nhess-14-2829-2014}{https://doi.org/10.5194/nhess-14-2829-2014} (accessed on 05/02/2021).
 
@@
 \newblock {\em Int. J. Wildland Fire} {\bf 2016}, {\em 25},~534–546.
-\newblock {doi:10.1071/WF14074}.
+\newblock
+  \href{https://doi.org/10.1071/WF14074}{https://doi.org/10.1071/WF14074} (accessed on 05/02/2021).
 
@@
   in Computational Geometry, Fluid Mechanics, Computer Vision, and Materials
-  Science}; Cambridge University Press: Cambridge, United Kingdom, 1999.
+  Science}; Cambridge,  1999.
 
@@
 Osher, S.; Fedkiw, R.
-\newblock {\em Level Set Methods and Dynamic Implicit Surfaces}; Springer:  New York, United States, 2006.
-
-\bibitem[Coen \em{et~al.}(2013)Coen, Cameron, Michalakes, Patton, Riggan, and
-  Yedinak]{coen_2013}
-Coen, J.L.; Cameron, M.; Michalakes, J.; Patton, E.G.; Riggan, P.J.; Yedinak,
-  K.M.
-\newblock WRF-Fire: Coupled Weather–Wildland Fire Modeling with the Weather
-  Research and Forecasting Model.
-\newblock {\em J. Appl. Meteorol. Climatol.} {\bf 2013}, {\em
-  52},~16--38.
-\newblock {doi:10.1175/JAMC-D-12-023.1}.
+\newblock {\em Level Set Methods and Dynamic Implicit Surfaces}; Springer,
+  2006.
 
@@
 \newblock {A physics-based approach to modelling grassland fires}.
-\newblock {\em Int. J. Wildland Fire} {\bf 2007}, {\em
+\newblock {\em International Journal of Wildland Fire} {\bf 2007}, {\em
   16},~1--22.
@@
 \bibitem[Arca \em{et~al.}(2019)Arca, G., M., M., and P.]{Arca_2019}
-Arca, B.; Ghisu, T.; Casula, M.; Salis, M.; Duce, P.
+Arca, B.; G., T.; M., C.; M., S.; P., D.
 \newblock A web-based wildfire simulator for operational applications.
-\newblock {\em Int. J. Wildland Fire} {\bf 2019}, {\em
+\newblock {\em International Journal of Wildland Fire} {\bf 2019}, {\em
   28},~99--112.
@@
 \newblock {The Equations of Motion for Thermally Driven, Buoyant Flows}.
-\newblock {\em J. Res.  NBS} {\bf 1978}, {\em 83},~297--308.
+\newblock {\em Journal of Research of the NBS} {\bf 1978}, {\em 83},~297--308.
 
@@
   transport}.
-\newblock {\em Fire Saf. J.} {\bf 2015}, {\em 78},~180--187.
+\newblock {\em Fire Safety Journal} {\bf 2015}, {\em 78},~180--187.
 
@@
   low-Mach flows.
-\newblock {\em J. Comput. Phys.} {\bf 2014}, {\em
+\newblock {\em Journal of Computational Physics} {\bf 2014}, {\em
   274},~413--431.
-\newblock {doi:10.1016/j.jcp.2014.06.019}.
+\newblock
+  \href{http://dx.doi.org/10.1016/j.jcp.2014.06.019}{http://dx.doi.org/10.1016/j.jcp.2014.06.019} (accessed on 05/02/2021).
 
@@
 Pope, S.B.
-\newblock {\em Turbulent Flows}; Cambridge University Press: Cambridge, United Kingdom, 2000.
+\newblock {\em Turbulent Flows}; Cambridge University Press,  2000.
+
+\bibitem[McGrattan \em{et~al.}(2013)McGrattan, Hostikka, McDermott, Floyd,
+  Weinschenk, and Overholt]{FDS_Tech_Guide}
+McGrattan, K.; Hostikka, S.; McDermott, R.; Floyd, J.; Weinschenk, C.;
+  Overholt, K.
+\newblock {\em {Fire Dynamics Simulator, Technical Reference Guide}}.
+\newblock National Institute of Standards and Technology, Gaithersburg,
+  Maryland, USA, and VTT Technical Research Centre of Finland, Espoo, Finland,
+  sixth ed.,  2013.
+\newblock Vol.~1: Mathematical Model; Vol.~2: Verification Guide; Vol.~3:
+  Validation Guide; Vol.~4: Software Quality Assurance.
 
@@
 \newblock {\em {Theoretical and Numerical Combustion}}, 2nd ed.; R.T. Edwards,
-  Inc.: Philadelphia, PA, USA,  2005.
+  Inc.: Philadelphia, Pennsylvania,  2005.
 
@@
   model}.
-\newblock {\em Bound.-Layer Meteorol.} {\bf 1980}, {\em 18},~495--527.
+\newblock {\em Boundary-Layer Meteorol.} {\bf 1980}, {\em 18},~495--527.
 
@@
 \newblock {Improved Subgrid Scale Models for Large Eddy Simulation}.
-\newblock   In Proceedings of the  AIAA 13th Fluid \& Plasma Dynamics Conference; American Institute of
-  Aeronautics and Astronautics,  Snowmass, CO, USA,   14--16 July 1980. doi:10.2514/6.1980-1357.
+\newblock  AIAA 13th Fluid \& Plasma Dynamics Conference; American Institute of
+  Aeronautics and Astronautics, ,  1980; AIAA-80-1357.
 
@@
   Gradient Tensor}.
-\newblock {\em Flow Turbul.  Combust.} {\bf 1999}, {\em
+\newblock {\em Flow, Turbulence, and Combustion} {\bf 1999}, {\em
   62},~183--200.
@@
 \newblock On the use of dynamic turbulence modelling in fire applications.
-\newblock {\em Comb.  Flame} {\bf 2020}, {\em 216},~9--23.
+\newblock {\em Comb. and Flame} {\bf 2020}, {\em 216},~9--23.
 
@@
   Emphasis on Soot Formation and Combustion}.
-\newblock {\em Symposium (International) on Combustion} {\bf 1977}, {\em
-  16:1},~719--729.
+\newblock  Proceedings of the Sixteenth Symposium (International) on
+  Combustion. Combustion Institute, Pittsburgh, Pennsylvania,  1977, pp.
+  719--729.
 
@@
 \newblock A Simple Reaction Time Scale for Under-Resolved Fire Dynamics.
-\newblock  In Proceedings of the Fire Safety Science---10th International
-  Symposium,  College Park, MD, USA,  19--24 June 2011; pp. 809--820.
-
-\bibitem[McGrattan \em{et~al.}(2013)McGrattan, Hostikka, McDermott, Floyd,
-  Weinschenk, and Overholt]{FDS_MathVal_Guides}
-McGrattan, K.; Hostikka, S.; McDermott, R.; Floyd, J.; Weinschenk, C.;
-  Overholt, K.
-\newblock {\em {Fire Dynamics Simulator, Technical Reference Guide, Volume 1: Mathematical Model, Volume 3: Validation}}, 6th ed.;
-\newblock National Institute of Standards and Technology: Gaithersburg,
-  MD, USA;  VTT Technical Research Centre of Finland: Espoo, Finland, 2013.
+\newblock  Fire Safety Science -- Proceedings of the 10th International
+  Symposium; ,  2011; pp. 809--820.
 
@@
 \newblock Flammability Limits of Premixed and Diffusion Flames. In {\em SFPE
-  Handbook of Fire Protection Engineering}, 5th ed.;  \mbox{Hurley, M., Ed.;}
-  Springer: New York, NY, USA, 2016.
+  Handbook of Fire Protection Engineering}, 5th ed.;  Hurley, M., Ed.;
+  Springer: New York,  2016.
 
@@
 Holman, J.
-\newblock {\em Heat Transfer}, 7th ed.; McGraw-Hill: New York, NY, USA, 1990.
+\newblock {\em Heat Transfer}, 7th ed.; McGraw-Hill: New York,  1990.
 
@@
 \newblock {\em Fundamentals of Heat and Mass Transfer}, 4th ed.; John Wiley and
-  Sons: New York,  NY, USA, 1996.
+  Sons: New York,  1996.
 
@@
 \newblock Cut Cells: Meshes and Solvers.
-\newblock {\em Handb. Numer. Anal.} {\bf 2017}, {\em 18},~1--22.
-\newblock {doi:10.1016/bs.hna.2016.10.008}.
+\newblock {\em Handbook of Numerical Analysis} {\bf 2017}, {\em 18},~1--22.
+\newblock
+  \href{https://doi.org/10.1016/bs.hna.2016.10.008}{https://doi.org/10.1016/bs.hna.2016.10.008} (accessed on 05/02/2021).
 
@@
 \newblock Finite volume methods. In {\em Solution of Equation in Rn (Part 3),
-  Techniques of Scientific Computing (Part 3)}; Handbook of Numerical Analysis;  Elsevier: Amsterdam, The Netherlands, 2000; Volume~7,  pp. 713 -- 1018.
-\newblock {doi:10.1016/S1570-8659(00)07005-8}.
+  Techniques of Scientific Computing (Part 3)}; Elsevier,  2000; Vol.~7, {\em
+  Handbook of Numerical Analysis}, pp. 713 -- 1018.
+\newblock
+  \href{https://doi.org/10.1016/S1570-8659(00)07005-8}{https://doi.org/10.1016/S1570-8659(00)07005-8} (accessed on 05/02/2021).
 
 \bibitem[LeVeque(2002)]{leveque_2002}
-\textls[-20]{LeVeque, R.
+LeVeque, R.
 \newblock {\em Finite Volume Methods for Hyperbolic Problems}; Cambridge Texts
-  in Applied Mathematics, Cambridge University Press: Cambridge, United Kingdom, 2002.}
+  in Applied Mathematics, Cambridge University Press,  2002.
 
@@
   Cartesian grids in large-eddy simulations.
-\newblock {\em Comput. Fluids} {\bf 2004}, {\em 33},~375--404.
+\newblock {\em Computers \& Fluids} {\bf 2004}, {\em 33},~375 -- 404.
 
@@
 \newblock {\em Finite Difference Methods for Ordinary and Partial Differential
-  Equations: Steady-State and Time-Dependent Problems}; \mbox{Other Titles} in Applied
-  Mathematics, Society for Industrial and Applied Mathematics: Philadelphia, United States, 2007.
+  Equations: Steady-State and Time-Dependent Problems}; Other Titles in Applied
+  Mathematics, Society for Industrial and Applied Mathematics,  2007.
 
@@
   Navier--Stokes equations on a staggered three-dimensional Cartesian grid.
-\newblock {\em J. Comput. Phys.} {\bf 2003}, {\em 184},~1--36.
-\newblock {doi:10.1016/S0021-9991(02)00013-X}.
+\newblock {\em Journal of Computational Physics} {\bf 2003}, {\em 184},~1 --
+  36.
+\newblock
+  \href{https://doi.org/10.1016/S0021-9991(02)00013-X}{https://doi.org/10.1016/S0021-9991(02)00013-X} (accessed on 05/02/2021).
 
@@
 \newblock An Analysis of the Fractional Step Method.
-\newblock {\em J. Comput. Phys.} {\bf 1993}, {\em 108},~51--58.
+\newblock {\em Journal of Computational Physics} {\bf 1993}, {\em 108},~51 --
+  58.
 
@@
   Three-Dimensional Complex Flow Simulations}.
-\newblock {\em J. Comput. Phys.} {\bf 2000}, {\em 161},~35--60.
+\newblock {\em Journal of Computational Physics} {\bf 2000}, {\em 161},~35--60.
 
@@
   Calorimeter}.
-\newblock {\em Fire Technol.} {\bf 2017}, {\em 53},~1795--1832.
+\newblock {\em Fire Technology} {\bf 2017}, {\em 53},~1795--1832.
 
@@
   Physical Model}.
-\newblock {\em Numer. Heat Transf. Part A} {\bf 2005}, {\em 47},~571--591.
+\newblock {\em Numerical Heat Transfer, Part A} {\bf 2005}, {\em 47},~571--591.
 
@@
   using a multiphase formulation}.
-\newblock {\em Combust. Flame} {\bf 2004}, {\em 138},~199--210.
+\newblock {\em Combustion and Flame} {\bf 2004}, {\em 138},~199--210.
 
@@
   wildland fires}.
-\newblock {\em Fire Technol.} {\bf 2016}, {\em 52},~731--751.
+\newblock {\em Fire Technology} {\bf 2016}, {\em 52},~731--751.
 
@@
 \newblock {Measurement of the Flow Resistance of Vegetation}.
-\newblock In {\em NIST Technical Note 2039}; National Institute of Standards and
-  Technology: Gaithersburg, MD, USA,  2019.
-\newblock
-  \href{doi:10.6028/NIST.TN.2039}{doi:10.6028/NIST.TN.2039}.
+\newblock {NIST Technical Note} 2039, National Institute of Standards and
+  Technology, Gaithersburg, Maryland,  2019.
+\newblock
+  \href{https://doi.org/10.6028/NIST.TN.2039}{https://doi.org/10.6028/NIST.TN.2039} (accessed on 05/02/2021).
 
@@
 \newblock {Predition of Fire-Spread in Grasslands}.
-\newblock {\em Int. J. Wildland Fire} {\bf 1993}, {\em
+\newblock {\em International Journal of Wildland Fire} {\bf 1993}, {\em
   8},~1--13.
@@
   fire-physics model.
-\newblock   In~Proceedings of the 6th International Fire Behavior and Fuels Conference; International
-  Association of Wildland Fire, Missoula, MT, USA, 29 April--3 May  2019.
+\newblock  6th International Fire Behavior and Fuels Conference; International
+  Association of Wildland Fire, ,  2019.
 
@@
 \newblock {Estimating Wildfire Behavior and Effects}.
-\newblock In \emph{Research Paper INT-30};  Intermountain Forest and Range Experiment
-  Station, USDA Forest Service: Ogden, Utah,  1976.
-\newblock
-   {Available online: \url{https://www.fs.fed.us/rm/pubs_int/int_gtr030.pdf} (accessed on 05/02/2021).}
+\newblock Research Paper INT-30, Intermountain Forest and Range Experiment
+  Station, USDA Forest Service, Ogden, Utah,  1976.
+\newblock
+  \href{https://www.fs.fed.us/rm/pubs_int/int_gtr030.pdf}{https://www.fs.fed.us/rm/pubs$\_$int/int$\_$gtr030.pdf} (accessed on 05/02/2021).
 
@@
 \newblock {Reformulation of forest fire spread equations in SI units}.
-\newblock In \emph{Research Note INT-292}; Intermountain Forest and Range Experiment
-  Station, USDA Forest Service: Ogden, Utah,  1980.
-\newblock
-   {Available online: \url{https://www.fs.usda.gov/treesearch/pubs/33592} (accessed on 05/02/2021).}
+\newblock Research Note INT-292, Intermountain Forest and Range Experiment
+  Station, USDA Forest Service, Ogden, Utah,  1980.
+\newblock
+  \href{https://www.fs.usda.gov/treesearch/pubs/33592}{https://www.fs.usda.gov/treesearch/pubs/33592} (accessed on 05/02/2021).
 
@@
   Turbulence.
-\newblock Ph.D. Thesis, The University of Manchester, Manchester, UK,  2008.
+\newblock PhD thesis, The University of Manchester, Manchester M60 1QD, United
+  Kingdom,  2008.
 
@@
 \newblock {A review of flux profile relationships}.
-\newblock {\em Bound.-Layer Meteorol.} {\bf 1974}, {\em 7},~363--372.
+\newblock {\em Boundary-Layer Meteorology} {\bf 1974}, {\em 7},~363--372.
 
 \bibitem[CAW()]{CAWFE}
-{\em CAWFE: Coupled Atmosphere-Wildland Fire Environment};
-\newblock National Center for Atmospheric Research. Available Online: \url{https://www.openwfm.org/wiki/CAWFE}  (accessed on 05/02/2021).%please add publisher location and year.
+{Coupled Atmosphere-Wildland Fire Environmant (CAWFE)}.
+\newblock National Center for Atmospheric Research.
 
@@
   Fire-Spread in Grasslands}.
-\newblock {\em Int. J. Wildland Fire} {\bf 1993}, {\em
+\newblock {\em International Journal of Wildland Fire} {\bf 1993}, {\em
   3},~31--44.
@@
 \bibitem[Arca \em{et~al.}(2007)Arca, P., M., G., M., and D.]{Arca_2007}
-Arca, B.; Duce, P.; Laconi, M.; Pellizzaro, G.; Salis, M.; Spano D.
+Arca, B.; P., D.; M., L.; G., P.; M., S.; D., S.
 \newblock Evaluation of FARSITE simulator Mediterranean marquis.
-\newblock {\em Int. J. Wildland Fire} {\bf 2007}, {\em
+\newblock {\em International Journal of Wildland Fire} {\bf 2007}, {\em
   16},~563--573.
@@
   combustible gas analysis}.
-\newblock {\em For. Sci.} {\bf 1982}, {\em 2},~404--420.
+\newblock {\em Forest Science} {\bf 1982}, {\em 2},~404--420.
 
@@
 \newblock In {\em SFPE Handbook of Fire Protection Engineering}, 5th ed.;
-  Springer: New York, NY, USA,  2016; Chapter~36, pp. 1143--1232.
+  Springer: New York,  2016; chapter~36, pp. 1143--1232.
 
@@
 \newblock {Thermal properties of soils}.
-\newblock In \emph{CRREL Monograph 81-1}; U.S. Army Corps of Engineers, Cold Regions
-  Research and Engineering Laboratory: Hanover, New Hampshire,  1981.
+\newblock CRREL Monograph 81-1, U.S. Army Corps of Engineers, Cold Regions
+  Research and Engineering Laboratory, Hanover, New Hampshire,  1981.
 
@@
   Dynamics}.
-\newblock  In Proceedings of the 10th U.S. National Combustion Meeting. Combustion
-  Institute, Pittsburgh, Pennsylvania,  College Park, MD, USA,  23--26 April 2017.
+\newblock  Proceedings of the 10th U.S. National Combustion Meeting. Combustion
+  Institute, Pittsburgh, Pennsylvania,  2017.
 
 \bibitem[QGI()]{QGIS}
- {Available online: \url{https://qgis.org/en/site/} (accessed on 05/02/2021).} %Please add accessed date and title
+\href{https://qgis.org/en/site/}{https://qgis.org/en/site/} (accessed on 05/02/2021).
 
 \bibitem[qgi()]{qgis2fds}
- {Available online: \url{https://github.com/firetools/qgis2fds/wiki/} (accessed on 05/02/2021).}
+\href{https://github.com/firetools/qgis2fds/wiki/}{https://github.com/firetools/qgis2fds/wiki/} (accessed on 05/02/2021).
 
@@
 \newblock {\em Fire} {\bf 2020}, {\em 3},~26.
-\newblock {doi:10.3390/fire3030026}.
+\newblock
+  \href{https://doi.org/10.3390/fire3030026}{https://doi.org/10.3390/fire3030026} (accessed on 05/02/2021).
 
@@
 
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}
-
-https://www.overleaf.com/project/5f9825767d349c000144f040
